---
title: "Spatiotemporal Variation of Adult C. californicus Shell Allometry"
author: "Malloy and Ul-Hasan and Lewis et al"
date: "Apr 9, 2019"
output:
  word_document: default
  pdf_document: default
---

Last edited [Sabah Ul-Hasan]: April 9, 2019

https://github.com/MichaelMalloy/C_californicus_Morphology

Contributors of data by order of contribution: MM, SUH, CAO, SC
Contributors of code by order of contribution: SUH, LL, MM

SuppTabs 2-4    Data, post-processing: Raw -> Clean -> SoCal  Lines 22 - 144
Figure 1a-b     Map of C. californicus range                  Lines 146 - 260
Figure 2b       Shell allometry (shell manual)                Lines 262 - 302
Supp Fig 1      Correlates for shell collection bias          Lines 304 - 331
Figure 3        Bergmann's rule: Morphology ~ temperature     Lines 333 -
Figure 4        Major factors ~ Morphology (SoCal)            Lines

Overview of data and post-processing
```{r}

###### Upload data ######
# install.packages("RCurl")
library("RCurl") # package needed for downloading data from Github repo
# packageVersion("RCurl") # v1.95.4.12 

Raw <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable2_Data-Raw.csv")) 

# Ensure blanks are read as NAs
###### Upload data ######


###### Data structure ######
str(Raw) # Quick view of RawData

# Plots of shell morphology measurements as a point of reference for outliers 
morphos = c("Globosity_WoverL", "Length_mm", "Width_mm", "Aperture_mm", "Thickness_mm") 

# install.packages("PerformanceAnalytics")
library("PerformanceAnalytics") # Need package for correlation charts
# packageVersion("PerformanceAnalytics") # v1.5.2

# install.packages("tidyverse")
library("tidyverse") # Need package pipes ( %>% or "then" )
# packageVersion("tidyverse") # v1.2.1
 
chart.Correlation(Raw %>% select(morphos)) # we can see outliers here
###### Data structure ######


###### Clean-up ######
# Subset for shells without aperture / thickness (iNat will be included)
Raw_globoshells = subset(Raw, is.na(Raw$Thickness_mm))
Raw_iNat = subset(Raw, Raw$Collection=="iNat") # 99 specifically from iNat

# Filter outliers (min and max values identified by viewing Raw)
# Confirmed outliers from operator error (measuring or datasheet input)
Clean_OutliersRemoved = Raw %>% # 2996 obs 
  filter(Thickness_mm < 2) %>% # low thickness
  filter(!(Width_mm < 15 & Thickness_mm > 2)) %>% # too thick for width
  filter(Aperture_mm < 6) %>%  # small aperture
  filter(!(Length_mm > 20 & Aperture_mm <2)) %>% # aperture too small for length
  rbind(Raw_globoshells) %>%   # adds 775 shells back in *length that's above the threshold?
  filter(Width_mm > 2) %>%  # too narrow (note this gets rid of iNat)
  rbind(Raw_iNat) %>% # add iNat back in
  filter(Globosity_WoverL > 0.42 & Globosity_WoverL < 0.71 ) %>% 
  data.frame # ensure it's a data frame
# Remove 25 shells from Pliocene (too old, few)
Clean_OutliersRemoved=subset(Clean_OutliersRemoved, !Epoch=="Pliocene")

# 50 shells of original 2996 removed (2946 obs)
length(Raw %>% filter(Thickness_mm >= 2) %>% .$Specimen_Number)
paste(length(Raw$Specimen_Number)-length(Clean_OutliersRemoved$Specimen_Number), "removed of", length(Raw$Specimen_Number), "shells")

# Ensure updated Clean dataset has 3 or more specimens per lot
Clean = Clean_OutliersRemoved %>% 
  group_by(Lot_Number) %>% 
  mutate(Specimens_per_Lot = length((Lot_Number))) %>%
  ungroup %>% 
  filter(Specimens_per_Lot > 2 | Collection == "iNat") # keep catalogs with > 2 in lot and iNat (1 per 'lot')

# 61 shells of original 2996 removed, 11 from Clean_OutliersRemoved (2935 obs)
paste(length(Raw$Specimen_Number)-length(Clean$Specimen_Number), "removed of", length(Raw$Specimen_Number), "shells")

# Validation check for NA values for final data set
cat("\n\n",
    sum(is.na(Clean$Globosity_WoverL)), "have no globosity ratio,\n", # 0
    sum(is.na(Clean$Width_mm)), "have no width measurement,\n", # 99 [iNat]
    sum(is.na(Clean$Length_mm)), "have no length measurement,\n", # 99 [iNat]
    sum(is.na(Clean$Thickness_mm)), "have no thickness measurement,\n", # 742 = 99 [iNat] + 338 PR + 273 Pleisto
    sum(is.na(Clean$Aperture_mm)), "have no aperture measurement" ) # 742 = 99 [iNat] + 338 PR + 273 Pleisto

# Order the data
Clean %>% group_by(Lot_Number) %>% summarise(N = length(Specimen_Number))
Clean %>% group_by(Epoch) %>% summarise(N = length(Specimen_Number)) 

# Re-do of correlation chart
chart.Correlation(Clean %>% select(morphos)) # much improved

# Parameters for SoCal points
counties = c("Sonoma", "San Mateo", "Monterey", "San Luis Obispo", "Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego", "Baja California")
Raw$County = factor(Raw$County, levels=counties) # order counties

socal = c("Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego")
Clean_SoCal = Clean %>% filter(County %in% socal) # 2597 obs for SoCal 

socal_locations = c("SB", "NLA", "LA", "SLA", "SD", "ANA", "SCZI", "SRI", "SMI", "CAT", "SBI", "SNI", "SCI") # socal location order
Clean_SoCal$Location = factor(Clean_SoCal$Location, levels=socal_locations) # order 

# Distribution of shells by county
Clean_SoCal %>% group_by(County) %>% summarise(N = length(Epoch)) %>% data.frame 
# Check correlation for SoCal subset
chart.Correlation(Clean_SoCal %>% select(morphos)) # Also looks good
###### Clean-up ######


###### Save cleaned-up data for downstream visualization and analysis ######

# Data structure for Clean and SoCal subset
sort(unique(Clean$Collection)) # Unique collections: 9
sort(unique(Clean$Catalog)) # Unique lots: 379 
sort(unique(Clean$County)) # Unique counties: 10
sort(unique(Clean$Location)) # Unique locations: 20
sort(unique(Clean$Land_Category)) # Unique location types: 3
sort(unique(Clean$Habitat)) # Unique habitats: 4
sort(unique(Clean$Depth)) # Unique depths: 3

sort(unique(Clean_SoCal$Collection)) # Unique collections: 9
sort(unique(Clean_SoCal$Catalog)) # Unique lots: 333 
sort(unique(Clean_SoCal$County)) # Unique counties: 5
sort(unique(Clean_SoCal$Location)) # Unique locations: 13

write.csv(Clean, file = "SupplementalTable3_Data-Clean.csv")
write.csv(Clean_SoCal, file = "SupplementalTable4_Data-SoCal.csv") 
# Validation check 
Clean <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable3_Data-Clean.csv")) 
Clean_SoCal <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable4_Data-SoCal.csv")) 

###### Save cleaned-up data for downstream visualization and analysis ######

```

Figure 1a-b: Maps of shell locations (Clean and SoCal Data) 
```{r}

# install.packages("ggplot2")
library("ggplot2")
# packageVersion("ggplot2") # v3.1.0.9000
# devtools::install_github("dkahle/ggmap") 
library("ggmap")
# packageVersion("ggmap") # v3.0.0.900
# install.packages("rgdal")
library("rgdal")
# packageVersion("rgdal") # v1.3.6

# This code is personalized by account and only runs if you have a key
# Instructions:https://github.com/dkahle/ggmap
# Visit your personal Google Maps Platform to retrieve key if you have one
# ggmap::register_google(key = " ") 

# Create parameters and background maps for Clean and Clean_SoCal
# Parameters for All points
# Subset data by Epoch to color code points
Anthro = subset(Clean, Epoch=="Anthropocene") # 1492
Holo = subset(Clean, Epoch=="Holocene") # 1138
Pleisto = subset(Clean, Epoch=="Pleistocene") # 305

All_bbox <- make_bbox(lat = Latitude, lon = Longitude, data = Clean)
All_bbox
#       left     bottom      right        top 
#   -122.9015   25.8435 -114.7285   38.7465 

# Load polygons for ecoregions and call them after downloaded in your folder
# shape files from http://www.marineregions.org/
NC_poly=readOGR( "northern_ecoregions" , layer="ecoregions") 
SCB_poly=readOGR( "southern_ecoregions" , layer="ecoregions") 
MT_poly=readOGR( "magdaelena_ecoregions" , layer="ecoregions") 

All=get_map(location=c(-125,20,-105,40), # adjust to include regions
            zoom = 5, 
            maptype = "terrain") 

SoCal_bbox = make_bbox(lat = Latitude, lon = Longitude, data = Clean_SoCal)
SoCal_bbox
#       left     bottom      right        top 
#      -120.6695   32.4425 -116.9405   34.5875  

SoCal=get_map(location=c(-121,32,-116.5,35), 
            zoom = 7, 
            maptype = "terrain") 

# Figure 1a
SoCal_box = data.frame(x=c(-121,-121,-116.5,-116.5,-121), 
       y=c(35,32,32,35,35)) # create box for zoom of SoCal

Clean_map = ggmap(All) +
  geom_polygon(data = NC_poly, aes(x = long, y = lat),
               fill = "white", alpha = 0.5) + # white
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) + # blue
  geom_polygon(data = MT_poly, aes(x = long, y = lat),
               fill = "#666666", alpha = 0.5) + # grey
  stat_density_2d(data = Clean, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # color fill, orange
  geom_point(data = Anthro, aes(x = Longitude, y = Latitude), 
               color = "#FC4E07",
             size = 1) + # aqua
  geom_point(data = Holo, aes(x = Longitude,y = Latitude), 
             color = "#E7B800", 
             size = 1) + # seagreen
  geom_point(data = Pleisto, aes(x = Longitude, y = Latitude), 
             color = "#00AFBB",
             size = 1) + # blue 
  geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude") 
Clean_map
ggsave("Figure1a_Data-Clean_Map.png", dpi = 800)

# Figure 1b
Anthro_SoCal = subset(Clean_SoCal, Epoch=="Anthropocene") # 1366
Holo_SoCal = subset(Clean_SoCal, Epoch=="Holocene") # 1020
Pleisto_SoCal = subset(Clean_SoCal, Epoch=="Pleistocene") # 211

CleanSoCal_map = ggmap(SoCal) +
  geom_polygon(data = NC_poly, aes(x = long, y = lat),
               fill = "white", alpha = 0.5) + # white
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) + # blue
  stat_density_2d(data = Clean_SoCal, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .35,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # color fill, orange
  geom_point(data = Anthro_SoCal, aes(x = Longitude, y = Latitude), 
               color = "#FC4E07",
             size = 2) + 
  geom_point(data = Holo_SoCal, aes(x = Longitude,y = Latitude), 
             color = "#E7B800", 
             size = 2) + 
  geom_point(data = Pleisto_SoCal, aes(x = Longitude, y = Latitude), 
             color = "#00AFBB",
             size = 2) + 
    geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude")
CleanSoCal_map
ggsave("Figure1b_Data-Clean-SoCal_Map.png", dpi = 800)

```

Figure 2b: Allometry - shell shape vs size (Clean) 
```{r}

# install.packages("GGally")
library(GGally)
# packageVersion("GGally") # v1.4.0
# install.packages("ggplot2")
library(ggplot2)
# packageVersion("ggplot2") # v3.1.0

# Create a function for including linear regression line
my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    geom_smooth(method=lm, color="black", ...) 
  p
}

# Subset data
Fig2 = Clean[,c("Epoch", "Globosity_WoverL", "Length_mm", "Width_mm", "Mean_SST")] 
# Rename columns
names(Fig2) = c("Epoch", "Globosity", "Length", "Width", "Mean SST") 
# Create figure (automatically excludes NAs)
Fig2_chart = ggpairs(Fig2, aes(color = Epoch, alpha = 0.85), lower = list(continuous = my_fn), upper = list(continuous = wrap("cor", size = 4.25)))
# Change colors manually with a loop through each plot
for(i in 1:Fig2_chart$nrow) {
  for(j in 1:Fig2_chart$ncol){
    Fig2_chart[i,j] <- Fig2_chart[i,j] + 
        scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
        scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB"))  
  }
}

Fig2_chart + 
  theme_classic() # theme for figure
ggsave("Figure2b_Data-Clean_Shell-Allometry.png", dpi = 800)

# Length = Width, globosity independent of length and width
# size = length=width, aperture and thickness are neglible as dependent on size (see categories)
# shape = globosity
```

Supplemental Figure 1: Cc shell morphology ~ batch effect (Clean) 
```{r}

SuppFig1 = Clean[,c("Collection", "Lot_Number", "Specimens_per_Lot", "Collector", "Globosity_WoverL", "Length_mm", "Width_mm", "Aperture_mm", "Thickness_mm", "Epoch")] 
# Rename columns
names(SuppFig1) = c("Coll", "Lot", "Specimens", "Collector", "Globosity", "Length", "Width", "Aperture", "Thickness", "Epoch") 

# Assign variables as numeric such that they are viewed as continuous
SuppFig1 = transform(SuppFig1, 
                 Coll = as.numeric(Coll), 
                 Lot = as.numeric(Lot),
                 Specimens = as.numeric(Specimens),
                 Collector = as.numeric(Collector),
                 Globosity = as.numeric(Globosity),
                 Length = as.numeric(Length),
                 Width = as.numeric(Width),
                 Aperture = as.numeric(Aperture),
                 Thickness = as.numeric(Thickness), 
                 Epoch = as.numeric(Epoch))

SuppFig1 = ggcorr(SuppFig1, geom = "blank", label = TRUE, hjust = 0.75, size = 3) +
  geom_point(size = 10, aes(color = coefficient > 0, alpha = abs(coefficient) > 0.5)) +
  scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
  guides(color = FALSE, alpha = FALSE)
SuppFig1
ggsave("SupplementalFigure1_Clean-Data_Batch-Effect.png", dpi = 800)

```

Figure 3: Relationship between shell morphology and temperature (Clean)
```{r}

Clean_Length = Clean[!is.na(Clean$Length_mm), ] # remove rows with NAs [iNat]
Clean_Shape = Clean[!is.na(Clean$Globosity_WoverL), ]

################################################
# STATISTICS
### Distribution of data
hist(Clean_Length$Length_mm) # fairly normal
hist(Clean_Length$Mean_SST) # slightly skewed

### Parametric (assumes normality)
# linear model (1st order polynomial)? No
x = lm(Length_mm~Mean_SST, Clean_Length)
shapiro.test(residuals(x)) # W = 0.97431, p-value < 2.2e-16
# p needs to be over 0.05
plot(x) # shows residuals, qqplot not great

# nonlinear models
# logarithmic model? No (about the same as before)
ya = lm(Length_mm~log(Mean_SST), Clean_Length) # log transform 
shapiro.test(residuals(ya)) # W = 0.9742, p-value < 2.2e-16
plot(ya) 
# log-log model? No (looks decent)
yb = lm(log(Length_mm)~log(Mean_SST), Clean_Length)  
shapiro.test(residuals(yb)) # W = 0.98499, p-value < 2.2e-16
plot(yb) 
# 2nd order polynomial model? No
z = lm(Length_mm~poly(Mean_SST,2), Clean_Length)
shapiro.test(residuals(z)) # W = 0.97539, p-value < 2.2e-16
plot(z)

summary(x) # r^2 = -0.0001501 
summary(ya) # r^2 = -0.0002482 
summary(yb) # r^2 =  0.0001369
summary(z) # r^2 = 0.002156 (best, but not by a lot...)
# 2nd order (z) didn't pass Shapiro's test (log-log), qqplots look bad

### Non-parametric (assumes non-normal, but weaker statistic)
cor.test(Clean_Length$Length_mm, Clean_Length$Mean_SST, method = "spearman") 
# S = 3885200000, p-value = 0.2418, rho = -0.02198427 ('r^2')
# p value needs to be less than 0.05 to be significant
# data is linear

cor.test(Clean_Shape$Globosity_WoverL, Clean_Shape$Mean_SST, method = "spearman") 
# S = 4451700000, p-value = 0.002214, rho = -0.05646121 
# data is not linear, but relationship to temperature is significant

################################################

locations = c("SM", "MB", "MO", "BS", "SO", "SB", "NLA", "LA", "SLA", "SD", "NBC", "SBC",  "ANA","SCZI","SRI","SMI","SCI","SNI","SBI","CAT")
# Coastal "SM", "MB", "MO", "BS", "SO", "SB", "NLA", "LA", "SLA", "SD", "NBC", "SBC"
# Northern islands (marine sanctuary) "ANA","SCZI","SRI","SMI"
# Southern islands "SCI","SNI","SBI","CAT"
Clean$Location = factor(Clean$Location, levels=locations) # order locations

Figure3a = ggplot(data = Clean, 
                      aes(x = Latitude, y = Mean_SST,
                      color = Epoch)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.3, fill="red") + # shade square
  geom_point() +
    geom_smooth() + # default of 95% confidence
    geom_smooth(method = lm, color = "black") + # 95% confidence
  theme(axis.line = element_line(colour = "black", size = 0.25)) +
  labs(x="Latitude", y="Mean SST")
Figure3a +
  theme_classic()
ggsave("Figure3a_Data-Clean_Latitude-MeanSST.png", dpi = 800)

Figure3b = ggplot(data = Clean, 
                      aes(x = Mean_SST, y = Globosity_WoverL,
                      color = Epoch)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_point() +
    geom_smooth() + # default of 95% confidence
    geom_smooth(method = lm, color = "black") + # 95% confidence
  theme(axis.line = element_line(colour = "black", size = 0.25)) +
  labs(x="Mean SST", y="Shape")
Figure3b +
  theme_classic()
ggsave("Figure3b_Data-Clean_SST-Shape.png", dpi = 800)

Figure3c = ggplot(data = Clean, 
                      aes(x = Mean_SST, y = Length_mm,
                      color = Epoch)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_point() +
    geom_smooth() + # default of 95% confidence
    geom_smooth(method = lm, color = "black") + # 95% confidence
  theme(axis.line = element_line(colour = "black", size = 0.25)) +
  labs(x="Mean SST", y="Size")
Figure3c +
  theme_classic()
ggsave("Figure3c_Data-Clean_SST-Size.png", dpi = 800)

```

Figure 4: Relationship between shell morphology and other factors (SoCal)
```{r}

# install.packages("corrr")
library(corrr)
# packageVersion("corrr") # v0.3.1

# Subset SoCal by Epoch
Fig4a = Pleisto_SoCal[,c("Ecoregion", "Location", "Latitude", "Mean_SST", "Globosity_WoverL", "Length_mm")] # No Land Category, Decade, or Population
Fig4b = Holo_SoCal[,c("Ecoregion", "Land_Category", "Location", "Latitude", "Decade", "Mean_SST", "Globosity_WoverL", "Length_mm", "Population_Density_NoBaja")] 
Fig4c = Anthro_SoCal[,c("Ecoregion", "Land_Category", "Location", "Latitude", "Decade", "Mean_SST", "Globosity_WoverL", "Length_mm", "Population_Density_NoBaja")] 

# Rename columns
names(Fig4a) = c("Ecoregion", "Location", "Latitude", "SST", "Shape", "Size") 
names(Fig4b) = c("Ecoregion", "Landscape", "Location", "Latitude", "Decade", "SST", "Shape", "Size", "Humans") 
names(Fig4c) = c("Ecoregion", "Landscape", "Location", "Latitude", "Decade", "SST", "Shape", "Size", "Humans") 

# Assign variables as numeric such that they are viewed as continuous
Fig4a_Pleisto = transform(Fig4a, 
                 Ecoregion = as.numeric(Ecoregion), 
                 Location = as.numeric(Location),
                 Latitude = as.numeric(Latitude),
                 SST = as.numeric(SST),
                 Shape = as.numeric(Shape),
                 Size = as.numeric(Size))
Fig4b_Holo = transform(Fig4b, 
                 Ecoregion = as.numeric(Ecoregion), 
                 Landscape = as.numeric(Landscape),
                 Location = as.numeric(Location),
                 Latitude = as.numeric(Latitude),
                 Decade = as.numeric(Decade),
                 SST = as.numeric(SST),
                 Shape = as.numeric(Shape),
                 Size = as.numeric(Size),
                 Humans = as.numeric(Humans))
Fig4c_Anthro = transform(Fig4c, 
                 Ecoregion = as.numeric(Ecoregion), 
                 Landscape = as.numeric(Landscape),
                 Location = as.numeric(Location),
                 Latitude = as.numeric(Latitude),
                 Decade = as.numeric(Decade),
                 SST = as.numeric(SST),
                 Shape = as.numeric(Shape),
                 Size = as.numeric(Size),
                 Humans = as.numeric(Humans))

# Produce networks and supporting correlations
Fig4a_Pleisto %>% correlate() %>% network_plot(min_cor=0.5)
ggsave("Figure4ai_Clean_SoCal-Data_Pleistocene-Network.png", dpi = 800)
Fig4aii = ggcorr(Fig4a_Pleisto, geom = "blank", label = TRUE, hjust = 0.75, size = 3) +
  geom_point(size = 10, aes(color = coefficient > 0, alpha = abs(coefficient) > 0.5)) +
  scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
  guides(color = FALSE, alpha = FALSE)
Fig4aii
ggsave("Figure4aii_Clean_SoCal-Data_Pleistocene-Chart.png", dpi = 800)

Fig4b_Holo %>% correlate() %>% network_plot(min_cor=0.5)
ggsave("Figure4bi_Clean_SoCal-Data_Holocene-Network.png", dpi = 800)
Fig4bii = ggcorr(Fig4b_Holo, geom = "blank", label = TRUE, hjust = 0.75, size = 3) +
  geom_point(size = 10, aes(color = coefficient > 0, alpha = abs(coefficient) > 0.5)) +
  scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
  guides(color = FALSE, alpha = FALSE)
Fig4bii
ggsave("Figure4bii_Clean_SoCal-Data_Holocene-Chart.png", dpi = 800)

Fig4c_Anthro %>% correlate() %>% network_plot(min_cor=0.5)
ggsave("Figure4ci_Clean_SoCal-Data_Anthropocene-Network.png", dpi = 800)
Fig4cii = ggcorr(Fig4c_Anthro, geom = "blank", label = TRUE, hjust = 0.75, size = 3) +
  geom_point(size = 10, aes(color = coefficient > 0, alpha = abs(coefficient) > 0.5)) +
  scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
  guides(color = FALSE, alpha = FALSE)
Fig4cii
ggsave("Figure4cii_Clean_SoCal-Data_Anthropocene-Chart.png", dpi = 800)

```
