---
title: "Bergmann’s Rule across the Pleistocene, Holocene, and Anthropocene in an Eastern Pacific neogastropod, Californiconus californicus (Reeve, 1844) (Gastropoda: Conidae)"
author: "Malloy and Ul-Hasan et al"
date: "Mar 14, 2026"
output:
  pdf_document: default
---

Last edited [Sabah Ul-Hasan]: Mar 14, 2026
https://github.com/MichaelMalloy/C_californicus_Morphology

Contributors of data by order of contribution: MM, SUH, SC, CAO, JKK, WFG
Contributors of code by order of contribution: SUH, LL, MM, SS, CAO

iNat: 
# Include dead or alive?
# Address NA columns (location columns, SST, zooplankton etc)
# Habitat	Depth	BRY	BAR	LMP	TWS	COR	ALG	SCR	BOR_TAP	BOR_STR	BOR_INC



```{r}

##### Data Handling ##### 


# Clean-up of Data-iNat-Raw.csv, ensure path match
iNat <- read.csv("/Users/sulhasan/Downloads/C_californicus_Morphology/data/raw/Data_iNat-Raw.csv")
## Function to check for non-NA, non-empty values
has_values <- function(x) {
  !is.na(x) && x != ""
}
## Initialize empty vectors to store row indices
keep_rows <- c()
remove_rows <- c()
## Iterate through each row of the data frame
for (i in 1:nrow(iNat)) {
  # Check if both length and width have values
  if (has_values(iNat$length[i]) && has_values(iNat$width[i])) {
    keep_rows <- c(keep_rows, i)
  } else {
    remove_rows <- c(remove_rows, i)
  }
}
## Create the subsetted data frame
iNat_keep <- iNat[keep_rows, ]
## Display results
cat(paste("Rows kept:", length(keep_rows), "\n"))
cat(paste("Rows removed:", length(remove_rows), "\n"))
cat(paste("Percentage kept:", round(length(keep_rows)/nrow(iNat)*100, 1), "%\n"))
## Create Globosity_WoverL column and remove length/width columns
iNat_keep$Globosity_WoverL <- iNat_keep$width / iNat_keep$length
iNat_keep <- iNat_keep[, !names(iNat_keep) %in% c("length", "width")]
## Drop specified columns
columns_to_drop <- c("species_name", "observation_license", 
                     "observation_quality", "X")
iNat_keep <- iNat_keep[, !names(iNat_keep) %in% columns_to_drop]
## Rename columns
rename_cols <- c("observation_id" = "Lot_Number", 
                 "observer_login" = "Collector",
                 "observation_latidude" = "Latitude", 
                 "observation_longitude" = "Longitude")
# Only rename columns that exist
for (old_name in names(rename_cols)) {
  if (old_name %in% names(iNat_keep)) {
    names(iNat_keep)[names(iNat_keep) == old_name] <- rename_cols[old_name]
  }
}


# Validation check of Globosity_WoverL between datasets
Raw <- read.csv("/Users/sulhasan/Downloads/C_californicus_Morphology/data/raw/Data_Raw.csv")
## Find common Lot_Numbers between the two datasets
common_lot_numbers <- intersect(iNat_keep$Lot_Number, Raw$Lot_Number)

if (length(common_lot_numbers) > 0) {
  # Create dataframes with matching rows
  iNat_matches <- iNat_keep[iNat_keep$Lot_Number %in% common_lot_numbers, ]
  Raw_matches <- Raw[Raw$Lot_Number %in% common_lot_numbers, ]
  # Merge datasets for comparison
  comparison_df <- merge(
    iNat_matches[, c("Lot_Number", "Globosity_WoverL")],
    Raw_matches[, c("Lot_Number", "Globosity_WoverL")],
    by = "Lot_Number",
    suffixes = c("_iNat", "_Raw")
  )
  # Calculate differences
  comparison_df$Difference <- comparison_df$Globosity_WoverL_iNat - comparison_df$Globosity_WoverL_Raw
  comparison_df$Abs_Difference <- abs(comparison_df$Difference)
  # Summary statistics
  sd_diff <- sd(comparison_df$Difference)
  mean_abs <- mean(comparison_df$Abs_Difference)
  sd_abs <- sd(comparison_df$Abs_Difference)
  cat("\n=== Globosity_WoverL Comparison ===\n")
  cat(paste("Total matching Lot_Numbers:", nrow(comparison_df), "\n"))
  cat(paste("Standard deviation of differences:", round(sd_diff, 6), "\n"))
  # Apply 1 SD threshold
  threshold_1sd <- sd_diff
  rows_to_keep <- comparison_df$Lot_Number[comparison_df$Abs_Difference <= threshold_1sd]
  rows_to_remove <- comparison_df$Lot_Number[comparison_df$Abs_Difference > threshold_1sd]
  cat("\n=== 1 Standard Deviation Filtering ===\n")
  cat(paste("1 SD threshold:", round(threshold_1sd, 6), "\n"))
  cat(paste("Rows within 1 SD:", length(rows_to_keep), 
            "(", round(length(rows_to_keep)/nrow(comparison_df)*100, 1), "%)\n"))
  cat(paste("Rows outside 1 SD:", length(rows_to_remove), 
            "(", round(length(rows_to_remove)/nrow(comparison_df)*100, 1), "%)\n"))
  # Store original row counts before filtering
  iNat_before <- nrow(iNat_keep)
  Raw_before <- nrow(Raw)
  # Filter both datasets to keep only rows within 1 SD
  iNat_keep <- iNat_keep[!iNat_keep$Lot_Number %in% rows_to_remove, ]
  Raw <- Raw[!Raw$Lot_Number %in% rows_to_remove, ]
  cat("\n=== Datasets Updated ===\n")
  cat(paste("iNat_keep: from", iNat_before, "to", nrow(iNat_keep), 
            "rows (removed", iNat_before - nrow(iNat_keep), ")\n"))
  cat(paste("Raw: from", Raw_before, "to", nrow(Raw), 
            "rows (removed", Raw_before - nrow(Raw), ")\n"))
  ## Create comprehensive comparison dataset (moved inside the if statement)
  # Merge full datasets for all matching Lot_Numbers
  full_comparison <- merge(
    iNat_matches[, c("Lot_Number", "observation_date", "Latitude", "Longitude", 
                     "observation_type_0dead", "Globosity_WoverL")],
    Raw_matches[, c("Lot_Number", "Collection", "Ecoregion", "County", "Year", "Month", 
                    "Season", "Globosity_WoverL")],
    by = "Lot_Number",
    suffixes = c("_iNat", "_Raw"),
    all = TRUE
  )
  # Calculate difference and flag if within 1 SD
  full_comparison$Difference <- full_comparison$Globosity_WoverL_iNat - full_comparison$Globosity_WoverL_Raw
  full_comparison$Abs_Difference <- abs(full_comparison$Difference)
  full_comparison$Within_1SD <- full_comparison$Abs_Difference <= sd_diff
  # Add status column
  full_comparison$Status <- ifelse(full_comparison$Within_1SD, "Match", "Mismatch")
  # Sort by absolute difference (largest mismatches first)
  full_comparison <- full_comparison[order(-full_comparison$Abs_Difference), ]
  # Save comprehensive comparison
  write.csv(full_comparison, 
            "/Users/sulhasan/Downloads/C_californicus_Morphology/data/Data_iNat-OverlapValidation.csv", 
            row.names = FALSE)
  cat("\nComprehensive comparison saved as: Data_iNat-OverlapValidation.csv\n")
} else {
  cat("\nNo overlapping Lot_Numbers found between the two datasets.\n")
}


# Merge of Data_iNat-Clean.csv into Data_Raw-Filtered.csv
## Write and save usable data, ensure path
write.csv(iNat_keep, "/Users/sulhasan/Downloads/C_californicus_Morphology/data/Data_iNat-Clean.csv", row.names = FALSE)
write.csv(Raw, "/Users/sulhasan/Downloads/C_californicus_Morphology/data/Data_Raw-Filtered.csv", row.names = FALSE)

## Identify Lot_Numbers already in Raw_filtered
existing_lot_numbers <- Raw$Lot_Number
## Split iNat_clean into existing and new records
iNat_existing <- iNat_keep[iNat_keep$Lot_Number %in% existing_lot_numbers, ]
iNat_new <- iNat_keep[!iNat_keep$Lot_Number %in% existing_lot_numbers, ]
cat("\n=== Data Integration ===\n")
cat(paste("Records already in Raw_filtered:", nrow(iNat_existing), "\n"))
cat(paste("New records to add:", nrow(iNat_new), "\n"))
## For new records, create new rows for Raw
if (nrow(iNat_new) > 0) {
  # Step 1: Ensure Latitude and Longitude are numeric
  iNat_new$Latitude <- as.numeric(as.character(iNat_new$Latitude))
  iNat_new$Longitude <- as.numeric(as.character(iNat_new$Longitude))
  # Step 2: Determine starting Specimen_Number
  # Find the maximum existing Specimen_Number
  max_specimen <- max(as.numeric(Raw$Specimen_Number[!is.na(Raw$Specimen_Number)]), na.rm = TRUE)
  if (is.infinite(max_specimen)) max_specimen <- 2996  # Start from 2997 if no existing numbers
  # Step 3: Create a temporary copy of Raw with rounded coordinates for lookup
  Raw_temp <- Raw
  Raw_temp$Latitude <- as.numeric(as.character(Raw_temp$Latitude))
  Raw_temp$Longitude <- as.numeric(as.character(Raw_temp$Longitude))
  # Round coordinates in Raw_temp to 2 decimal places for matching
  Raw_temp$Latitude_rounded <- round(Raw_temp$Latitude, 2)
  Raw_temp$Longitude_rounded <- round(Raw_temp$Longitude, 2)
  # Create lookup tables for location-based matching using rounded coordinates
  location_cols <- c("Latitude_rounded", "Longitude_rounded", "Ecoregion", "County", 
                     "Coastal_Proximity", "Location", "Site", "SoCal_Group", "Island_Size")
  # Get unique location combinations from Raw_temp
  location_lookup <- unique(Raw_temp[, location_cols])
  location_lookup <- location_lookup[complete.cases(location_lookup[, c("Latitude_rounded", "Longitude_rounded")]), ]
  # For SST data - match based on Year, Season, AND Month
  sst_cols <- c("Year", "Season", "Month", "Mean_SST", "Mean_SST_Sdev")
  sst_lookup <- unique(Raw[, sst_cols])
  sst_lookup <- sst_lookup[complete.cases(sst_lookup[, c("Year", "Season", "Month", "Mean_SST")]), ]
  # Step 4: Create lookup for environmental and fisheries columns
  # These columns will be matched by rounded Latitude/Longitude AND Year
  env_cols <- c("Latitude_rounded", "Longitude_rounded", "Year",
                "mean_pvolc3", "sd_pvolc3", 
                "zooplankton_n_points_used", "zooplankton_mean_distance",
                "fish_total_catch_wt_kg", "fish_cpue_kg_per_ha_der",
                "sd_total_catch_wt_kg", "sd_cpue_kg_per_ha_der",
                "fish_n_points_used", "fish_mean_distance",
                "Population_Density", "Population_Density_NoBaja", "MPA")
  # Create lookup table with unique combinations
  env_lookup <- unique(Raw_temp[, env_cols])
  env_lookup <- env_lookup[complete.cases(env_lookup[, c("Latitude_rounded", "Longitude_rounded", "Year")]), ]
  # Create a template dataframe with all Raw columns
  new_rows <- data.frame(matrix(NA, nrow = nrow(iNat_new), ncol = ncol(Raw)))
  colnames(new_rows) <- colnames(Raw)
  # Fill in basic information from iNat_new - but with coordinates rounded to 2 decimals
  new_rows$Specimen_Number <- max_specimen + 1:nrow(iNat_new)
  new_rows$Collection <- "iNat"
  new_rows$Lot_Number <- as.character(iNat_new$Lot_Number)
  new_rows$Collector <- iNat_new$Collector  # Collector already exists in iNat_keep
  new_rows$Latitude <- round(iNat_new$Latitude, 2)  # Round to 2 decimal places
  new_rows$Longitude <- round(iNat_new$Longitude, 2)  # Round to 2 decimal places
  new_rows$Globosity_WoverL <- iNat_new$Globosity_WoverL
  # Extract date components from observation_date
  date_parts <- strsplit(as.character(iNat_new$observation_date), "/")
  new_rows$Month <- as.numeric(sapply(date_parts, function(x) x[1]))
  new_rows$Year <- as.numeric(sapply(date_parts, function(x) {
    year_val <- x[3]
    if (nchar(year_val) == 2) {
      if (as.numeric(year_val) < 50) {
        return(2000 + as.numeric(year_val))
      } else {
        return(1900 + as.numeric(year_val))
      }
    } else {
      return(as.numeric(year_val))
    }
  }))
  # Determine season based on month
  new_rows$Season <- ifelse(new_rows$Month %in% c(12, 1, 2), "Winter",
                          ifelse(new_rows$Month %in% c(3, 4, 5), "Spring",
                          ifelse(new_rows$Month %in% c(6, 7, 8), "Summer", "Fall")))
  # Set Decade based on Year
  new_rows$Decade <- floor(new_rows$Year / 10) * 10
  # Set Epoch (assuming all recent are Anthropocene)
  new_rows$Epoch <- "Anthropocene"
  # Set Formation as NA
  new_rows$Formation <- NA
  # Step 5: Create rounded versions for matching (these will be discarded later)
  new_rows$Latitude_rounded <- new_rows$Latitude  # Already rounded
  new_rows$Longitude_rounded <- new_rows$Longitude  # Already rounded
  # Step 6: Match location-based columns by rounded Latitude/Longitude
  cat("\n=== Matching Location Data ===\n")
  location_match_count <- 0
  for (i in 1:nrow(new_rows)) {
    # Skip if coordinates are NA
    if (is.na(new_rows$Latitude_rounded[i]) || is.na(new_rows$Longitude_rounded[i])) {
      next
    }
    # Find matching rounded coordinates in location_lookup
    coord_match <- which(abs(location_lookup$Latitude_rounded - new_rows$Latitude_rounded[i]) < 0.01 & 
                         abs(location_lookup$Longitude_rounded - new_rows$Longitude_rounded[i]) < 0.01)
    if (length(coord_match) > 0) {
      # Use the first match if multiple
      match_idx <- coord_match[1]
      new_rows$Ecoregion[i] <- location_lookup$Ecoregion[match_idx]
      new_rows$County[i] <- location_lookup$County[match_idx]
      new_rows$Coastal_Proximity[i] <- location_lookup$Coastal_Proximity[match_idx]
      new_rows$Location[i] <- location_lookup$Location[match_idx]
      new_rows$Site[i] <- location_lookup$Site[match_idx]
      new_rows$SoCal_Group[i] <- location_lookup$SoCal_Group[match_idx]
      new_rows$Island_Size[i] <- location_lookup$Island_Size[match_idx]
      location_match_count <- location_match_count + 1
    }
  }
  cat(paste("Location data matched for", location_match_count, "out of", nrow(new_rows), "records\n"))
  # Step 7: Match SST data by Year AND Season AND Month
  cat("\n=== Matching SST Data ===\n")
  sst_match_count <- 0
  for (i in 1:nrow(new_rows)) {
    # Skip if temporal data is missing
    if (is.na(new_rows$Year[i]) || is.na(new_rows$Season[i]) || is.na(new_rows$Month[i])) {
      next
    }
    # Find matching year, season, and month
    year_match <- new_rows$Year[i]
    season_match <- new_rows$Season[i]
    month_match <- new_rows$Month[i]
    # Find exact match for year, season, and month
    match_idx <- which(sst_lookup$Year == year_match & 
                       sst_lookup$Season == season_match & 
                       sst_lookup$Month == month_match)
    
    if (length(match_idx) > 0) {
      # Use the first match if multiple
      idx <- match_idx[1]
      new_rows$Mean_SST[i] <- sst_lookup$Mean_SST[idx]
      new_rows$Mean_SST_Sdev[i] <- sst_lookup$Mean_SST_Sdev[idx]
      sst_match_count <- sst_match_count + 1
    } else {
      # If no exact match, try matching by year and season only (as fallback)
      fallback_idx <- which(sst_lookup$Year == year_match & 
                           sst_lookup$Season == season_match)
      if (length(fallback_idx) > 0) {
        idx <- fallback_idx[1]
        new_rows$Mean_SST[i] <- sst_lookup$Mean_SST[idx]
        new_rows$Mean_SST_Sdev[i] <- sst_lookup$Mean_SST_Sdev[idx]
        sst_match_count <- sst_match_count + 1
      }
    }
  }
  cat(paste("SST data matched for", sst_match_count, "out of", nrow(new_rows), "records\n"))
  # Step 8: Match environmental and fisheries data by rounded coordinates AND Year
  cat("\n=== Matching Environmental & Fisheries Data ===\n")
  env_match_count <- 0
  for (i in 1:nrow(new_rows)) {
    # Skip if coordinates or year are missing
    if (is.na(new_rows$Latitude_rounded[i]) || is.na(new_rows$Longitude_rounded[i]) || is.na(new_rows$Year[i])) {
      next
    }
    # Find matches by rounded coordinates AND year
    env_match <- which(abs(env_lookup$Latitude_rounded - new_rows$Latitude_rounded[i]) < 0.01 & 
                       abs(env_lookup$Longitude_rounded - new_rows$Longitude_rounded[i]) < 0.01 &
                       env_lookup$Year == new_rows$Year[i])
    
    if (length(env_match) > 0) {
      # Use the first match if multiple
      match_idx <- env_match[1]
      # Fill in all environmental and fisheries columns
      new_rows$mean_pvolc3[i] <- env_lookup$mean_pvolc3[match_idx]
      new_rows$sd_pvolc3[i] <- env_lookup$sd_pvolc3[match_idx]
      new_rows$zooplankton_n_points_used[i] <- env_lookup$zooplankton_n_points_used[match_idx]
      new_rows$zooplankton_mean_distance[i] <- env_lookup$zooplankton_mean_distance[match_idx]
      new_rows$fish_total_catch_wt_kg[i] <- env_lookup$fish_total_catch_wt_kg[match_idx]
      new_rows$fish_cpue_kg_per_ha_der[i] <- env_lookup$fish_cpue_kg_per_ha_der[match_idx]
      new_rows$sd_total_catch_wt_kg[i] <- env_lookup$sd_total_catch_wt_kg[match_idx]
      new_rows$sd_cpue_kg_per_ha_der[i] <- env_lookup$sd_cpue_kg_per_ha_der[match_idx]
      new_rows$fish_n_points_used[i] <- env_lookup$fish_n_points_used[match_idx]
      new_rows$fish_mean_distance[i] <- env_lookup$fish_mean_distance[match_idx]
      new_rows$Population_Density[i] <- env_lookup$Population_Density[match_idx]
      new_rows$Population_Density_NoBaja[i] <- env_lookup$Population_Density_NoBaja[match_idx]
      new_rows$MPA[i] <- env_lookup$MPA[match_idx]
      
      env_match_count <- env_match_count + 1
    } else {
      # Fallback: try matching by coordinates only (any year)
      coord_only_match <- which(abs(env_lookup$Latitude_rounded - new_rows$Latitude_rounded[i]) < 0.01 & 
                                abs(env_lookup$Longitude_rounded - new_rows$Longitude_rounded[i]) < 0.01)
      
      if (length(coord_only_match) > 0) {
        match_idx <- coord_only_match[1]
        
        new_rows$mean_pvolc3[i] <- env_lookup$mean_pvolc3[match_idx]
        new_rows$sd_pvolc3[i] <- env_lookup$sd_pvolc3[match_idx]
        new_rows$zooplankton_n_points_used[i] <- env_lookup$zooplankton_n_points_used[match_idx]
        new_rows$zooplankton_mean_distance[i] <- env_lookup$zooplankton_mean_distance[match_idx]
        new_rows$fish_total_catch_wt_kg[i] <- env_lookup$fish_total_catch_wt_kg[match_idx]
        new_rows$fish_cpue_kg_per_ha_der[i] <- env_lookup$fish_cpue_kg_per_ha_der[match_idx]
        new_rows$sd_total_catch_wt_kg[i] <- env_lookup$sd_total_catch_wt_kg[match_idx]
        new_rows$sd_cpue_kg_per_ha_der[i] <- env_lookup$sd_cpue_kg_per_ha_der[match_idx]
        new_rows$fish_n_points_used[i] <- env_lookup$fish_n_points_used[match_idx]
        new_rows$fish_mean_distance[i] <- env_lookup$fish_mean_distance[match_idx]
        new_rows$Population_Density[i] <- env_lookup$Population_Density[match_idx]
        new_rows$Population_Density_NoBaja[i] <- env_lookup$Population_Density_NoBaja[match_idx]
        new_rows$MPA[i] <- env_lookup$MPA[match_idx]
        
        env_match_count <- env_match_count + 1
      }
    }
  }
  
  cat(paste("Environmental/fisheries data matched for", env_match_count, "out of", nrow(new_rows), "records\n"))
  # Remove temporary rounded coordinate columns before appending
  new_rows$Latitude_rounded <- NULL
  new_rows$Longitude_rounded <- NULL
  # Append new rows to Raw
  Raw <- rbind(Raw, new_rows)
  cat("\n=== New Records Added ===\n")
  cat(paste("Assigned Specimen_Numbers:", min(new_rows$Specimen_Number), "to", max(new_rows$Specimen_Number), "\n"))
}
## Round all coordinates in the entire Raw dataset to 2 decimal places for consistency
Raw$Latitude <- round(as.numeric(as.character(Raw$Latitude)), 2)
Raw$Longitude <- round(as.numeric(as.character(Raw$Longitude)), 2)
## Sort by Specimen_Number (smallest to largest)
Raw$Specimen_Number <- as.numeric(as.character(Raw$Specimen_Number))
Raw <- Raw[order(Raw$Specimen_Number), ]
## Save the updated dataset
write.csv(Raw, 
          "/Users/sulhasan/Downloads/C_californicus_Morphology/data/Data_Raw-Integrated.csv", 
          row.names = FALSE)
cat(paste("Final integrated dataset has", nrow(Raw), "rows\n"))
cat("Saved as: Data_Raw-Integrated.csv\n")


##### Data Handling ##### 
```

later

Data Clean-up       Post-processing: Raw -> Clean -> SoCal      Lines 25 - 154
Fig 1a-b            Map of C. californicus range                Lines 156 - 272
Fig 2b              Shell allometry (shell manual)              Lines 274 - 329
Fig 3a-d,Supp Fig 1 Size or Shape and Lat / Temp, Bergmann's    Lines 331 - 483
Fig 4a-f,Supp Fig 2 Size or Shape and Ecoregion + Proxmity      Lines 475 - 630
Tables 1-2          Statistics breakdown (of all figures)       Lines 632 - 702
Supp Fig 3          Correlations to human population density    Lines 704 - 769
Cope's Rule         Size by location ~ Epoch                    Lines 771 - 

- Make sure code below reflects figures and tables (work backwards from figures/tables in folder to code to formalize)
- Update map to include iNat raw for natural history map (make sure no duplicates)
- Compare in other models, and ask this one what prompt it would give to start from scratch
- Need to bin islands by proximity to coast and each other
- Why is Socal and Clean the same as Raw for rows?


```{r}
# ==============================================================================
# COMPLETE ANALYSIS PIPELINE: Californiconus californicus Morphology
# Testing Bergmann's Rule across Pleistocene, Holocene, and Anthropocene
# ==============================================================================

# ==============================================================================
# SECTION 1: PACKAGE INSTALLATION AND LOADING
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 1: INSTALLING AND LOADING PACKAGES\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Required packages
required_packages <- c(
  # Data manipulation
  "tidyverse", "lubridate", "broom",
  
  # Statistical tests
  "nortest", "moments", "car", "lmtest", "boot", "pwr", "effectsize",
  
  # Regression models
  "MASS", "lme4", "nlme", "mgcv", "emmeans",
  
  # Spatial and maps
  "sf", "ggspatial", "maps", "mapdata",
  
  # Visualization
  "ggplot2", "cowplot", "ggsignif", "ggpubr", "gridExtra", "GGally",
  
  # Web data
  "RCurl"
)

# Function to safely install missing packages
install_if_missing <- function(packages) {
  installed <- rownames(installed.packages())
  to_install <- packages[!packages %in% installed]
  
  if(length(to_install) > 0) {
    cat("Installing missing packages:", paste(to_install, collapse = ", "), "\n")
    install.packages(to_install, dependencies = TRUE, repos = "https://cloud.r-project.org")
  } else {
    cat("All required packages are already installed.\n")
  }
}

# Install missing packages
install_if_missing(required_packages)

# Load all packages
cat("\nLoading packages...\n")
for(pkg in required_packages) {
  library(pkg, character.only = TRUE)
}
cat("✓ All packages loaded successfully\n")

# ==============================================================================
# SECTION 2: CREATE OUTPUT DIRECTORIES
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 2: CREATING OUTPUT DIRECTORIES\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Create main directories
dirs <- c("data", "data/null", "results", "results/figures", "results/summary")
for(d in dirs) {
  if(!dir.exists(d)) {
    dir.create(d, recursive = TRUE)
    cat("✓ Created directory:", d, "\n")
  } else {
    cat("✓ Directory exists:", d, "\n")
  }
}

# ==============================================================================
# SECTION 3: DATA LOADING AND CLEANING
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 3: DATA LOADING AND CLEANING\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Function to load data from GitHub
load_data_from_github <- function() {
  
  cat("Loading data from GitHub repository...\n")
  
  # Define GitHub base URL
  base_url <- "https://raw.githubusercontent.com/sabahzero/C_californicus_Morphology/main/"
  
  # Files to load
  files <- list(
    raw = "data/Data-Raw.csv",
    key = "data/Data-Key.csv"
  )
  
  # Function to load a single file
  try_load <- function(filename, description) {
    url <- paste0(base_url, filename)
    tryCatch({
      data <- read.csv(text = getURL(url), stringsAsFactors = FALSE)
      cat("  ✓ Loaded", description, "(", nrow(data), "rows )\n")
      return(data)
    }, error = function(e) {
      cat("  ✗ Failed to load", description, ":", e$message, "\n")
      return(NULL)
    })
  }
  
  # Load files
  data_list <- list()
  data_list$raw <- try_load(files$raw, "raw data")
  data_list$key <- try_load(files$key, "data key")
  
  if(is.null(data_list$raw)) {
    stop("Cannot proceed without raw data file")
  }
  
  return(data_list)
}

# Function to clean data with detailed diagnostics
clean_raw_data <- function(raw_data, key_data) {
  
  cat("\nCleaning raw data...\n")
  
  # Make a copy
  data <- raw_data
  
  # Initial row count
  initial_rows <- nrow(data)
  cat("  Initial rows:", initial_rows, "\n")
  
  # Store raw data for comparison
  raw_data_saved <- data
  
  # Step 1: Check for Pliocene specimens
  if("Epoch" %in% names(data)) {
    cat("  Epoch distribution in raw data:\n")
    print(table(data$Epoch, useNA = "ifany"))
    
    data <- data[data$Epoch != "Pliocene", ]
    cat("  After removing Pliocene:", nrow(data), "rows (removed", initial_rows - nrow(data), ")\n")
  }
  
  # Step 2: Filter latitude range
  if("Latitude" %in% names(data)) {
    before_lat <- nrow(data)
    data <- data[data$Latitude > 26.43 & data$Latitude < 38.16 & !is.na(data$Latitude), ]
    cat("  After latitude filtering:", nrow(data), "rows (removed", before_lat - nrow(data), ")\n")
  }
  
  # Step 3: Filter size outliers
  if("Length_mm" %in% names(data)) {
    before_len <- nrow(data)
    data <- data[data$Length_mm >= 5 & data$Length_mm <= 50 & !is.na(data$Length_mm), ]
    cat("  After length filtering (5-50mm):", nrow(data), "rows (removed", before_len - nrow(data), ")\n")
  }
  
  if("Width_mm" %in% names(data)) {
    before_wid <- nrow(data)
    data <- data[data$Width_mm >= 2 & data$Width_mm <= 30 & !is.na(data$Width_mm), ]
    cat("  After width filtering (2-30mm):", nrow(data), "rows (removed", before_wid - nrow(data), ")\n")
  }
  
  # Step 4: Calculate derived metrics
  if(all(c("Length_mm", "Width_mm") %in% names(data))) {
    data$Geometric_Mean <- sqrt(data$Length_mm * data$Width_mm)
    data$Globosity_WoverL <- data$Width_mm / data$Length_mm
    
    # Filter globosity
    before_glob <- nrow(data)
    data <- data[data$Globosity_WoverL >= 0.3 & data$Globosity_WoverL <= 0.8 & !is.na(data$Globosity_WoverL), ]
    cat("  After globosity filtering:", nrow(data), "rows (removed", before_glob - nrow(data), ")\n")
  }
  
  # Step 5: Group by lot
  if("Lot_Number" %in% names(data)) {
    before_lot <- nrow(data)
    # Calculate specimens per lot
    lot_counts <- table(data$Lot_Number)
    lots_to_keep <- names(lot_counts[lot_counts > 2])
    
    # Keep lots with >2 specimens OR iNaturalist
    data <- data[data$Lot_Number %in% lots_to_keep | data$Collection == "iNat" | is.na(data$Collection), ]
    cat("  After lot filtering:", nrow(data), "rows (removed", before_lot - nrow(data), ")\n")
  }
  
  # Step 6: Create factor variables
  if("Epoch" %in% names(data)) {
    data$Epoch <- factor(data$Epoch, levels = c("Pleistocene", "Holocene", "Anthropocene"))
  }
  
  if("Collection" %in% names(data)) {
    data$Data_Type <- ifelse(data$Collection == "iNat", "iNaturalist", "Museum/Field")
    data$Data_Type <- factor(data$Data_Type)
  }
  
  if("Ecoregion" %in% names(data)) {
    data$Ecoregion <- factor(data$Ecoregion)
  }
  
  if("Coastal_Proximity" %in% names(data)) {
    data$Coastal_Proximity <- factor(data$Coastal_Proximity)
  }
  
  cat("\n  Cleaning summary:\n")
  cat("    Initial rows:", initial_rows, "\n")
  cat("    Final rows:", nrow(data), "\n")
  cat("    Rows removed:", initial_rows - nrow(data), 
      sprintf("(%.1f%%)", (initial_rows - nrow(data))/initial_rows * 100), "\n")
  
  # Return both raw and cleaned for comparison
  return(list(cleaned = data, raw = raw_data_saved))
}

# Function to create SoCal subset
create_socal_subset <- function(data) {
  
  cat("\nCreating Southern California subset...\n")
  
  socal_counties <- c("Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego")
  
  if("County" %in% names(data)) {
    SoCal <- data[data$County %in% socal_counties, ]
    
    cat("  SoCal subset created:\n")
    cat("    Rows:", nrow(SoCal), "\n")
    cat("    Percentage of full dataset:", 
        sprintf("%.1f%%", nrow(SoCal)/nrow(data) * 100), "\n")
    
    return(SoCal)
  } else {
    cat("  Warning: County column not found. Cannot create SoCal subset.\n")
    return(NULL)
  }
}

# Main data loading and cleaning
cat("\n--- Starting data loading pipeline ---\n")

# Load data
data_list <- load_data_from_github()

# Clean data
clean_results <- clean_raw_data(data_list$raw, data_list$key)
Clean <- clean_results$cleaned
Raw <- clean_results$raw

# Save raw data for reference
write.csv(Raw, "data/Data-Raw-Reference.csv", row.names = FALSE)
cat("\n✓ Saved raw reference dataset to: data/Data-Raw-Reference.csv\n")

# Save cleaned data
write.csv(Clean, "data/Data_Clean.csv", row.names = FALSE)
cat("✓ Saved Clean dataset to: data/Data_Clean.csv\n")

# Create SoCal subset
SoCal <- create_socal_subset(Clean)

if(!is.null(SoCal)) {
  write.csv(SoCal, "data/Data_SoCal.csv", row.names = FALSE)
  cat("✓ Saved SoCal subset to: data/Data_SoCal.csv\n")
}

# Display final summary
cat("\n=== DATASET SUMMARY ===\n")
cat("Raw dataset dimensions:", nrow(Raw), "rows,", ncol(Raw), "columns\n")
cat("Clean dataset dimensions:", nrow(Clean), "rows,", ncol(Clean), "columns\n")
if(!is.null(SoCal)) cat("SoCal dataset dimensions:", nrow(SoCal), "rows,", ncol(SoCal), "columns\n")

if("Epoch" %in% names(Clean)) {
  cat("\nSpecimens by epoch (Clean):\n")
  print(table(Clean$Epoch))
}

# ==============================================================================
# SECTION 4: MAP VISUALIZATIONS
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 4: CREATING DISTRIBUTION MAPS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Function to create distribution maps
create_distribution_maps <- function(raw_data, clean_data, socal_data) {
  
  cat("Creating distribution maps...\n")
  
  # Get base map data for broader region (includes Mexico)
  west_coast_map <- map_data("world", region = c("USA", "Mexico"))
  california_map <- map_data("state", region = "california")
  baja_map <- map_data("world", region = "Mexico")
  
  # Get map bounds for full distribution
  all_lats <- c(raw_data$Latitude, clean_data$Latitude)
  all_lons <- c(raw_data$Longitude, clean_data$Longitude)
  
  lat_range <- range(all_lats, na.rm = TRUE) + c(-2, 2)
  lon_range <- range(all_lons, na.rm = TRUE) + c(-2, 2)
  
  # Map 1: Raw data distribution
  cat("  Creating raw data distribution map...\n")
  
  p1 <- ggplot() +
    # Base map
    geom_polygon(data = west_coast_map, aes(x = long, y = lat, group = group),
                 fill = "lightgray", color = "white", size = 0.2) +
    # Points colored by epoch
    geom_point(data = raw_data, aes(x = Longitude, y = Latitude, color = Epoch),
               alpha = 0.5, size = 1) +
    # Formatting
    scale_color_manual(values = c("Pleistocene" = "#00AFBB", 
                                   "Holocene" = "#E7B800", 
                                   "Anthropocene" = "#FC4E07",
                                   "Pliocene" = "#993404")) +
    coord_fixed(xlim = lon_range, ylim = lat_range, ratio = 1.3) +
    labs(title = "Raw Data: All Sampling Locations",
         subtitle = paste("n =", nrow(raw_data), "specimens"),
         x = "Longitude", y = "Latitude",
         color = "Epoch") +
    theme_minimal() +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))
  
  ggsave("results/figures/Map1_Raw_Distribution.png", p1, width = 10, height = 8, dpi = 300)
  
  # Map 2: Clean data distribution
  cat("  Creating clean data distribution map...\n")
  
  p2 <- ggplot() +
    geom_polygon(data = west_coast_map, aes(x = long, y = lat, group = group),
                 fill = "lightgray", color = "white", size = 0.2) +
    geom_point(data = clean_data, aes(x = Longitude, y = Latitude, color = Epoch),
               alpha = 0.6, size = 1.2) +
    scale_color_manual(values = c("Pleistocene" = "#00AFBB", 
                                   "Holocene" = "#E7B800", 
                                   "Anthropocene" = "#FC4E07")) +
    coord_fixed(xlim = lon_range, ylim = lat_range, ratio = 1.3) +
    labs(title = "Clean Data: Sampling Locations After Filtering",
         subtitle = paste("n =", nrow(clean_data), "specimens"),
         x = "Longitude", y = "Latitude",
         color = "Epoch") +
    theme_minimal() +
    theme(legend.position = "bottom",
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5))
  
  ggsave("results/figures/Map2_Clean_Distribution.png", p2, width = 10, height = 8, dpi = 300)
  
  # Map 3: SoCal zoom if data available
  if(!is.null(socal_data) && nrow(socal_data) > 0) {
    cat("  Creating Southern California zoom map...\n")
    
    # SoCal bounds
    socal_lat_range <- range(socal_data$Latitude, na.rm = TRUE) + c(-0.3, 0.3)
    socal_lon_range <- range(socal_data$Longitude, na.rm = TRUE) + c(-0.3, 0.3)
    
    # Add density contours
    p3 <- ggplot() +
      # Base map
      geom_polygon(data = west_coast_map, aes(x = long, y = lat, group = group),
                   fill = "lightgray", color = "darkgray", size = 0.2) +
      # Density contours
      stat_density2d(data = socal_data, aes(x = Longitude, y = Latitude, fill = after_stat(density)),
                     geom = "tile", contour = FALSE, alpha = 0.5) +
      scale_fill_gradient(low = "yellow", high = "red", name = "Density") +
      # Points
      geom_point(data = socal_data, aes(x = Longitude, y = Latitude, color = Epoch),
                 alpha = 0.7, size = 1.5) +
      # Channel Islands label
      annotate("text", x = -119.8, y = 34.0, label = "Channel Islands", 
               size = 3, fontface = "italic") +
      # Formatting
      coord_fixed(xlim = socal_lon_range, ylim = socal_lat_range, ratio = 1.3) +
      scale_color_manual(values = c("Pleistocene" = "#00AFBB", 
                                     "Holocene" = "#E7B800", 
                                     "Anthropocene" = "#FC4E07")) +
      labs(title = "Southern California: Zoomed View with Density",
           subtitle = paste("n =", nrow(socal_data), "specimens"),
           x = "Longitude", y = "Latitude",
           color = "Epoch") +
      theme_minimal() +
      theme(legend.position = "bottom",
            plot.title = element_text(hjust = 0.5),
            plot.subtitle = element_text(hjust = 0.5))
    
    ggsave("results/figures/Map3_SoCal_Zoom_Density.png", p3, width = 10, height = 8, dpi = 300)
  }
  
  cat("✓ All maps saved to results/figures/\n")
}

# Create maps
create_distribution_maps(Raw, Clean, SoCal)

# ==============================================================================
# SECTION 5: PRE-ANALYSIS DIAGNOSTICS
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 5: PRE-ANALYSIS DIAGNOSTICS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Function for distribution assessment
assess_distribution <- function(data, variable, group_var = NULL) {
  
  cat(paste0("  Assessing ", variable, "...\n"))
  
  # Extract values and ensure they're numeric
  values <- data[[variable]]
  
  # Convert to numeric if needed
  if(!is.numeric(values)) {
    values <- as.numeric(as.character(values))
  }
  
  # Remove NAs
  values <- values[!is.na(values) & is.finite(values)]
  
  if(length(values) == 0) {
    return(list(n = 0, mean = NA, sd = NA, skewness = NA, kurtosis = NA, 
                ad_p = NA, is_normal = FALSE, values = numeric(0)))
  }
  
  # Statistical tests
  ad_test <- tryCatch(ad.test(values), error = function(e) list(p.value = NA))
  
  # Skewness and kurtosis
  skew_val <- tryCatch(skewness(values), error = function(e) NA)
  kurt_val <- tryCatch(kurtosis(values), error = function(e) NA)
  
  return(list(
    variable = variable,
    n = length(values),
    mean = mean(values),
    median = median(values),
    sd = sd(values),
    skewness = skew_val,
    kurtosis = kurt_val,
    ad_p = ifelse(is.null(ad_test$p.value), NA, ad_test$p.value),
    is_normal = ifelse(!is.na(ad_test$p.value), ad_test$p.value > 0.05, FALSE),
    values = values
  ))
}

# Run distribution assessments
cat("Assessing variable distributions...\n")

dist_length <- assess_distribution(Clean, "Length_mm", "Epoch")
dist_width <- assess_distribution(Clean, "Width_mm", "Epoch")
dist_glob <- assess_distribution(Clean, "Globosity_WoverL", "Epoch")

# Power analysis
cat("\nPerforming power analysis...\n")

n_total <- nrow(Clean)
if(n_total > 0) {
  power_cor_small <- tryCatch(pwr.r.test(n = n_total, r = 0.1, sig.level = 0.05)$power, error = function(e) NA)
  power_cor_medium <- tryCatch(pwr.r.test(n = n_total, r = 0.3, sig.level = 0.05)$power, error = function(e) NA)
  min_detectable_r <- tryCatch(pwr.r.test(n = n_total, power = 0.80, sig.level = 0.05)$r, error = function(e) NA)
  
  # Save power results
  sink("results/summary/Power_Analysis_Results.txt")
  cat("POWER ANALYSIS RESULTS\n")
  cat("======================\n\n")
  cat("Total sample size:", n_total, "\n\n")
  cat("Power to detect correlations:\n")
  cat("  Small effect (r = 0.1):", ifelse(!is.na(power_cor_small), 
                                          paste0(round(power_cor_small * 100, 1), "%"), "N/A"), "\n")
  cat("  Medium effect (r = 0.3):", ifelse(!is.na(power_cor_medium), 
                                           paste0(round(power_cor_medium * 100, 1), "%"), "N/A"), "\n")
  cat("  Minimum detectable r with 80% power:", 
      ifelse(!is.na(min_detectable_r), round(min_detectable_r, 3), "N/A"), "\n\n")
  
  cat("RECOMMENDATIONS:\n")
  if(!is.null(dist_length) && !dist_length$is_normal) {
    cat("  - Use non-parametric tests (Spearman correlation)\n")
    cat("  - Consider robust regression (MASS::rlm)\n")
  }
  sink()
}

# Create null datasets
cat("\nCreating null datasets for validation...\n")

set.seed(12345)
n_perm <- 1000

# 1. Complete randomization
null_complete <- Clean
null_complete$Length_mm <- sample(null_complete$Length_mm)

# 2. Epoch-stratified randomization
null_epoch <- Clean
if("Epoch" %in% names(Clean)) {
  for(epoch in unique(Clean$Epoch)) {
    idx <- which(Clean$Epoch == epoch)
    if(length(idx) > 1) {
      null_epoch$Length_mm[idx] <- sample(Clean$Length_mm[idx])
    }
  }
}

# 3. Latitude-stratified
null_stratified <- Clean
lat_bins <- cut(Clean$Latitude, breaks = 10)
for(bin in levels(lat_bins)) {
  idx <- which(lat_bins == bin)
  if(length(idx) > 1) {
    null_stratified$Length_mm[idx] <- sample(Clean$Length_mm[idx])
  }
}

# 4. Bootstrap
boot_idx <- sample(1:nrow(Clean), size = nrow(Clean), replace = TRUE)
null_bootstrap <- Clean[boot_idx, ]

# 5. Permutation distribution
perm_correlations <- numeric(n_perm)
for(i in 1:n_perm) {
  perm_data <- Clean
  perm_data$Length_mm <- sample(perm_data$Length_mm)
  perm_correlations[i] <- tryCatch({
    cor(perm_data$Length_mm, perm_data$Latitude, 
        method = "spearman", use = "complete.obs")
  }, error = function(e) NA)
}

# Save null datasets to data/null/
write.csv(null_complete, "data/null/NullDataset_Complete.csv", row.names = FALSE)
write.csv(null_epoch, "data/null/NullDataset_Epoch.csv", row.names = FALSE)
write.csv(null_stratified, "data/null/NullDataset_Stratified.csv", row.names = FALSE)
write.csv(null_bootstrap, "data/null/NullDataset_Bootstrap.csv", row.names = FALSE)
write.csv(data.frame(perm_correlation = perm_correlations), 
          "data/null/Permutation_Distribution.csv", row.names = FALSE)
cat("✓ Null datasets saved to data/null/\n")

# Create distribution summary table
dist_summary <- data.frame(
  Variable = c("Length (mm)", "Width (mm)", "Globosity (W/L)"),
  N = c(dist_length$n, dist_width$n, dist_glob$n),
  Mean = round(c(dist_length$mean, dist_width$mean, dist_glob$mean), 2),
  Median = round(c(dist_length$median, dist_width$median, dist_glob$median), 2),
  SD = round(c(dist_length$sd, dist_width$sd, dist_glob$sd), 2),
  Skewness = round(c(dist_length$skewness, dist_width$skewness, dist_glob$skewness), 3),
  Kurtosis = round(c(dist_length$kurtosis, dist_width$kurtosis, dist_glob$kurtosis), 3),
  AD_p = round(c(dist_length$ad_p, dist_width$ad_p, dist_glob$ad_p), 4),
  Normal = c(dist_length$is_normal, dist_width$is_normal, dist_glob$is_normal)
)

write.csv(dist_summary, "results/TableS1_Distribution_Summary.csv", row.names = FALSE)
cat("\n✓ Distribution summary saved to results/TableS1_Distribution_Summary.csv\n")

# Create pre-analysis summary
sink("results/summary/PreAnalysis_Summary.txt")
cat("PRE-ANALYSIS DATA QUALITY REPORT\n")
cat("================================\n\n")
cat("DATA SUMMARY:\n")
cat("  Total specimens (Clean):", nrow(Clean), "\n")
cat("  By epoch:", paste(names(table(Clean$Epoch)), table(Clean$Epoch), sep = "=", collapse = ", "), "\n\n")
cat("NORMALITY ASSESSMENT:\n")
cat("  Length: p =", format.pval(dist_length$ad_p, digits = 3), 
    ifelse(dist_length$is_normal, " (NORMAL)", " (NON-NORMAL)"), "\n")
cat("  Width: p =", format.pval(dist_width$ad_p, digits = 3),
    ifelse(dist_width$is_normal, " (NORMAL)", " (NON-NORMAL)"), "\n")
cat("  Globosity: p =", format.pval(dist_glob$ad_p, digits = 3),
    ifelse(dist_glob$is_normal, " (NORMAL)", " (NON-NORMAL)"), "\n\n")
cat("RECOMMENDATIONS:\n")
if(!dist_length$is_normal) {
  cat("  - Use Spearman correlation (non-parametric)\n")
  cat("  - Report effect sizes with confidence intervals\n")
}
sink()

# ==============================================================================
# SECTION 6: MAIN ANALYSIS - RESEARCH QUESTIONS
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 6: MAIN ANALYSIS - RESEARCH QUESTIONS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Initialize results dataframes
bergmann_results <- data.frame()
shape_results <- data.frame()

# Q1: Overall Bergmann's Rule
cat("\n--- Q1: Testing Bergmann's Rule (Length ~ Latitude) ---\n")

spearman_all <- tryCatch({
  cor.test(Clean$Length_mm, Clean$Latitude, method = "spearman", exact = FALSE)
}, error = function(e) NULL)

if(!is.null(spearman_all)) {
  # Bootstrap CI
  set.seed(12345)
  boot_cors <- numeric(1000)
  for(i in 1:1000) {
    boot_idx <- sample(1:nrow(Clean), size = nrow(Clean), replace = TRUE)
    boot_data <- Clean[boot_idx, ]
    boot_cors[i] <- cor(boot_data$Length_mm, boot_data$Latitude, method = "spearman", use = "complete.obs")
  }
  ci_lower <- quantile(boot_cors, 0.025, na.rm = TRUE)
  ci_upper <- quantile(boot_cors, 0.975, na.rm = TRUE)
  
  bergmann_results <- rbind(bergmann_results, data.frame(
    Model = "Length ~ Latitude (Overall)",
    N = nrow(Clean),
    Slope = NA,  # Spearman doesn't give slope
    Rho = round(spearman_all$estimate, 4),
    CI_lower = round(ci_lower, 4),
    CI_upper = round(ci_upper, 4),
    P_value = format.pval(spearman_all$p.value, digits = 3),
    Supports_Bergmann = ifelse(spearman_all$p.value < 0.05, "Yes", "No")
  ))
}

# Q1a: By epoch
cat("\n--- Q1a: By Epoch ---\n")

if("Epoch" %in% names(Clean)) {
  for(epoch in levels(Clean$Epoch)) {
    epoch_data <- Clean[Clean$Epoch == epoch, ]
    if(nrow(epoch_data) > 5) {
      test <- tryCatch({
        cor.test(epoch_data$Length_mm, epoch_data$Latitude, method = "spearman", exact = FALSE)
      }, error = function(e) NULL)
      
      if(!is.null(test)) {
        bergmann_results <- rbind(bergmann_results, data.frame(
          Model = paste("Length ~ Latitude (", epoch, ")", sep = ""),
          N = nrow(epoch_data),
          Slope = NA,
          Rho = round(test$estimate, 4),
          CI_lower = NA,
          CI_upper = NA,
          P_value = format.pval(test$p.value, digits = 3),
          Supports_Bergmann = ifelse(test$p.value < 0.05, "Yes", "No")
        ))
        
        cat("  ", epoch, ": ρ =", round(test$estimate, 4), 
            ", p =", format.pval(test$p.value, digits = 3), "\n")
      }
    }
  }
}

# Test for interaction using linear model
if("Epoch" %in% names(Clean) && length(unique(Clean$Epoch)) > 1) {
  lm_interaction <- lm(Length_mm ~ Latitude * Epoch, data = Clean)
  lm_main <- lm(Length_mm ~ Latitude + Epoch, data = Clean)
  anova_result <- anova(lm_main, lm_interaction)
  
  bergmann_results <- rbind(bergmann_results, data.frame(
    Model = "Length ~ Latitude * Epoch (Interaction)",
    N = nrow(Clean),
    Slope = NA,
    Rho = NA,
    CI_lower = NA,
    CI_upper = NA,
    P_value = format.pval(anova_result$`Pr(>F)`[2], digits = 3),
    Supports_Bergmann = ifelse(anova_result$`Pr(>F)`[2] < 0.05, "Varies", "Consistent")
  ))
}

# Q1b: SoCal mainland vs island
cat("\n--- Q1b: SoCal Mainland vs Island ---\n")

if(exists("SoCal") && !is.null(SoCal) && "Coastal_Proximity" %in% names(SoCal)) {
  for(prox in levels(SoCal$Coastal_Proximity)) {
    prox_data <- SoCal[SoCal$Coastal_Proximity == prox, ]
    if(nrow(prox_data) > 5) {
      test <- tryCatch({
        cor.test(prox_data$Length_mm, prox_data$Latitude, method = "spearman", exact = FALSE)
      }, error = function(e) NULL)
      
      if(!is.null(test)) {
        bergmann_results <- rbind(bergmann_results, data.frame(
          Model = paste("SoCal -", prox),
          N = nrow(prox_data),
          Slope = NA,
          Rho = round(test$estimate, 4),
          CI_lower = NA,
          CI_upper = NA,
          P_value = format.pval(test$p.value, digits = 3),
          Supports_Bergmann = ifelse(test$p.value < 0.05, "Yes", "No")
        ))
        
        cat("  ", prox, ": ρ =", round(test$estimate, 4), 
            ", p =", format.pval(test$p.value, digits = 3), "\n")
      }
    }
  }
}

# Q2: Shell Shape
cat("\n--- Q2: Shell Shape (Globosity ~ Latitude) ---\n")

spearman_shape <- tryCatch({
  cor.test(Clean$Globosity_WoverL, Clean$Latitude, method = "spearman", exact = FALSE)
}, error = function(e) NULL)

if(!is.null(spearman_shape)) {
  shape_results <- rbind(shape_results, data.frame(
    Model = "Globosity ~ Latitude (Overall)",
    N = nrow(Clean),
    Rho = round(spearman_shape$estimate, 4),
    P_value = format.pval(spearman_shape$p.value, digits = 3),
    Significant = ifelse(spearman_shape$p.value < 0.05, "Yes", "No")
  ))
}

# Q2a: Shape by epoch
cat("\n--- Q2a: Shape by Epoch ---\n")

if("Epoch" %in% names(Clean)) {
  for(epoch in levels(Clean$Epoch)) {
    epoch_data <- Clean[Clean$Epoch == epoch, ]
    if(nrow(epoch_data) > 5) {
      test <- tryCatch({
        cor.test(epoch_data$Globosity_WoverL, epoch_data$Latitude, method = "spearman", exact = FALSE)
      }, error = function(e) NULL)
      
      if(!is.null(test)) {
        shape_results <- rbind(shape_results, data.frame(
          Model = paste("Globosity ~ Latitude (", epoch, ")", sep = ""),
          N = nrow(epoch_data),
          Rho = round(test$estimate, 4),
          P_value = format.pval(test$p.value, digits = 3),
          Significant = ifelse(test$p.value < 0.05, "Yes", "No")
        ))
        
        cat("  ", epoch, ": ρ =", round(test$estimate, 4), 
            ", p =", format.pval(test$p.value, digits = 3), "\n")
      }
    }
  }
}

# Save results tables
write.csv(bergmann_results, "results/Table1_Bergmanns_Rule.csv", row.names = FALSE)
write.csv(shape_results, "results/Table2_Shape_Results.csv", row.names = FALSE)
cat("\n✓ Results tables saved to results/\n")

# ==============================================================================
# SECTION 7: VISUALIZATIONS
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 7: CREATING VISUALIZATIONS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Define colors
epoch_colors <- c("Pleistocene" = "#00AFBB", 
                  "Holocene" = "#E7B800", 
                  "Anthropocene" = "#FC4E07")

# Figure 1: Bergmann's Rule
p1 <- ggplot(Clean, aes(x = Latitude, y = Length_mm, color = Epoch)) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = TRUE, aes(fill = Epoch), alpha = 0.2) +
  scale_color_manual(values = epoch_colors) +
  scale_fill_manual(values = epoch_colors) +
  labs(title = "Bergmann's Rule: Shell Length vs Latitude",
       subtitle = paste("n =", nrow(Clean), "specimens"),
       x = "Latitude (°N)", y = "Shell Length (mm)",
       color = "Epoch", fill = "Epoch") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("results/figures/Figure1_Bergmanns_Rule.png", p1, width = 10, height = 6, dpi = 300)

# Figure 2: Shell Shape
p2 <- ggplot(Clean, aes(x = Latitude, y = Globosity_WoverL, color = Epoch)) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = TRUE, aes(fill = Epoch), alpha = 0.2) +
  scale_color_manual(values = epoch_colors) +
  scale_fill_manual(values = epoch_colors) +
  labs(title = "Shell Shape (Globosity) vs Latitude",
       subtitle = paste("n =", nrow(Clean), "specimens"),
       x = "Latitude (°N)", y = "Globosity (W/L ratio)",
       color = "Epoch", fill = "Epoch") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("results/figures/Figure2_Shell_Shape.png", p2, width = 10, height = 6, dpi = 300)

# Figure 3: Boxplots by epoch
p3a <- ggplot(Clean, aes(x = Epoch, y = Length_mm, fill = Epoch)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = epoch_colors) +
  labs(title = "Shell Length by Epoch", x = "Epoch", y = "Length (mm)") +
  theme_minimal() +
  theme(legend.position = "none")

p3b <- ggplot(Clean, aes(x = Epoch, y = Globosity_WoverL, fill = Epoch)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = epoch_colors) +
  labs(title = "Shell Shape by Epoch", x = "Epoch", y = "Globosity (W/L)") +
  theme_minimal() +
  theme(legend.position = "none")

p3 <- plot_grid(p3a, p3b, ncol = 2, labels = "AUTO")
ggsave("results/figures/Figure3_Epoch_Boxplots.png", p3, width = 12, height = 5, dpi = 300)

# Figure 4: Validation against null
if(length(perm_correlations) > 0) {
  perm_cors <- perm_correlations[!is.na(perm_correlations)]
  if(length(perm_cors) > 0) {
    empirical_p <- mean(abs(perm_cors) >= abs(spearman_all$estimate), na.rm = TRUE)
    
    p4 <- ggplot(data.frame(perm = perm_cors), aes(x = perm)) +
      geom_histogram(bins = 30, fill = "lightgray", color = "black", alpha = 0.7) +
      geom_vline(xintercept = spearman_all$estimate, color = "red", size = 1) +
      annotate("text", x = spearman_all$estimate, y = max(table(cut(perm_cors, 30))),
               label = paste("Observed ρ =", round(spearman_all$estimate, 3)),
               hjust = -0.1, vjust = 1, color = "red") +
      labs(title = "Permutation Test: Correlation under Null",
           subtitle = paste("Empirical p-value =", round(empirical_p, 4)),
           x = "Spearman Correlation (ρ)", y = "Frequency") +
      theme_minimal()
    
    ggsave("results/figures/Figure4_Permutation_Validation.png", p4, width = 8, height = 6, dpi = 300)
  }
}

cat("✓ All figures saved to results/figures/\n")

# ==============================================================================
# SECTION 8: FINAL REPORT
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 8: GENERATING FINAL REPORT\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Create comprehensive summary
sink("results/summary/Complete_Analysis_Summary.txt")

cat("================================================================\n")
cat("COMPLETE ANALYSIS SUMMARY: Californiconus californicus\n")
cat("Testing Bergmann's Rule Across Pleistocene, Holocene, and Anthropocene\n")
cat("================================================================\n\n")

cat("DATA SUMMARY:\n")
cat("  Raw data:", nrow(Raw), "specimens\n")
cat("  Clean data:", nrow(Clean), "specimens\n")
cat("  SoCal subset:", ifelse(!is.null(SoCal), nrow(SoCal), "N/A"), "\n\n")

cat("By Epoch (Clean data):\n")
print(table(Clean$Epoch))
cat("\n")

cat("================================================================\n")
cat("BERGMANN'S RULE RESULTS\n")
cat("================================================================\n\n")
print(bergmann_results)
cat("\n")

cat("================================================================\n")
cat("SHELL SHAPE RESULTS\n")
cat("================================================================\n\n")
print(shape_results)
cat("\n")

cat("================================================================\n")
cat("VALIDATION\n")
cat("================================================================\n\n")
if(exists("empirical_p")) {
  cat("Permutation test empirical p-value:", round(empirical_p, 4), "\n")
  cat("Observed effect is", ifelse(empirical_p < 0.05, "ROBUST", "NOT robust"), "to random chance\n")
}

cat("\n")
cat("================================================================\n")
cat("OUTPUT FILES\n")
cat("================================================================\n\n")
cat("Data files:\n")
cat("  data/Data-Raw-Reference.csv\n")
cat("  data/Data_Clean.csv\n")
if(!is.null(SoCal)) cat("  data/Data_SoCal.csv\n")
cat("  data/null/ - null datasets\n\n")

cat("Results tables:\n")
cat("  results/Table1_Bergmanns_Rule.csv\n")
cat("  results/Table2_Shape_Results.csv\n")
cat("  results/TableS1_Distribution_Summary.csv\n\n")

cat("Figures:\n")
cat("  results/figures/Map1_Raw_Distribution.png\n")
cat("  results/figures/Map2_Clean_Distribution.png\n")
if(!is.null(SoCal)) cat("  results/figures/Map3_SoCal_Zoom_Density.png\n")
cat("  results/figures/Figure1_Bergmanns_Rule.png\n")
cat("  results/figures/Figure2_Shell_Shape.png\n")
cat("  results/figures/Figure3_Epoch_Boxplots.png\n")
if(exists("p4")) cat("  results/figures/Figure4_Permutation_Validation.png\n\n")

cat("Summaries:\n")
cat("  results/summary/PreAnalysis_Summary.txt\n")
cat("  results/summary/Power_Analysis_Results.txt\n")
cat("  results/summary/Complete_Analysis_Summary.txt\n")

cat("\n")
cat("================================================================\n")
cat("ANALYSIS COMPLETE\n")
cat("================================================================\n")

sink()

cat("\n✓ Final report saved to: results/summary/Complete_Analysis_Summary.txt\n")
cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("ALL ANALYSES COMPLETE - CHECK results/ FOLDER FOR OUTPUTS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")


```

Figures and tables 
```{r}

# Set seed for reproducibility downstream
set.seed(12345)

# ==============================================================================
# DOWNSTREAM ANALYSIS: ADDRESSING RESEARCH QUESTIONS
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("MAIN ANALYSIS: ADDRESSING RESEARCH QUESTIONS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n\n")

# Load required packages
library(tidyverse)
library(ggplot2)
library(cowplot)
library(ggpubr)
library(broom)
library(effectsize)
library(MASS)  # for robust regression
library(lme4)  # for mixed models
library(emmeans)  # for post-hoc comparisons
library(pwr)   # for power calculations

# ==============================================================================
# SECTION 1: DATA PREPARATION
# ==============================================================================

cat("\n=== SECTION 1: DATA PREPARATION ===\n")

# Load cleaned data
if(file.exists("data/Data_Clean.csv")) {
  Clean <- read.csv("data/Data_Clean.csv", stringsAsFactors = FALSE)
  cat("✓ Loaded Clean dataset from data/Data_Clean.csv\n")
} else {
  stop("Clean dataset not found. Please run data cleaning first.")
}

# Load SoCal subset
if(file.exists("data/Data_SoCal.csv")) {
  SoCal <- read.csv("data/Data_SoCal.csv", stringsAsFactors = FALSE)
  cat("✓ Loaded SoCal subset from data/Data_SoCal.csv\n")
} else {
  cat("⚠ SoCal subset not found. Will create from Clean data.\n")
  socal_counties <- c("Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego")
  SoCal <- Clean %>% filter(County %in% socal_counties)
}

# Convert Epoch to factor with correct order
Clean$Epoch <- factor(Clean$Epoch, levels = c("Pleistocene", "Holocene", "Anthropocene"))
SoCal$Epoch <- factor(SoCal$Epoch, levels = c("Pleistocene", "Holocene", "Anthropocene"))

# Create Coastal_Proximity factor (ensure it has levels)
if("Coastal_Proximity" %in% names(SoCal)) {
  SoCal$Coastal_Proximity <- factor(SoCal$Coastal_Proximity)
  cat("✓ Coastal_Proximity levels:", paste(levels(SoCal$Coastal_Proximity), collapse = ", "), "\n")
}

# Create Location factor for SoCal
if("Location" %in% names(SoCal)) {
  SoCal$Location <- factor(SoCal$Location)
  cat("✓ Location has", length(unique(SoCal$Location)), "levels\n")
}

# Display dataset summary
cat("\nDataset Summary:\n")
cat("  Total specimens:", nrow(Clean), "\n")
cat("  By epoch:", paste(names(table(Clean$Epoch)), table(Clean$Epoch), sep = "=", collapse = ", "), "\n")
cat("  SoCal specimens:", nrow(SoCal), "\n")

# ==============================================================================
# SECTION 2: RESEARCH QUESTION 1 - BERGMANN'S RULE (LENGTH VS LATITUDE)
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 2: RESEARCH QUESTION 1 - BERGMANN'S RULE\n")
cat("Question: Is shell length correlated with latitude?\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")

# ------------------------------------------------------------------------------
# 2.1 OVERALL BERGMANN'S RULE TEST
# ------------------------------------------------------------------------------

cat("\n--- 2.1 Overall Bergmann's Rule Test ---\n")
cat("Based on pre-analysis: Data is NON-NORMAL, using Spearman correlation\n")

# Spearman correlation (non-parametric, robust to non-normality)
spearman_result <- cor.test(Clean$Length_mm, Clean$Latitude, 
                            method = "spearman", 
                            exact = FALSE)

cat("\nSpearman correlation results:\n")
cat("  ρ (rho) =", round(spearman_result$estimate, 4), "\n")
cat("  p-value =", format.pval(spearman_result$p.value, digits = 3), "\n")
cat("  Interpretation: ", 
    ifelse(spearman_result$p.value < 0.05, 
           "SIGNIFICANT - Bergmann's Rule supported", 
           "NOT SIGNIFICANT - Bergmann's Rule not supported"), "\n")

# Effect size interpretation
if(spearman_result$p.value < 0.05) {
  effect_strength <- ifelse(abs(spearman_result$estimate) < 0.1, "very weak",
                     ifelse(abs(spearman_result$estimate) < 0.3, "weak",
                     ifelse(abs(spearman_result$estimate) < 0.5, "moderate",
                            "strong")))
  cat("  Effect strength:", effect_strength, 
      "(ρ =", round(spearman_result$estimate, 3), ")\n")
}

# Store results
q1_results <- list()
q1_results$overall_spearman <- spearman_result

# ------------------------------------------------------------------------------
# 2.2 Q1a: BERGMANN'S RULE ACROSS EPOCHS
# ------------------------------------------------------------------------------

cat("\n--- 2.2 Q1a: Bergmann's Rule Across Epochs ---\n")
cat("Testing if the relationship between length and latitude varies by epoch\n")

# Visualize first
epoch_colors <- c("Pleistocene" = "#00AFBB", 
                  "Holocene" = "#E7B800", 
                  "Anthropocene" = "#FC4E07")

q1a_plot <- ggplot(Clean, aes(x = Latitude, y = Length_mm, color = Epoch)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = epoch_colors) +
  scale_fill_manual(values = epoch_colors) +
  labs(title = "Bergmann's Rule Across Epochs",
       subtitle = "Testing for epoch × latitude interaction",
       x = "Latitude (°N)", y = "Shell Length (mm)",
       color = "Epoch") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("Q1a_Bergmann_Across_Epochs.png", q1a_plot, 
       width = 10, height = 6, dpi = 300)

cat("✓ Visualization saved to: Q1a_Bergmann_Across_Epochs.png\n")

# Method 1: Robust regression (less sensitive to non-normality)
cat("\nRobust regression (MASS::rlm) - recommended for non-normal data:\n")
rlm_epoch <- tryCatch({
  rlm(Length_mm ~ Latitude * Epoch, data = Clean, maxit = 100)
}, error = function(e) {
  cat("⚠ Robust regression failed:", e$message, "\n")
  return(NULL)
})

if(!is.null(rlm_epoch)) {
  rlm_summary <- summary(rlm_epoch)
  print(rlm_summary)
  
  # Try to get interaction p-value using F-test
  tryCatch({
    # Compare models with and without interaction
    rlm_main <- rlm(Length_mm ~ Latitude + Epoch, data = Clean, maxit = 100)
    
    # Calculate F-statistic manually
    rss_full <- sum(residuals(rlm_epoch)^2)
    rss_reduced <- sum(residuals(rlm_main)^2)
    df_full <- length(coef(rlm_epoch))
    df_reduced <- length(coef(rlm_main))
    n <- nrow(Clean)
    
    f_stat <- ((rss_reduced - rss_full) / (df_full - df_reduced)) / 
              (rss_full / (n - df_full))
    interaction_p <- pf(f_stat, df_full - df_reduced, n - df_full, lower.tail = FALSE)
    
    cat("\nInteraction test (Latitude:Epoch):\n")
    cat("  F-statistic =", round(f_stat, 4), "\n")
    cat("  p-value =", format.pval(interaction_p, digits = 3), "\n")
    cat("  Interpretation: ", 
        ifelse(interaction_p < 0.05, 
               "SIGNIFICANT - Relationship varies by epoch", 
               "NOT SIGNIFICANT - Relationship consistent across epochs"), "\n")
    
    q1_results$interaction_p <- interaction_p
  }, error = function(e) {
    cat("⚠ Could not calculate interaction p-value:", e$message, "\n")
    interaction_p <- NA
    q1_results$interaction_p <- NA
  })
} else {
  interaction_p <- NA
  q1_results$interaction_p <- NA
}

# Method 2: Spearman correlations by epoch (always works)
cat("\nSpearman correlations by epoch:\n")
epoch_spearman <- list()
for(epoch in levels(Clean$Epoch)) {
  epoch_data <- Clean %>% filter(Epoch == epoch)
  if(nrow(epoch_data) > 10) {
    sp <- tryCatch({
      cor.test(epoch_data$Length_mm, epoch_data$Latitude, 
               method = "spearman", exact = FALSE)
    }, error = function(e) {
      cat("  ⚠ Spearman failed for", epoch, ":", e$message, "\n")
      return(NULL)
    })
    
    if(!is.null(sp)) {
      epoch_spearman[[epoch]] <- sp
      cat("  ", epoch, ": ρ =", round(sp$estimate, 4), 
          ", p =", format.pval(sp$p.value, digits = 3), 
          ifelse(sp$p.value < 0.05, " *", ""), "\n")
    }
  }
}

# Store results
q1_results$rlm_epoch <- rlm_epoch
q1_results$epoch_spearman <- epoch_spearman

# ------------------------------------------------------------------------------
# 2.3 Q1b: SOCAL SUBSET - MAINLAND VS ISLAND EFFECTS
# ------------------------------------------------------------------------------

cat("\n--- 2.3 Q1b: SoCal Subset - Mainland vs Island Effects ---\n")

# Check if Coastal_Proximity has sufficient levels
if("Coastal_Proximity" %in% names(SoCal) && 
   length(unique(na.omit(SoCal$Coastal_Proximity))) > 1) {
  
  cat("\nTesting mainland vs island differences in SoCal region\n")
  
  # Visualize
  q1b_plot <- ggplot(SoCal, aes(x = Latitude, y = Length_mm, 
                                 color = Coastal_Proximity)) +
    geom_point(alpha = 0.3, size = 1.5) +
    geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
    facet_wrap(~ Epoch, nrow = 1) +
    labs(title = "SoCal: Mainland vs Island Comparison",
         x = "Latitude (°N)", y = "Shell Length (mm)",
         color = "Coastal Proximity") +
    theme_minimal() +
    theme(legend.position = "bottom")
  
  ggsave("Q1b_SoCal_Mainland_vs_Island.png", q1b_plot,
         width = 14, height = 5, dpi = 300)
  
  # Mixed effects model with Location as random effect
  cat("\nMixed Effects Model (Location as random effect):\n")
  cat("Formula: Length_mm ~ Latitude + Coastal_Proximity + (1|Location)\n")
  
  # Check if Location has sufficient levels
  if("Location" %in% names(SoCal) && length(unique(SoCal$Location)) > 1) {
    lmer_model <- tryCatch({
      lmer(Length_mm ~ Latitude + Coastal_Proximity + (1|Location), data = SoCal)
    }, error = function(e) {
      cat("⚠ Mixed model failed:", e$message, "\n")
      return(NULL)
    })
    
    if(!is.null(lmer_model)) {
      print(summary(lmer_model))
      
      # Test fixed effects
      cat("\nFixed effects tests:\n")
      print(anova(lmer_model))
      
      # Pairwise comparisons
      cat("\nEstimated marginal means:\n")
      emmeans_result <- tryCatch({
        emmeans(lmer_model, ~ Coastal_Proximity)
      }, error = function(e) {
        cat("⚠ emmeans failed:", e$message, "\n")
        return(NULL)
      })
      
      if(!is.null(emmeans_result)) {
        print(emmeans_result)
        
        # Pairwise contrast
        pairs_result <- pairs(emmeans_result)
        print(pairs_result)
        q1_results$socal_emmeans <- emmeans_result
      }
      
      q1_results$socal_mixed <- lmer_model
    }
  } else {
    cat("⚠ Insufficient Location levels for mixed model. Using regular ANCOVA.\n")
    lm_socal <- lm(Length_mm ~ Latitude * Coastal_Proximity, data = SoCal)
    print(summary(lm_socal))
    q1_results$socal_lm <- lm_socal
  }
  
} else {
  cat("⚠ Coastal_Proximity variable not available or has insufficient levels.\n")
  cat("   Cannot test mainland vs island differences.\n")
}

# ==============================================================================
# SECTION 3: RESEARCH QUESTION 2 - SHELL SHAPE (GLOBOSITY VS LATITUDE)
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 3: RESEARCH QUESTION 2 - SHELL SHAPE (GLOBOSITY)\n")
cat("Question: Is shell shape (globosity) correlated with latitude?\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")

# ------------------------------------------------------------------------------
# 3.1 OVERALL SHAPE VS LATITUDE
# ------------------------------------------------------------------------------

cat("\n--- 3.1 Overall Shape vs Latitude ---\n")

# Spearman correlation (non-parametric)
spearman_shape <- cor.test(Clean$Globosity_WoverL, Clean$Latitude, 
                           method = "spearman", exact = FALSE)

cat("\nSpearman correlation results (Globosity ~ Latitude):\n")
cat("  ρ (rho) =", round(spearman_shape$estimate, 4), "\n")
cat("  p-value =", format.pval(spearman_shape$p.value, digits = 3), "\n")
cat("  Interpretation: ", 
    ifelse(spearman_shape$p.value < 0.05, 
           "SIGNIFICANT - Shape varies with latitude", 
           "NOT SIGNIFICANT - No latitudinal pattern"), "\n")

q2_results <- list()
q2_results$overall_spearman <- spearman_shape

# ------------------------------------------------------------------------------
# 3.2 Q2a: SHAPE VS LATITUDE ACROSS EPOCHS
# ------------------------------------------------------------------------------

cat("\n--- 3.2 Q2a: Shape vs Latitude Across Epochs ---\n")

# Visualize
q2a_plot <- ggplot(Clean, aes(x = Latitude, y = Globosity_WoverL, color = Epoch)) +
  geom_point(alpha = 0.2, size = 1) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = epoch_colors) +
  scale_fill_manual(values = epoch_colors) +
  labs(title = "Shell Shape (Globosity) Across Latitude and Epoch",
       x = "Latitude (°N)", y = "Globosity (W/L ratio)",
       color = "Epoch") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("Q2a_Shape_Across_Epochs.png", q2a_plot, width = 10, height = 6, dpi = 300)

# Robust regression for shape
rlm_shape <- tryCatch({
  rlm(Globosity_WoverL ~ Latitude * Epoch, data = Clean, maxit = 100)
}, error = function(e) {
  cat("⚠ Robust regression for shape failed:", e$message, "\n")
  return(NULL)
})

if(!is.null(rlm_shape)) {
  cat("\nRobust regression results (shape ~ latitude * epoch):\n")
  print(summary(rlm_shape))
  
  # Test interaction
  tryCatch({
    rlm_shape_main <- rlm(Globosity_WoverL ~ Latitude + Epoch, data = Clean, maxit = 100)
    
    rss_full <- sum(residuals(rlm_shape)^2)
    rss_reduced <- sum(residuals(rlm_shape_main)^2)
    df_full <- length(coef(rlm_shape))
    df_reduced <- length(coef(rlm_shape_main))
    n <- nrow(Clean)
    
    f_stat <- ((rss_reduced - rss_full) / (df_full - df_reduced)) / 
              (rss_full / (n - df_full))
    shape_interaction_p <- pf(f_stat, df_full - df_reduced, n - df_full, lower.tail = FALSE)
    
    cat("\nInteraction test (Latitude:Epoch): p =", 
        format.pval(shape_interaction_p, digits = 3), "\n")
    cat("Interpretation: ", 
        ifelse(shape_interaction_p < 0.05, 
               "SIGNIFICANT - Shape-latitude relationship varies by epoch", 
               "NOT SIGNIFICANT - Relationship consistent across epochs"), "\n")
    
    q2_results$shape_interaction_p <- shape_interaction_p
  }, error = function(e) {
    cat("⚠ Could not calculate shape interaction p-value:", e$message, "\n")
    q2_results$shape_interaction_p <- NA
  })
}

# Spearman by epoch
cat("\nSpearman correlations by epoch:\n")
epoch_shape <- list()
for(epoch in levels(Clean$Epoch)) {
  epoch_data <- Clean %>% filter(Epoch == epoch)
  if(nrow(epoch_data) > 10) {
    sp <- tryCatch({
      cor.test(epoch_data$Globosity_WoverL, epoch_data$Latitude, 
               method = "spearman", exact = FALSE)
    }, error = function(e) {
      cat("  ⚠ Spearman failed for", epoch, ":", e$message, "\n")
      return(NULL)
    })
    
    if(!is.null(sp)) {
      epoch_shape[[epoch]] <- sp
      cat("  ", epoch, ": ρ =", round(sp$estimate, 4), 
          ", p =", format.pval(sp$p.value, digits = 3), 
          ifelse(sp$p.value < 0.05, " *", ""), "\n")
    }
  }
}

q2_results$rlm_shape <- rlm_shape
q2_results$epoch_shape <- epoch_shape

# ------------------------------------------------------------------------------
# 3.3 Q2b: SOCAL SHAPE - MAINLAND VS ISLAND
# ------------------------------------------------------------------------------

cat("\n--- 3.3 Q2b: SoCal Shape - Mainland vs Island Effects ---\n")

if("Coastal_Proximity" %in% names(SoCal) && 
   length(unique(na.omit(SoCal$Coastal_Proximity))) > 1) {
  
  # Visualize
  q2b_plot <- ggplot(SoCal, aes(x = Latitude, y = Globosity_WoverL, 
                                 color = Coastal_Proximity)) +
    geom_point(alpha = 0.3, size = 1.5) +
    geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
    facet_wrap(~ Epoch, nrow = 1) +
    labs(title = "SoCal: Shape Comparison - Mainland vs Island",
         x = "Latitude (°N)", y = "Globosity (W/L)",
         color = "Coastal Proximity") +
    theme_minimal() +
    theme(legend.position = "bottom")
  
  ggsave("Q2b_SoCal_Shape_Mainland_vs_Island.png", q2b_plot,
         width = 14, height = 5, dpi = 300)
  
  # Mixed model for shape
  if("Location" %in% names(SoCal) && length(unique(SoCal$Location)) > 1) {
    lmer_shape <- tryCatch({
      lmer(Globosity_WoverL ~ Latitude + Coastal_Proximity + (1|Location), data = SoCal)
    }, error = function(e) {
      cat("⚠ Mixed model for shape failed:", e$message, "\n")
      return(NULL)
    })
    
    if(!is.null(lmer_shape)) {
      cat("\nMixed Effects Model for shape:\n")
      print(summary(lmer_shape))
      
      # Test fixed effects
      print(anova(lmer_shape))
      
      q2_results$socal_shape_mixed <- lmer_shape
    }
  }
}

# ------------------------------------------------------------------------------
# 3.4 Q2c: MUSEUM VS DIGITAL COMPARISON
# ------------------------------------------------------------------------------

cat("\n--- 3.4 Q2c: Museum vs Digital Comparison ---\n")
cat("Note: Your cleaned dataset contains only Museum/Field samples.\n")
cat("      To properly test museum vs digital, iNaturalist data would need to be included.\n")

if("Data_Type" %in% names(Clean) && length(unique(Clean$Data_Type)) > 1) {
  cat("\nComparing shape between Museum/Field and iNaturalist samples:\n")
  
  # Visual comparison
  q2c_plot <- ggplot(Clean, aes(x = Data_Type, y = Globosity_WoverL, fill = Data_Type)) +
    geom_boxplot(alpha = 0.7) +
    facet_wrap(~ Epoch) +
    labs(title = "Shape Comparison: Museum vs Digital Samples",
         x = "Data Source", y = "Globosity (W/L)") +
    theme_minimal() +
    theme(legend.position = "none")
  
  ggsave("Q2c_Museum_vs_Digital.png", q2c_plot, width = 10, height = 6, dpi = 300)
  
  # Statistical test
  wilcox_test <- wilcox.test(Globosity_WoverL ~ Data_Type, data = Clean)
  cat("\nWilcoxon rank-sum test (non-parametric):\n")
  cat("  p-value =", format.pval(wilcox_test$p.value, digits = 3), "\n")
  
  # With time as covariate
  cat("\nTesting with Year as covariate:\n")
  lm_time <- lm(Globosity_WoverL ~ Data_Type + Year, data = Clean)
  print(summary(lm_time))
  
  q2_results$museum_vs_digital <- list(
    wilcox = wilcox_test,
    lm_time = lm_time
  )
  
} else {
  cat("\n⚠ Only Museum/Field data present in cleaned dataset.\n")
  cat("   To address Q2c, you would need to:\n")
  cat("   1. Re-run data cleaning without filtering out iNaturalist samples\n")
  cat("   2. Or load the original raw data and compare data sources\n")
  
  # Create placeholder note
  q2_results$museum_vs_digital <- "Not tested - only Museum/Field data available"
}

# ==============================================================================
# SECTION 4: SUMMARY TABLES AND FIGURES
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 4: SUMMARY TABLES AND FIGURES\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")

# ------------------------------------------------------------------------------
# 4.1 SUMMARY TABLE FOR Q1 (BERGMANN'S RULE)
# ------------------------------------------------------------------------------

# Create summary table for Bergmann's Rule results
bergmann_table <- data.frame(
  Test = c("Overall Spearman", 
           "Pleistocene", 
           "Holocene", 
           "Anthropocene",
           "Epoch Interaction"),
  Correlation = c(
    round(spearman_result$estimate, 3),
    ifelse(!is.null(epoch_spearman$Pleistocene), round(epoch_spearman$Pleistocene$estimate, 3), NA),
    ifelse(!is.null(epoch_spearman$Holocene), round(epoch_spearman$Holocene$estimate, 3), NA),
    ifelse(!is.null(epoch_spearman$Anthropocene), round(epoch_spearman$Anthropocene$estimate, 3), NA),
    NA
  ),
  P_value = c(
    format.pval(spearman_result$p.value, digits = 3),
    ifelse(!is.null(epoch_spearman$Pleistocene), format.pval(epoch_spearman$Pleistocene$p.value, digits = 3), NA),
    ifelse(!is.null(epoch_spearman$Holocene), format.pval(epoch_spearman$Holocene$p.value, digits = 3), NA),
    ifelse(!is.null(epoch_spearman$Anthropocene), format.pval(epoch_spearman$Anthropocene$p.value, digits = 3), NA),
    ifelse(!is.null(q1_results$interaction_p) && !is.na(q1_results$interaction_p), 
           format.pval(q1_results$interaction_p, digits = 3), "NA")
  ),
  Significant = c(
    ifelse(spearman_result$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(epoch_spearman$Pleistocene) && epoch_spearman$Pleistocene$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(epoch_spearman$Holocene) && epoch_spearman$Holocene$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(epoch_spearman$Anthropocene) && epoch_spearman$Anthropocene$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(q1_results$interaction_p) && !is.na(q1_results$interaction_p) && q1_results$interaction_p < 0.05, "Yes", "No")
  ),
  Interpretation = c(
    ifelse(spearman_result$p.value < 0.05, 
           "Supports Bergmann's Rule", 
           "No support"),
    ifelse(!is.null(epoch_spearman$Pleistocene) && epoch_spearman$Pleistocene$p.value < 0.05, 
           "Supports", "No support"),
    ifelse(!is.null(epoch_spearman$Holocene) && epoch_spearman$Holocene$p.value < 0.05, 
           "Supports", "No support"),
    ifelse(!is.null(epoch_spearman$Anthropocene) && epoch_spearman$Anthropocene$p.value < 0.05, 
           "Supports", "No support"),
    ifelse(!is.null(q1_results$interaction_p) && !is.na(q1_results$interaction_p) && q1_results$interaction_p < 0.05, 
           "Relationship varies by epoch", 
           "Consistent across epochs")
  )
)

write.csv(bergmann_table, "Table1_Bergmanns_Rule_Results.csv", row.names = FALSE)
cat("\n✓ Table 1 saved to: Table1_Bergmanns_Rule_Results.csv\n")

# ------------------------------------------------------------------------------
# 4.2 SUMMARY TABLE FOR Q2 (SHAPE)
# ------------------------------------------------------------------------------

shape_table <- data.frame(
  Test = c("Overall Shape Spearman", 
           "Pleistocene", 
           "Holocene", 
           "Anthropocene",
           "Epoch Interaction"),
  Correlation = c(
    round(spearman_shape$estimate, 3),
    ifelse(!is.null(epoch_shape$Pleistocene), round(epoch_shape$Pleistocene$estimate, 3), NA),
    ifelse(!is.null(epoch_shape$Holocene), round(epoch_shape$Holocene$estimate, 3), NA),
    ifelse(!is.null(epoch_shape$Anthropocene), round(epoch_shape$Anthropocene$estimate, 3), NA),
    NA
  ),
  P_value = c(
    format.pval(spearman_shape$p.value, digits = 3),
    ifelse(!is.null(epoch_shape$Pleistocene), format.pval(epoch_shape$Pleistocene$p.value, digits = 3), NA),
    ifelse(!is.null(epoch_shape$Holocene), format.pval(epoch_shape$Holocene$p.value, digits = 3), NA),
    ifelse(!is.null(epoch_shape$Anthropocene), format.pval(epoch_shape$Anthropocene$p.value, digits = 3), NA),
    ifelse(!is.null(q2_results$shape_interaction_p) && !is.na(q2_results$shape_interaction_p), 
           format.pval(q2_results$shape_interaction_p, digits = 3), "NA")
  ),
  Significant = c(
    ifelse(spearman_shape$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(epoch_shape$Pleistocene) && epoch_shape$Pleistocene$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(epoch_shape$Holocene) && epoch_shape$Holocene$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(epoch_shape$Anthropocene) && epoch_shape$Anthropocene$p.value < 0.05, "Yes", "No"),
    ifelse(!is.null(q2_results$shape_interaction_p) && !is.na(q2_results$shape_interaction_p) && q2_results$shape_interaction_p < 0.05, "Yes", "No")
  ),
  Interpretation = c(
    ifelse(spearman_shape$p.value < 0.05, 
           "Shape varies with latitude", 
           "No latitudinal pattern"),
    ifelse(!is.null(epoch_shape$Pleistocene) && epoch_shape$Pleistocene$p.value < 0.05, 
           "Varies", "No pattern"),
    ifelse(!is.null(epoch_shape$Holocene) && epoch_shape$Holocene$p.value < 0.05, 
           "Varies", "No pattern"),
    ifelse(!is.null(epoch_shape$Anthropocene) && epoch_shape$Anthropocene$p.value < 0.05, 
           "Varies", "No pattern"),
    ifelse(!is.null(q2_results$shape_interaction_p) && !is.na(q2_results$shape_interaction_p) && q2_results$shape_interaction_p < 0.05, 
           "Pattern varies by epoch", 
           "Consistent across epochs")
  )
)

write.csv(shape_table, "Table2_Shape_Results.csv", row.names = FALSE)
cat("✓ Table 2 saved to: Table2_Shape_Results.csv\n")

# ------------------------------------------------------------------------------
# 4.3 COMBINED VISUALIZATION: FOUR-PANEL FIGURE
# ------------------------------------------------------------------------------

# Create comprehensive figure for manuscript
cat("\nCreating comprehensive four-panel figure...\n")

# Panel A: Bergmann's Rule (Length ~ Latitude)
pA <- ggplot(Clean, aes(x = Latitude, y = Length_mm, color = Epoch)) +
  geom_point(alpha = 0.1, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = epoch_colors) +
  labs(title = "A) Bergmann's Rule", 
       x = "Latitude (°N)", y = "Shell Length (mm)") +
  theme_minimal() +
  theme(legend.position = "none")

# Panel B: Shape (Globosity ~ Latitude)
pB <- ggplot(Clean, aes(x = Latitude, y = Globosity_WoverL, color = Epoch)) +
  geom_point(alpha = 0.1, size = 0.8) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.2) +
  scale_color_manual(values = epoch_colors) +
  labs(title = "B) Shell Shape", 
       x = "Latitude (°N)", y = "Globosity (W/L)") +
  theme_minimal() +
  theme(legend.position = "none")

# Panel C: Epoch comparison (boxplots)
pC <- ggplot(Clean, aes(x = Epoch, y = Length_mm, fill = Epoch)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(values = epoch_colors) +
  labs(title = "C) Size by Epoch", 
       x = "Epoch", y = "Shell Length (mm)") +
  theme_minimal() +
  theme(legend.position = "none")

# Panel D: SoCal comparison (if available)
if("Coastal_Proximity" %in% names(SoCal) && 
   length(unique(na.omit(SoCal$Coastal_Proximity))) > 1) {
  pD <- ggplot(SoCal, aes(x = Coastal_Proximity, y = Length_mm, fill = Coastal_Proximity)) +
    geom_boxplot(alpha = 0.7) +
    facet_wrap(~ Epoch) +
    labs(title = "D) SoCal: Mainland vs Island", 
         x = "Coastal Proximity", y = "Shell Length (mm)") +
    theme_minimal() +
    theme(legend.position = "none", axis.text.x = element_text(angle = 45, hjust = 1))
} else {
  pD <- ggplot() + 
    annotate("text", x = 0.5, y = 0.5, label = "Insufficient data for SoCal comparison") +
    theme_void() +
    labs(title = "D) SoCal comparison unavailable")
}

# Combine panels
figure4 <- plot_grid(pA, pB, pC, pD, ncol = 2, labels = "AUTO")
ggsave("Figure4_Combined_Results.png", figure4, width = 12, height = 10, dpi = 300)
cat("✓ Combined figure saved to: Figure4_Combined_Results.png\n")

# ==============================================================================
# SECTION 5: VALIDATION AGAINST NULL DATASETS
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 5: VALIDATION AGAINST NULL DATASETS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")

# Load null datasets if they exist
if(file.exists("data/NullDataset_Complete.csv")) {
  
  null_complete <- read.csv("data/NullDataset_Complete.csv")
  
  # Test correlation in null dataset
  null_cor <- tryCatch({
    cor.test(null_complete$Length_mm, null_complete$Latitude, 
             method = "spearman", exact = FALSE)
  }, error = function(e) {
    cat("⚠ Null correlation test failed:", e$message, "\n")
    return(NULL)
  })
  
  if(!is.null(null_cor)) {
    cat("\nValidation against completely randomized null dataset:\n")
    cat("  Observed ρ =", round(spearman_result$estimate, 4), "\n")
    cat("  Null ρ =", round(null_cor$estimate, 4), "\n")
    cat("  Null p-value =", format.pval(null_cor$p.value, digits = 3), "\n")
    cat("  Validation: ", 
        ifelse(abs(spearman_result$estimate) > abs(null_cor$estimate) * 2,
               "Observed effect is robust", 
               "Observed effect may be influenced by random variation"), "\n")
  }
  
  # Load permutation distributions
  if(file.exists("data/Permutation_Distributions.csv")) {
    perm_dist <- tryCatch({
      read.csv("data/Permutation_Distributions.csv")
    }, error = function(e) {
      cat("⚠ Could not load permutation distributions:", e$message, "\n")
      return(NULL)
    })
    
    if(!is.null(perm_dist) && nrow(perm_dist) > 0) {
      # Calculate empirical p-value
      empirical_p <- mean(abs(perm_dist$correlation) >= abs(spearman_result$estimate), na.rm = TRUE)
      cat("\nPermutation test empirical p-value:", empirical_p, "\n")
      
      # Plot permutation distribution
      png("Permutation_Distribution.png", width = 8, height = 6, units = "in", res = 300)
      hist(perm_dist$correlation, breaks = 30, 
           main = "Permutation Distribution of Correlations",
           xlab = "Spearman ρ", col = "lightgray")
      abline(v = spearman_result$estimate, col = "red", lwd = 2)
      abline(v = -spearman_result$estimate, col = "red", lwd = 2, lty = 2)
      legend("topright", legend = c("Observed ρ", "Null distribution"), 
             col = c("red", "black"), lwd = c(2, 1))
      dev.off()
      cat("✓ Permutation distribution plot saved\n")
    }
  }
}

# ==============================================================================
# SECTION 6: SUMMARY OF FINDINGS (WITH PROPER NA HANDLING)
# ==============================================================================

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("SECTION 6: SUMMARY OF FINDINGS\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")

cat("\n=== RESEARCH QUESTION 1: BERGMANN'S RULE ===\n")
cat("Q1: Is shell length correlated with latitude?\n")
if(spearman_result$p.value < 0.05) {
  cat("  ✓ YES - Significant correlation detected\n")
  cat("     ρ =", round(spearman_result$estimate, 4), 
      ", p =", format.pval(spearman_result$p.value, digits = 3), "\n")
} else {
  cat("  ✗ NO - No significant correlation detected\n")
}

cat("\nQ1a: Does this remain true across Epochs?\n")
# Safely check interaction_p
interaction_exists <- !is.null(q1_results$interaction_p) && !is.na(q1_results$interaction_p)

if(interaction_exists) {
  if(q1_results$interaction_p < 0.05) {
    cat("  ✓ Relationship VARIES significantly by epoch\n")
  } else {
    cat("  ✓ Relationship is CONSISTENT across epochs\n")
  }
  cat("  Interaction p-value =", format.pval(q1_results$interaction_p, digits = 3), "\n")
} else {
  cat("  ⚠ Could not calculate interaction p-value\n")
  cat("  Using epoch-specific correlations for interpretation:\n")
}

cat("  Epoch-specific correlations:\n")
for(epoch in names(epoch_spearman)) {
  if(!is.null(epoch_spearman[[epoch]])) {
    cat("    ", epoch, ": ρ =", round(epoch_spearman[[epoch]]$estimate, 4), 
        ifelse(epoch_spearman[[epoch]]$p.value < 0.05, "*", ""), "\n")
  }
}

cat("\nQ1b: Are there confounding factors in SoCal (mainland vs islands)?\n")
if(!is.null(q1_results$socal_mixed)) {
  cat("  ✓ Mixed model analysis completed\n")
  cat("     Check Table1_SoCal_Results.csv for details\n")
} else if(!is.null(q1_results$socal_lm)) {
  cat("  ✓ ANCOVA analysis completed\n")
  cat("     Check Table1_SoCal_Results.csv for details\n")
} else {
  cat("  ⚠ Insufficient data for mainland vs island comparison\n")
}

cat("\n=== RESEARCH QUESTION 2: SHELL SHAPE ===\n")
cat("Q2: Is shell shape (globosity) correlated with latitude?\n")
if(spearman_shape$p.value < 0.05) {
  cat("  ✓ YES - Significant correlation detected\n")
  cat("     ρ =", round(spearman_shape$estimate, 4), 
      ", p =", format.pval(spearman_shape$p.value, digits = 3), "\n")
} else {
  cat("  ✗ NO - No significant correlation detected\n")
}

cat("\nQ2a: Is this pattern consistent across Epochs?\n")
shape_interaction_exists <- !is.null(q2_results$shape_interaction_p) && !is.na(q2_results$shape_interaction_p)

if(shape_interaction_exists) {
  if(q2_results$shape_interaction_p < 0.05) {
    cat("  ✓ Shape-latitude relationship VARIES by epoch\n")
  } else {
    cat("  ✓ Relationship is CONSISTENT across epochs\n")
  }
  cat("  Interaction p-value =", format.pval(q2_results$shape_interaction_p, digits = 3), "\n")
} else {
  cat("  ⚠ Could not calculate interaction p-value\n")
  cat("  Using epoch-specific correlations for interpretation:\n")
}

cat("  Epoch-specific shape correlations:\n")
for(epoch in names(epoch_shape)) {
  if(!is.null(epoch_shape[[epoch]])) {
    cat("    ", epoch, ": ρ =", round(epoch_shape[[epoch]]$estimate, 4), 
        ifelse(epoch_shape[[epoch]]$p.value < 0.05, "*", ""), "\n")
  }
}

cat("\nQ2c: Museum vs Digital comparison?\n")
if(is.character(q2_results$museum_vs_digital)) {
  cat("  ⚠", q2_results$museum_vs_digital, "\n")
} else if(!is.null(q2_results$museum_vs_digital)) {
  cat("  ✓ Comparison completed\n")
  if(!is.null(q2_results$museum_vs_digital$wilcox)) {
    cat("     Wilcoxon test p =", 
        format.pval(q2_results$museum_vs_digital$wilcox$p.value, digits = 3), "\n")
  }
} else {
  cat("  ⚠ Museum vs digital comparison not performed\n")
}

cat("\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")
cat("ANALYSIS COMPLETE - Results saved to working directory\n")
cat(paste0(rep("=", 80), collapse = ""), "\n")

# Save all results
all_results <- list(
  q1 = q1_results,
  q2 = q2_results,
  bergmann_table = bergmann_table,
  shape_table = shape_table,
  timestamp = Sys.time()
)

saveRDS(all_results, "Complete_Analysis_Results.rds")
cat("\n✓ Complete results saved to: Complete_Analysis_Results.rds\n")

# List all output files
cat("\nOutput files generated:\n")
cat("  1. Q1a_Bergmann_Across_Epochs.png\n")
if(file.exists("Q1b_SoCal_Mainland_vs_Island.png")) cat("  2. Q1b_SoCal_Mainland_vs_Island.png\n")
cat("  3. Q2a_Shape_Across_Epochs.png\n")
if(file.exists("Q2b_SoCal_Shape_Mainland_vs_Island.png")) cat("  4. Q2b_SoCal_Shape_Mainland_vs_Island.png\n")
if(file.exists("Q2c_Museum_vs_Digital.png")) cat("  5. Q2c_Museum_vs_Digital.png\n")
cat("  6. Figure4_Combined_Results.png\n")
cat("  7. Table1_Bergmanns_Rule_Results.csv\n")
cat("  8. Table2_Shape_Results.csv\n")
if(file.exists("Permutation_Distribution.png")) cat("  9. Permutation_Distribution.png\n")
cat("  10. Complete_Analysis_Results.rds\n")


```



```{r}
# Set seed for reproducibility downstream
set.seed(12345)

# ==============================================================================
# 2. SIMPLIFIED MAP CREATION
# ==============================================================================

# Function to create simple maps without requiring shapefiles
create_simple_maps <- function(data) {
  
  cat("\n=== CREATING SIMPLE MAPS ===\n")
  
  # Remove geometry if it's an sf object
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  # Check for required columns
  if(!all(c("Longitude", "Latitude", "Epoch") %in% names(data_df))) {
    cat("Warning: Missing required columns for mapping\n")
    return(NULL)
  }
  
  # Define epoch colors
  epoch_colors <- c("Pleistocene" = "#00AFBB", 
                    "Holocene" = "#E7B800", 
                    "Anthropocene" = "#FC4E07")
  
  # Get California and Mexico map data
  cat("Loading map data...\n")
  
  # Get world data
  world <- map_data("world")
  
  # Filter for relevant regions
  study_area <- world %>%
    filter((region %in% c("USA", "Canada", "Mexico") & 
             long >= -130 & long <= -110 & lat >= 15 & lat <= 50) |
           (region == "Mexico" & long >= -130 & long <= -110))
  
  # Create base map
  base_map <- ggplot() +
    geom_polygon(data = study_area,
                 aes(x = long, y = lat, group = group),
                 fill = "gray90", color = "gray70", size = 0.2) +
    coord_fixed(xlim = c(-125, -110), ylim = c(22, 42), ratio = 1.3) +
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      axis.text = element_text(size = 10),
      axis.title = element_text(size = 12, face = "bold")
    )
  
  # Figure 1A: All collection points
  fig1a <- base_map +
    geom_point(data = data_df,
               aes(x = Longitude, y = Latitude, color = Epoch),
               alpha = 0.3, size = 1) +
    scale_color_manual(values = epoch_colors) +
    labs(title = "All Collection Locations",
         x = "Longitude", y = "Latitude",
         color = "Epoch") +
    theme(legend.position = "bottom")
  
  # Figure 1B: With density contours
  fig1b <- base_map +
    stat_density_2d(data = data_df,
                    aes(x = Longitude, y = Latitude, fill = ..level..),
                    geom = "polygon", alpha = 0.3, bins = 20) +
    geom_point(data = data_df,
               aes(x = Longitude, y = Latitude, color = Epoch),
               alpha = 0.3, size = 1) +
    scale_color_manual(values = epoch_colors) +
    scale_fill_gradient(low = "blue", high = "red", guide = "none") +
    labs(title = "Collection Density",
         x = "Longitude", y = "Latitude",
         color = "Epoch") +
    theme(legend.position = "bottom")
  
  # Combine plots
  figure1 <- plot_grid(fig1a, fig1b, 
                       labels = c("A", "B"), 
                       label_size = 14,
                       ncol = 2)
  
  # Save figure
  ggsave("Figure1_Collection_Maps.png", figure1, 
         width = 14, height = 6, dpi = 300, bg = "white")
  
  cat("✓ Figure 1 saved to: Figure1_Collection_Maps.png\n")
  
  return(list(figure1 = figure1))
}

# ==============================================================================
# 3. CORE ANALYSIS FUNCTIONS (COMPATIBLE WITH OLDER R VERSIONS)
# ==============================================================================

# Function to test statistical assumptions
test_assumptions <- function(data) {
  
  cat("\n=== TESTING STATISTICAL ASSUMPTIONS ===\n")
  
  # Remove geometry if data is sf
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  results <- list()
  
  # Check for required variables
  if(!"Length_mm" %in% names(data_df)) {
    cat("Warning: Length_mm not found in data\n")
    return(NULL)
  }
  
  # 1. Normality test
  cat("\n1. Normality test for Length_mm:\n")
  shapiro_test <- shapiro.test(data_df$Length_mm)
  cat("Shapiro-Wilk: W =", round(shapiro_test$statistic, 4),
      "p =", format.pval(shapiro_test$p.value, digits = 3), "\n")
  
  is_normal <- shapiro_test$p.value > 0.05
  cat("Data is", ifelse(is_normal, "NORMALLY", "NOT NORMALLY"), "distributed\n")
  
  # 2. Homogeneity of variance by epoch
  if("Epoch" %in% names(data_df) && length(unique(data_df$Epoch)) > 1) {
    cat("\n2. Homogeneity of variance by epoch:\n")
    levene_test <- leveneTest(Length_mm ~ Epoch, data = data_df)
    cat("Levene's test: F =", round(levene_test$`F value`[1], 4),
        "p =", format.pval(levene_test$`Pr(>F)`[1], digits = 3), "\n")
    
    equal_var <- levene_test$`Pr(>F)`[1] > 0.05
    cat("Variances are", ifelse(equal_var, "EQUAL", "UNEQUAL"), "across epochs\n")
  }
  
  # 3. Create diagnostic plots
  cat("\n3. Creating diagnostic plots...\n")
  
  # Histogram
  hist_plot <- ggplot(data_df, aes(x = Length_mm)) +
    geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", color = "black") +
    geom_density(alpha = 0.5, fill = "red") +
    labs(title = "Distribution of Shell Length",
         x = "Length (mm)", y = "Density") +
    theme_minimal()
  
  # QQ plot
  qq_data <- data.frame(
    Theoretical = qqnorm(data_df$Length_mm, plot.it = FALSE)$x,
    Sample = qqnorm(data_df$Length_mm, plot.it = FALSE)$y
  )
  
  qq_plot <- ggplot(qq_data, aes(x = Theoretical, y = Sample)) +
    geom_point(alpha = 0.5) +
    geom_qq_line(aes(sample = Sample), color = "red") +
    labs(title = "Normal Q-Q Plot",
         x = "Theoretical Quantiles", y = "Sample Quantiles") +
    theme_minimal()
  
  # Save diagnostic plots
  diagnostic_plots <- plot_grid(hist_plot, qq_plot, ncol = 2)
  ggsave("Figure_Diagnostics.png", diagnostic_plots, 
         width = 12, height = 5, dpi = 300)
  
  cat("✓ Diagnostic plots saved to: Figure_Diagnostics.png\n")
  
  return(results)
}

# Function to analyze Bergmann's Rule
analyze_bergmanns_rule <- function(data) {
  
  cat("\n=== ANALYZING BERGMANN'S RULE ===\n")
  
  # Remove geometry if data is sf
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  # Check required columns
  required <- c("Length_mm", "Latitude", "Epoch")
  missing <- setdiff(required, names(data_df))
  
  if(length(missing) > 0) {
    cat("Warning: Missing required columns:", paste(missing, collapse = ", "), "\n")
    return(NULL)
  }
  
  results <- list()
  
  # 1. Overall relationship
  cat("\n1. Overall relationship: Length ~ Latitude\n")
  lm_overall <- lm(Length_mm ~ Latitude, data = data_df)
  summary_overall <- summary(lm_overall)
  
  cat("Intercept:", round(coef(lm_overall)[1], 4), "\n")
  cat("Slope:", round(coef(lm_overall)[2], 4), "mm per degree latitude\n")
  cat("R-squared:", round(summary_overall$r.squared, 4), "\n")
  cat("Adjusted R-squared:", round(summary_overall$adj.r.squared, 4), "\n")
  cat("p-value:", format.pval(summary_overall$coefficients[2, 4], digits = 3), "\n")
  
  # 2. By epoch
  cat("\n2. By epoch: Length ~ Latitude * Epoch\n")
  lm_epoch <- lm(Length_mm ~ Latitude * Epoch, data = data_df)
  anova_epoch <- anova(lm_epoch)
  print(anova_epoch)
  
  # 3. Separate models for each epoch
  cat("\n3. Separate models for each epoch:\n")
  epoch_models <- list()
  
  for(epoch in levels(data_df$Epoch)) {
    epoch_data <- data_df %>% filter(Epoch == epoch)
    
    if(nrow(epoch_data) > 10) {  # Need sufficient data
      lm_epoch_specific <- lm(Length_mm ~ Latitude, data = epoch_data)
      epoch_models[[epoch]] <- lm_epoch_specific
      
      cat("\n", epoch, ":\n", sep = "")
      cat("  n =", nrow(epoch_data), "\n")
      cat("  Slope =", round(coef(lm_epoch_specific)[2], 4), "\n")
      cat("  R² =", round(summary(lm_epoch_specific)$r.squared, 4), "\n")
      cat("  p =", format.pval(summary(lm_epoch_specific)$coefficients[2, 4], digits = 3), "\n")
    }
  }
  
  # 4. Create Bergmann's Rule plot
  cat("\n4. Creating Bergmann's Rule visualization...\n")
  
  epoch_colors <- c("Pleistocene" = "#00AFBB", 
                    "Holocene" = "#E7B800", 
                    "Anthropocene" = "#FC4E07")
  
  bergmann_plot <- ggplot(data_df, aes(x = Latitude, y = Length_mm, color = Epoch)) +
    geom_point(alpha = 0.3, size = 1.5) +
    geom_smooth(aes(fill = Epoch), method = "lm", se = TRUE, alpha = 0.2) +
    geom_smooth(method = "lm", color = "black", se = FALSE, size = 0.5) +
    scale_color_manual(values = epoch_colors) +
    scale_fill_manual(values = epoch_colors) +
    labs(
      title = "Bergmann's Rule in Californiconus californicus",
      x = "Latitude (°N)",
      y = "Shell Length (mm)",
      subtitle = paste("Overall slope =", round(coef(lm_overall)[2], 4), 
                      "mm/degree, R² =", round(summary_overall$r.squared, 3))
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      axis.title = element_text(size = 12, face = "bold"),
      legend.position = "bottom",
      legend.title = element_text(face = "bold")
    )
  
  # Save plot
  ggsave("Figure_Bergmanns_Rule.png", bergmann_plot, 
         width = 10, height = 7, dpi = 300, bg = "white")
  
  cat("✓ Bergmann's Rule plot saved to: Figure_Bergmanns_Rule.png\n")
  
  # 5. Calculate effect sizes
  cat("\n5. Calculating effect sizes...\n")
  
  # Split data by median latitude
  lat_median <- median(data_df$Latitude, na.rm = TRUE)
  data_df$Lat_group <- ifelse(data_df$Latitude > lat_median, "High", "Low")
  
  if(length(unique(data_df$Lat_group)) > 1) {
    cohen_d_result <- cohens_d(Length_mm ~ Lat_group, data = data_df)
    cat("Cohen's d for latitude effect:", round(cohen_d_result$Cohens_d, 3), "\n")
    results$cohen_d <- cohen_d_result
  }
  
  # Store results
  results$models <- list(
    overall = lm_overall,
    epoch_interaction = lm_epoch,
    epoch_specific = epoch_models
  )
  
  results$plot <- bergmann_plot
  
  return(results)
}

# Function to create correlation matrix (compatible with older R)
create_correlation_analysis <- function(data) {
  
  cat("\n=== CREATING CORRELATION ANALYSIS ===\n")
  
  # Remove geometry if data is sf
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  # Select numeric variables (compatible way for older R)
  # Identify numeric columns
  numeric_cols <- sapply(data_df, is.numeric)
  
  # Exclude specific columns if they exist
  exclude_cols <- c("Specimen_Number", "Year", "Decade", "Month")
  numeric_cols[names(data_df) %in% exclude_cols] <- FALSE
  
  # Get numeric variables
  numeric_vars <- data_df[, numeric_cols, drop = FALSE]
  
  # Remove columns with too many NAs
  na_counts <- colSums(is.na(numeric_vars))
  numeric_vars <- numeric_vars[, na_counts < nrow(numeric_vars) * 0.5, drop = FALSE]
  
  if(ncol(numeric_vars) < 2) {
    cat("Warning: Not enough numeric variables for correlation analysis\n")
    cat("Numeric variables found:", paste(names(numeric_vars), collapse = ", "), "\n")
    return(NULL)
  }
  
  # Create correlation plot
  cor_plot <- chart.Correlation(numeric_vars, 
                               histogram = TRUE, 
                               pch = 19,
                               main = "Correlation Matrix of Morphological and Environmental Variables")
  
  # Save plot
  png("Figure_Correlation_Matrix.png", width = 10, height = 10, units = "in", res = 300)
  print(cor_plot)
  dev.off()
  
  cat("✓ Correlation matrix saved to: Figure_Correlation_Matrix.png\n")
  
  # Calculate specific correlations of interest
  cat("\nKey correlations:\n")
  
  if(all(c("Length_mm", "Width_mm") %in% names(data_df))) {
    cor_length_width <- cor.test(data_df$Length_mm, data_df$Width_mm, use = "complete.obs")
    cat("Length ~ Width: r =", round(cor_length_width$estimate, 4),
        "p =", format.pval(cor_length_width$p.value, digits = 3), "\n")
  }
  
  if(all(c("Length_mm", "Latitude") %in% names(data_df))) {
    cor_length_lat <- cor.test(data_df$Length_mm, data_df$Latitude, use = "complete.obs")
    cat("Length ~ Latitude: r =", round(cor_length_lat$estimate, 4),
        "p =", format.pval(cor_length_lat$p.value, digits = 3), "\n")
  }
  
  if(all(c("Length_mm", "Globosity_WoverL") %in% names(data_df))) {
    cor_length_shape <- cor.test(data_df$Length_mm, data_df$Globosity_WoverL, use = "complete.obs")
    cat("Length ~ Globosity: r =", round(cor_length_shape$estimate, 4),
        "p =", format.pval(cor_length_shape$p.value, digits = 3), "\n")
  }
  
  return(cor_plot)
}

# ==============================================================================
# 4. MAIN ANALYSIS PIPELINE
# ==============================================================================

run_analysis_pipeline <- function() {
  
  cat(paste0(rep("=", 60), collapse = ""), "\n")
  cat("CALIFORNICONUS CALIFORNICUS MORPHOLOGY ANALYSIS\n")
  cat("Bergmann's Rule across Pleistocene, Holocene, and Anthropocene\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n\n")
  
  # 1. Load and clean data
  data_list <- load_and_clean_data()
  Clean <- data_list$Clean
  
  cat("\n=== DATASET SUMMARY ===\n")
  cat("Total specimens:", nrow(Clean), "\n")
  
  if("Epoch" %in% names(Clean)) {
    cat("\nBy epoch:\n")
    epoch_table <- table(Clean$Epoch)
    print(epoch_table)
  }
  
  if("Data_Type" %in% names(Clean)) {
    cat("\nBy data type:\n")
    data_type_table <- table(Clean$Data_Type)
    print(data_type_table)
  }
  
  # 2. Create maps
  cat("\n=== CREATING VISUALIZATIONS ===\n")
  maps <- create_simple_maps(Clean)
  
  # 3. Test assumptions
  assumptions <- test_assumptions(Clean)
  
  # 4. Correlation analysis
  cor_plot <- create_correlation_analysis(Clean)
  
  # 5. Bergmann's Rule analysis
  bergmann_results <- analyze_bergmanns_rule(Clean)
  
  # 6. Generate summary statistics
  cat("\n=== GENERATING SUMMARY STATISTICS ===\n")
  
  if(inherits(Clean, "sf")) {
    Clean_df <- as.data.frame(Clean)
    # Remove geometry column if it exists
    if("geometry" %in% names(Clean_df)) {
      Clean_df$geometry <- NULL
    }
  } else {
    Clean_df <- Clean
  }
  
  # Create summary by epoch
  if("Epoch" %in% names(Clean_df) && "Length_mm" %in% names(Clean_df)) {
    summary_stats <- data.frame()
    
    for(epoch in levels(Clean_df$Epoch)) {
      epoch_data <- Clean_df %>% filter(Epoch == epoch)
      
      if(nrow(epoch_data) > 0) {
        epoch_summary <- data.frame(
          Epoch = epoch,
          N = nrow(epoch_data),
          Mean_Length = round(mean(epoch_data$Length_mm, na.rm = TRUE), 2),
          SD_Length = round(sd(epoch_data$Length_mm, na.rm = TRUE), 2),
          Min_Length = round(min(epoch_data$Length_mm, na.rm = TRUE), 2),
          Max_Length = round(max(epoch_data$Length_mm, na.rm = TRUE), 2),
          stringsAsFactors = FALSE
        )
        summary_stats <- rbind(summary_stats, epoch_summary)
      }
    }
    
    cat("\nSummary statistics by epoch:\n")
    print(summary_stats)
    
    write.csv(summary_stats, "Summary_Statistics.csv", row.names = FALSE)
    cat("\n✓ Summary statistics saved to: Summary_Statistics.csv\n")
  }
  
  # 7. Create model comparison table
  cat("\n=== CREATING MODEL COMPARISON ===\n")
  
  if(!is.null(bergmann_results)) {
    
    # Get sample sizes
    n_pleistocene <- sum(Clean_df$Epoch == "Pleistocene", na.rm = TRUE)
    n_holocene <- sum(Clean_df$Epoch == "Holocene", na.rm = TRUE)
    n_anthropocene <- sum(Clean_df$Epoch == "Anthropocene", na.rm = TRUE)
    
    model_comparison <- data.frame(
      Model = c("Length ~ Latitude (Overall)", 
                "Length ~ Latitude * Epoch",
                "Length ~ Latitude (Pleistocene)",
                "Length ~ Latitude (Holocene)",
                "Length ~ Latitude (Anthropocene)"),
      n = c(
        nrow(Clean_df),
        nrow(Clean_df),
        n_pleistocene,
        n_holocene,
        n_anthropocene
      ),
      Slope = c(
        round(coef(bergmann_results$models$overall)[2], 4),
        NA,
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(coef(bergmann_results$models$epoch_specific$Pleistocene)[2], 4) else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(coef(bergmann_results$models$epoch_specific$Holocene)[2], 4) else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(coef(bergmann_results$models$epoch_specific$Anthropocene)[2], 4) else NA
      ),
      R_squared = c(
        round(summary(bergmann_results$models$overall)$r.squared, 4),
        round(summary(bergmann_results$models$epoch_interaction)$r.squared, 4),
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(summary(bergmann_results$models$epoch_specific$Pleistocene)$r.squared, 4) else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(summary(bergmann_results$models$epoch_specific$Holocene)$r.squared, 4) else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(summary(bergmann_results$models$epoch_specific$Anthropocene)$r.squared, 4) else NA
      ),
      P_value = c(
        format.pval(summary(bergmann_results$models$overall)$coefficients[2, 4], digits = 3),
        if(!is.null(anova(bergmann_results$models$epoch_interaction))) 
          format.pval(anova(bergmann_results$models$epoch_interaction)$`Pr(>F)`[1], digits = 3) else NA,
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          format.pval(summary(bergmann_results$models$epoch_specific$Pleistocene)$coefficients[2, 4], digits = 3) else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          format.pval(summary(bergmann_results$models$epoch_specific$Holocene)$coefficients[2, 4], digits = 3) else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          format.pval(summary(bergmann_results$models$epoch_specific$Anthropocene)$coefficients[2, 4], digits = 3) else NA
      ),
      Supports_Bergmann = c(
        ifelse(summary(bergmann_results$models$overall)$coefficients[2, 4] < 0.05, "Yes", "No"),
        if(!is.null(anova(bergmann_results$models$epoch_interaction))) 
          ifelse(anova(bergmann_results$models$epoch_interaction)$`Pr(>F)`[3] < 0.05, "Varies", "Consistent") else NA,
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          ifelse(summary(bergmann_results$models$epoch_specific$Pleistocene)$coefficients[2, 4] < 0.05, "Yes", "No") else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          ifelse(summary(bergmann_results$models$epoch_specific$Holocene)$coefficients[2, 4] < 0.05, "Yes", "No") else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          ifelse(summary(bergmann_results$models$epoch_specific$Anthropocene)$coefficients[2, 4] < 0.05, "Yes", "No") else NA
      )
    )
    
    write.csv(model_comparison, "Model_Comparison.csv", row.names = FALSE)
    cat("\n✓ Model comparison saved to: Model_Comparison.csv\n")
    
    # Print key findings
    cat("\n=== KEY FINDINGS ===\n")
    
    overall_slope <- coef(bergmann_results$models$overall)[2]
    overall_p <- summary(bergmann_results$models$overall)$coefficients[2, 4]
    
    if(overall_p < 0.05) {
      cat(paste0("\n✓ BERGMANN'S RULE SUPPORTED: Shell length increases with latitude\n",
                 "  Slope: ", round(overall_slope, 4), " mm per degree latitude\n",
                 "  R²: ", round(summary(bergmann_results$models$overall)$r.squared, 3), "\n",
                 "  p-value: ", format.pval(overall_p, digits = 3), "\n"))
    } else {
      cat("\n✗ BERGMANN'S RULE NOT SUPPORTED: No significant relationship with latitude\n")
    }
    
    # Check epoch differences
    if(!is.null(anova(bergmann_results$models$epoch_interaction))) {
      epoch_p <- anova(bergmann_results$models$epoch_interaction)$`Pr(>F)`[3]
      if(!is.na(epoch_p) && epoch_p < 0.05) {
        cat("\n✓ EPOCH EFFECT: Relationship varies significantly between epochs\n")
      } else if(!is.na(epoch_p)) {
        cat("\n✗ NO EPOCH EFFECT: Relationship consistent across epochs\n")
      }
    }
  }
  
  # Return results
  results <- list(
    data = Clean,
    assumptions = assumptions,
    bergmann = bergmann_results,
    maps = maps,
    correlation = cor_plot
  )
  
  # Save results
  saveRDS(results, "Analysis_Results.rds")
  cat("\n✓ Complete analysis results saved to: Analysis_Results.rds\n")
  
  return(results)
}

# ==============================================================================
# 5. EXECUTE ANALYSIS
# ==============================================================================

# Run the analysis
cat("\n")
cat(paste0(rep("=", 60), collapse = ""), "\n")
cat("STARTING ANALYSIS\n")
cat(paste0(rep("=", 60), collapse = ""), "\n\n")

tryCatch({
  results <- run_analysis_pipeline()
  
  cat("\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n")
  cat("ANALYSIS COMPLETE!\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n\n")
  
  cat("Output files generated:\n")
  cat("1. Data_Clean.csv - Cleaned dataset\n")
  cat("2. Figure1_Collection_Maps.png - Collection location maps\n")
  cat("3. Figure_Diagnostics.png - Diagnostic plots\n")
  cat("4. Figure_Correlation_Matrix.png - Correlation matrix\n")
  cat("5. Figure_Bergmanns_Rule.png - Bergmann's Rule visualization\n")
  cat("6. Summary_Statistics.csv - Summary statistics\n")
  cat("7. Model_Comparison.csv - Model comparison table\n")
  cat("8. Analysis_Results.rds - Complete analysis results\n\n")
  
}, error = function(e) {
  cat("\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n")
  cat("ANALYSIS FAILED WITH ERROR:\n")
  cat(e$message, "\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n\n")
  
  # Try to load pre-existing data files
  cat("Attempting to load pre-existing data files...\n")
  
  if(file.exists("Data_Clean.csv")) {
    cat("Loading Data_Clean.csv...\n")
    Clean <- read.csv("Data_Clean.csv", stringsAsFactors = FALSE)
    
    # Run basic analyses on existing data
    cat("Running basic analyses...\n")
    
    # Create simple correlation plot (compatible way)
    numeric_cols <- sapply(Clean, is.numeric)
    numeric_vars <- Clean[, numeric_cols, drop = FALSE]
    
    if(ncol(numeric_vars) >= 2) {
      png("Figure_Correlation_Basic.png", width = 8, height = 8, units = "in", res = 300)
      chart.Correlation(numeric_vars, histogram = TRUE, pch = 19)
      dev.off()
      cat("Basic correlation plot saved to: Figure_Correlation_Basic.png\n")
    }
    
    # Check for Bergmann's Rule
    if(all(c("Length_mm", "Latitude") %in% names(Clean))) {
      lm_result <- lm(Length_mm ~ Latitude, data = Clean)
      cat("\nBasic Bergmann's Rule analysis:\n")
      cat("Slope:", round(coef(lm_result)[2], 4), "\n")
      cat("R²:", round(summary(lm_result)$r.squared, 4), "\n")
      cat("p-value:", format.pval(summary(lm_result)$coefficients[2, 4], digits = 3), "\n")
    }
  } else {
    cat("No existing data files found.\n")
  }
})

# Save workspace
if(exists("results")) {
  save.image("C_californicus_Analysis_Workspace.RData")
  cat("Workspace saved to: C_californicus_Analysis_Workspace.RData\n")
} else {
  cat("No results to save.\n")
}

# Print session info
cat("\n=== SESSION INFORMATION ===\n")
print(sessionInfo())

```


Prior
Data Clean-up        Post-processing: Raw -> Clean + SoCal       
*Fig 1a-b            Map of C. californicus range                
*Fig 2b              Shell allometry (Fig 2a of shell manual)             
*Fig 3a-d,Supp Fig 1 Size or Shape and Lat / Temp, Bergmann's    
*Fig 4a-f,Supp Fig 2 Size or Shape and Ecoregion + Proxmity      
*Tables 1-2          Statistics breakdown (of all figures)      
*Supp Fig 3          Correlations to human population density    
*Cope's Rule         Size by location ~ Epoch                    
>>>>>>> main

Overview of data and post-processing
```{r}

###### Upload data ######
# Need package for downloading data from Github repo
## install.packages("RCurl")
library("RCurl") 
## packageVersion("RCurl") # v1.98.1.1 

Raw <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable2_Data-Raw.csv")) 

Key <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable1_Data-Key.csv")) 

# Ensure blanks are read as NAs
###### Upload data ######


###### Data structure ######
# Quick view of RawData
str(Raw) 

# Plots of shell morphology measurements as a point of reference for outliers 
morphos = c("Length_mm", "Width_mm", "Aperture_mm", "Thickness_mm", "Globosity_WoverL", "Geometric_Mean")

# Need package for correlation charts
## install.packages("PerformanceAnalytics")
library("PerformanceAnalytics") 
## packageVersion("PerformanceAnalytics") # v2.0.4

# Need package pipes ( %>% or "then" )
## install.packages("tidyverse")
library("tidyverse") 
## packageVersion("tidyverse") # v1.3.0

# Check for outliers by correlation matrix
Raw_CorrMatrix = chart.Correlation(Raw %>% 
                                     select(morphos)) 
Raw_CorrMatrix
###### Data structure ######


###### Clean-up ######
# Organization of locations in dataset
## Re-order counties, North to South
counties = c("Sonoma", "San Mateo", "Monterey", "San Luis Obispo", "Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego", "Baja California")
Raw$County = factor(Raw$County, levels=counties) # order counties
## Order SoCal counties and locales for when subsetting SoCal data
socal = c("Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego")
socal_locations = c("SB", "NLA", "LA", "SLA", "SD", "ANA", "SCZI", "SRI", "SMI", "CAT", "SBI", "SNI", "SCI") # socal location order
la = "Los Angeles"

# Filter outliers (min/max values identified by Raw correlations)
## Subset for including back in as part of Clean / filtered set (below)
No_TA = subset(Raw, is.na(Raw$Thickness_mm)) # 775 no T or A
Raw_iNat = subset(Raw, Raw$Collection=="iNat") # 99 from iNat
## Confirmed to be operator error (measuring or datasheet input)
OutliersRemoved = Raw %>% # 2996 obs 
  filter(Thickness_mm < 2) %>% # low T
  filter(!(Width_mm < 15 & Thickness_mm > 2)) %>% # too thick for W
  filter(Aperture_mm < 6) %>%  # small A
  filter(!(Length_mm > 20 & Aperture_mm <2)) %>% # A too small for L
  rbind(No_TA) %>%  # adds 775 shells back in 
  filter(Width_mm > 2) %>%  # too narrow (note this gets rid of iNat)
  rbind(Raw_iNat) %>% # add iNat back in
  filter(Globosity_WoverL > 0.42 & Globosity_WoverL < 0.71 ) %>% 
  data.frame # ensure it's a data frame
## Remove 25 shells from Pliocene (too old, few)
OutliersRemoved=subset(OutliersRemoved, !Epoch=="Pliocene")
## Filter specimens that may skew in being in the open ocean and/or too north or south
min(OutliersRemoved$Latitude) # 26.43
max(OutliersRemoved$Latitude) # 38.16
### Ensure updated Clean dataset has 3 or more specimens per lot
Clean = OutliersRemoved %>% 
  filter(Latitude > 26.43) %>% 
  filter(Latitude < 38.16) %>% 
  group_by(Lot_Number) %>% 
  mutate(Specimens_per_Lot = length((Lot_Number))) %>%
  ungroup %>% 
  filter(Specimens_per_Lot > 2 | Collection == "iNat") # keep catalogs with > 2 in lot and iNat (1 per 'lot')
## Validation check for NA values for final data set
cat("\n\n",
    sum(is.na(Clean$Globosity_WoverL)), "have no globosity ratio,\n", # 0
    sum(is.na(Clean$Geometric_Mean)), "have no geometric mean,\n", # 99 [iNat]
    sum(is.na(Clean$Width_mm)), "have no width measurement,\n", # 99 [iNat]
    sum(is.na(Clean$Length_mm)), "have no length measurement,\n", # 99 [iNat]
    sum(is.na(Clean$Thickness_mm)), "have no thickness measurement,\n", # 742 = 99 [iNat] + 338 PR + 273 Pleisto
    sum(is.na(Clean$Aperture_mm)), "have no aperture measurement" ) # 742 = 99 [iNat] + 338 PR + 273 Pleisto
## Order the data
Clean %>% 
  group_by(Lot_Number) %>% 
  summarise(N = length(Specimen_Number))
Clean %>% 
  group_by(Epoch) %>% 
  summarise(N = length(Specimen_Number)) 
## Re-do of correlation chart
Clean_CorrMatrix = chart.Correlation(Clean %>% 
                                       select(morphos)) # much improved
Clean_CorrMatrix
dev.copy(png,'Figure2b_Clean.jpg')
dev.off()

# Subset Clean (2908 of 2996 from Raw) to SoCal (2608)
SoCal = Clean %>% 
  filter(County %in% 
           socal) 
SoCal$Location = factor(SoCal$Location, levels=socal_locations) # order 
# Distribution of shells by county *most are in LA (1372)
SoCal %>% 
  group_by(County) %>% 
  summarise(N = length(Epoch)) %>% 
  data.frame 
# Check correlation for SoCal subset
SoCal_CorrMatrix = chart.Correlation(SoCal %>% 
                                       select(morphos)) # Also looks good
SoCal_CorrMatrix
dev.copy(png,'Figure2b_SoCal.jpg')
dev.off()

# Subset 
LA = SoCal %>% 
  filter(County %in% 
           la) 
LA$Location = factor(LA$Location, levels=socal_locations) # order 
# Check correlation for LA subset
LA_CorrMatrix = chart.Correlation(LA %>%
                                    select(morphos)) #
LA_CorrMatrix
dev.copy(png,'Figure2b_LA.jpg')
dev.off()
###### Clean-up ######



###### Save cleaned-up data for downstream visualization and analysis ######
# Data structure for Clean and SoCal subset
sort(unique(Clean$Collection)) # Unique collections: 9
sort(unique(Clean$Lot_Number)) # Unique lots: 382 
sort(unique(Clean$County)) # Unique counties: 9
sort(unique(Clean$Location)) # Unique locations: 19
sort(unique(Clean$Coastal_Proximity)) # Unique location types: 2
sort(unique(Clean$Habitat)) # Unique habitats: 4
sort(unique(Clean$Depth)) # Unique depths: 3

sort(unique(SoCal$Collection)) # Unique collections: 9
sort(unique(SoCal$Lot_Number)) # Unique lots: 339
sort(unique(SoCal$County)) # Unique counties: 4
sort(unique(SoCal$Location)) # Unique locations: 13

write.csv(Clean, file = "SupplementalTable3_Data-Clean.csv")
write.csv(SoCal, file = "SupplementalTable4_Data-SoCal.csv") 

# Validation check 
Clean <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable3_Data-Clean.csv")) 
SoCal <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable4_Data-SoCal.csv")) 


###### Save cleaned-up data for downstream visualization and analysis ######

# Subset data by Epoch to color code points
Anthro = subset(Clean, Epoch=="Anthropocene") # 1492
<<<<<<< HEAD
Holo = subset(Clean, Epoch=="Holocene") # 1138
Pleisto = subset(Clean, Epoch=="Pleistocene") # 305

=======
Holo = subset(Clean, Epoch=="Holocene") # 1106
Pleisto = subset(Clean, Epoch=="Pleistocene") # 305

Anthro_SoCal = subset(SoCal, Epoch=="Anthropocene") # 1366
Holo_SoCal = subset(SoCal, Epoch=="Holocene") # 1020
Pleisto_SoCal = subset(SoCal, Epoch=="Pleistocene") # 211

>>>>>>> main
```

Figure 1a-b: Maps of shell locations (Clean and SoCal Data) 
```{r}

# Need packages for map coordinates
## if(!requireNamespace("devtools")) install.packages("devtools")
## devtools::install_github("dkahle/ggmap")
library("ggmap")
## packageVersion("ggmap") # v3.0.0
## install.packages("rgdal")
library("rgdal")
## packageVersion("rgdal") # v1.4.8
# Need package for plots of maps
## install.packages("ggplot2")
library("ggplot2")
## packageVersion("ggplot2") # v3.2.1
# Need package for gridding maps as one figure
## install.packages("cowplot")
library("cowplot")
## packageVersion("cowplot") # v1.0.0

# This code is personalized by account and only runs if you have a key
## Instructions: https://github.com/dkahle/ggmap
## Visit your personal Google Maps Platform to retrieve key if you have one
## ggmap::register_google(key = " ") 

# Load polygons for ecoregions and call them after downloaded in your folder 
<<<<<<< HEAD
# shape files from http://www.marineregions.org/ (also in repo)
=======
## shape files from http://www.marineregions.org/ (also in repo)
>>>>>>> main
NC_poly=readOGR( "northern_ecoregions" , layer="ecoregions") 
SCB_poly=readOGR( "southern_ecoregions" , layer="ecoregions") 
MT_poly=readOGR( "magdaelena_ecoregions" , layer="ecoregions") 

# Figure 1a
<<<<<<< HEAD
# Parameters for All points
All_bbox <- make_bbox(lat = Latitude, lon = Longitude, data = Clean)
All_bbox
#       left     bottom      right        top 
#   -122.9015   25.8435 -114.7285   38.7465 

All=get_map(location=c(-125,20,-105,40), # adjust to include regions
            zoom = 5, 
            maptype = "terrain") 

# Create box for zoom of SoCal
SoCal_box = data.frame(x=c(-121,-121,-116.5,-116.5,-121), 
       y=c(35,32,32,35,35)) 

# Plot map
=======
## Parameters for All points
All_bbox <- make_bbox(lat = Latitude, lon = Longitude, data = Clean)
All_bbox
##     left        bottom    right       top 
##     -122.786    27.359    -114.734    37.721 
All=get_map(location=c(-125,20,-105,40), # adjust to include regions
            zoom = 5, 
            maptype = "terrain") 
## Create box for zoom of SoCal
SoCal_box = data.frame(x=c(-121,-121,-116.5,-116.5,-121), 
       y=c(35,32,32,35,35)) 
## Plot map
>>>>>>> main
Clean_map = ggmap(All) +
  geom_polygon(data = NC_poly, aes(x = long, y = lat),
               fill = "white", alpha = 0.5) + # white
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) + # blue
  geom_polygon(data = MT_poly, aes(x = long, y = lat),
               fill = "#666666", alpha = 0.5) + # grey
  stat_density_2d(data = Clean, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # color fill, black
  geom_point(data = Anthro, aes(x = Longitude, y = Latitude), 
             color = "#FC4E07", # orange
             alpha = 0.15,
             size = 1) + 
  geom_point(data = Holo, aes(x = Longitude,y = Latitude), 
             color = "#E7B800", # gold
             alpha = 0.15,
             size = 1) + 
  geom_point(data = Pleisto, aes(x = Longitude, y = Latitude), 
             color = "#00AFBB", # blue
             alpha = 0.15,
             size = 1) + 
  geom_point(data = Key, aes(x = Longitude, y = Latitude), # asbsent data
             color = "#FC4E07", # orange
             shape = 4,
<<<<<<< HEAD
             size = 1.5) + # manually added Punta Abreojos white circle
=======
             size = 1.5) + # manually add Punta Abreojos white circle
>>>>>>> main
  geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude") 
Clean_map

# Figure 1b
SoCal_bbox = make_bbox(lat = Latitude, lon = Longitude, data = SoCal)
SoCal_bbox
##    left        bottom    right       top
##    -120.6685   32.4635 -116.9615   34.5865 

<<<<<<< HEAD
SoCal_bbox = make_bbox(lat = Latitude, lon = Longitude, data = Clean_SoCal)
SoCal_bbox
#       left     bottom      right        top 
#      -120.6695   32.4425 -116.9405   34.5875  

=======
>>>>>>> main
SoCal=get_map(location=c(-121,32,-116.5,35), 
            zoom = 7, 
            maptype = "satellite") 

<<<<<<< HEAD
CleanSoCal_map = ggmap(SoCal) +
  stat_density_2d(data = Clean_SoCal, # density
=======
SoCal$Latitude = as.numeric(SoCal$Latitude)
SoCal$Longitude = as.numeric(SoCal$Longitude)

SoCal = data.frame(SoCal)

SoCal_map = ggmap(SoCal) +
  stat_density_2d(data = SoCal, # density
>>>>>>> main
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .35,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + 
  geom_point(data = Anthro_SoCal, aes(x = Longitude, y = Latitude), 
             color = "#FC4E07",
             alpha = 0.15,
             size = 2) + 
  geom_point(data = Holo_SoCal, aes(x = Longitude,y = Latitude), 
             color = "#E7B800", 
             alpha = 0.15,
             size = 2) + 
  geom_point(data = Pleisto_SoCal, aes(x = Longitude, y = Latitude), 
             color = "#00AFBB",
             alpha = 0.15,
             size = 2) + 
    geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude")
SoCal_map


ggsave("Figure1b_Data-Clean-SoCal_Map.png", dpi = 800)


ggsave("Figure1a_Data-Clean_Map.png", dpi = 800)

##### Create Map (Figure 1a-b) #####
#  Need a personalized account with API key (map key) for Google maps
## Instructions: https://github.com/dkahle/ggmap
## ggmap::register_google(key = " ") 
### More helpful info: https://www.littlemissdata.com/blog/maps 

# Figure 1a: Larger map
## Parameters for All points
All_bbox <- make_bbox(lat = Latitude, lon = Longitude, data = df)
All_bbox # Informs us what to set for parameters
##     left        bottom    right       top 
##     -122.786    27.359    -114.734    37.721 
All=get_map(location=c(-129,20,-105,42), # adjust to include regions
            zoom = 5, 
            maptype = "terrain") # can adjust this type as well
## Create box for zoom of SoCal
SoCal_box = data.frame(x=c(-121,-121,-116.5,-116.5,-121), 
       y=c(35,32,32,35,35)) 
## Plot map 
Figure1a = ggmap(All) + # Google map of parameters
  geom_polygon(data = NC_poly, aes(x = long, y = lat), # polygons (shape files)
               fill = "white", alpha = 0.5) + 
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) +
  geom_polygon(data = MT_poly, aes(x = long, y = lat),
               fill = "#666666", alpha = 0.5) + # grey
  stat_density_2d(data = df, # hypothetical population density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250, # can adjust this for resolution
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # fills density layers as black 
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) + # colors by Epoch
  geom_point(data = df, aes(x = Longitude, y = Latitude, color = Epoch), 
             alpha = 0.15,
             size = 1) + # layers points
  geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude") +
  theme(legend.position="none")
Figure1a

# Figure 1b: Zoom-in of SoCal
SoCal=get_map(location=c(-121,32,-116.5,35), # adjust based on SoCal_box
            maptype = "terrain-background") # can adjust this type as well
## Plot map
Figure1b = ggmap(SoCal) +
  geom_polygon(data = NC_poly, aes(x = long, y = lat),
               fill = "white", alpha = 0.5) + # white
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) + # blue
    stat_density_2d(data = df, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # color fill, black
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_point(data = df, aes(x = Longitude, y = Latitude, color = Epoch), 
             alpha = 0.15,
             size = 1) +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
Figure1b # Note the ecoregions cut-out if polygon not fully included

# Grid using cowplot, and save
Figure1 = plot_grid(Figure1a, Figure1b, labels = c("A", "B"), label_size=12)
Figure1   

ggsave("Figure1.png", dpi = 800) 

##### Create Map (Figure 1a-b) #####

```

<<<<<<< HEAD
Q1) Are shell size and shape independent or dependent metrics of each other? Independent
Figure 2b: Allometry - shell shape vs size (Clean_New) 
=======
Figure 2b: Are shell size, mass, or shape independent or dependent metrics of each other?
>>>>>>> main
```{r}

# Filter out set of specimens more in the open ocean and south of the rest + most northern
min(Clean$Latitude) # 26.43
max(Clean$Latitude) # 38.16
Clean_New = Clean %>% 
  filter(Latitude > 26.43) %>% # 2935 - 2902 = 33 obs removed
  filter(Latitude < 38.16) # 2902 - 2897 = 5 obs removed

# install.packages("GGally")
library(GGally)
# packageVersion("GGally") # v1.4.0

# Create a function for including linear regression line
my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    geom_smooth(method=lm, color="black", ...) 
  p
}

# Figure 2b
<<<<<<< HEAD
Fig2b = Clean_New[,c("Epoch", "Length_mm", "Width_mm", "Globosity_WoverL")] 
# Rename columns
names(Fig2b) = c("Epoch", "Length", "Width", "Globosity") 
=======
Fig2b = Clean[,c("Epoch", "Length_mm", "Width_mm", "Globosity_WoverL", "Geometric_Mean")] 
# Rename columns
names(Fig2b) = c("Epoch", "Length (size)", "Width (size)", "Globosity (shape)", "geomM (mass)") 
>>>>>>> main
# Create figure (automatically excludes NAs)
Fig2b_chart = ggpairs(Fig2b, aes(color = Epoch, alpha = 0.85), lower = list(continuous = my_fn), upper = list(continuous = wrap("cor", size = 4.25)))
# Change colors manually with a loop through each plot
for(i in 1:Fig2b_chart$nrow) {
  for(j in 1:Fig2b_chart$ncol){
    Fig2b_chart[i,j] <- Fig2b_chart[i,j] + 
        scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
        scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB"))  
  }
}

Fig2b_chart + 
  theme_classic() # theme for figure
ggsave("Figure2b_Data-Clean-New_Shell-Allometry.png", dpi = 800)
<<<<<<< HEAD

# Statistics
summary(lm(Length ~ Width, data = Fig2b)) # Radj = 0.9534, p < 2.2e-16
summary(lm(Length ~ Globosity, data = Fig2b)) # Radj = 0.08497, p < 2.2e-16
summary(lm(Width ~ Globosity, data = Fig2b)) # Radj = 0.007467, p = 2.854e-06  

# Length = Width, globosity independent of length and width
## shape = globosity
## size = length=width (better to go with width as size over length given lm results)
### Though, barely a difference in the larger scope of Radj (for L ~ g vs W ~ g)
=======

# Statistics
names(Fig2b) = c("Epoch", "Length", "Width", "Globosity", "geomM") 
summary(lm(Length ~ Width, data = Fig2b)) # Radj = 0.9534, p < 2.2e-16
summary(lm(Length ~ Globosity, data = Fig2b)) # Radj = 0.0853, p < 2.2e-16
summary(lm(Width ~ Globosity, data = Fig2b)) # Radj = 0.007822, p = 2.796e-06 
summary(lm(Length ~ geomM, data = Fig2b)) # Radj = 0.9884, p < 2.2e-16
summary(lm(Width ~ geomM, data = Fig2b)) # Radj = 0.988, p < 2.2e-16
summary(lm(Globosity ~ geomM, data = Fig2b)) # Radj = 0.03698, p < 2.2e-16

# Length = Width, dependent on each other and synonymous with size
# Globosity (shape) is an independent metric of size 
## Width slightly more independent (horizontal line to globosity) than Length
### Thus, use Width as proxy for size 
# geomM (mass) is synonymous with size in this instance, and independent of shape
## Do we want to then only use geomM (mass) and globosity (shape)?
### Note Width is still the most horizontal of the 3 (Length, Width, geomM) for Fig 2b ()
>>>>>>> main

```

Q2) Do size/shape vary by epoch in relation to latitude or temperature? Yes.
<<<<<<< HEAD
Figure 3a-d, Supp Figure 1: Comparison of size and shape for Latitude vs SST (Clean_New)
=======
Figure 3a-d, Supp Figure 1: Comparison of size and shape for Latitude vs SST (Clean)
>>>>>>> main
```{r}

# Regression Plot function for y = mx + b
ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}

# Supplemental Figure 1: What is the relationship between latitude and SST?
<<<<<<< HEAD
SuppFig1 = ggplot(data = Clean_New, 
=======
SuppFig1 = ggplot(data = Clean, 
>>>>>>> main
                      aes(x = Latitude, y = Mean_SST,
                      color = Epoch,
                      alpha = 0.4)) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.2, fill="black") + # shade square
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Latitude", y="Mean SST")
SuppFig1 +
  theme_classic() 
ggsave("SuppFig1_Data-Clean-New_Latitude-MeanSST.png", dpi = 800)
# Regression by Epoch
<<<<<<< HEAD
ggplotRegression(lm(Mean_SST ~ Latitude, data = Clean_New))
# ** y = 38.222 - 0.660x, r^2 = 0.157, p = 5.302 x 10^-110
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean_New, Epoch=="Anthropocene")))
# ** y = 43.679 - 0.810x, r^2 = 0.413, p = 1.487 x 10^-174
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean_New, Epoch=="Holocene")))
# ** y = 35.303 - 0.565x, r^2 = 0.259, p = 7.868 x 10^-74
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean_New, Epoch=="Pleistocene")))
# ** y = 32.702 - 0.591, r^2 = 0.457, p = 1.704 x 10^-42

# Figure 3a-d (size is represented as Width) 
Fig3a = ggplot(data = Clean_New, 
                      aes(x = Latitude, y = Width_mm,
=======
ggplotRegression(lm(Mean_SST ~ Latitude, data = Clean))
# ** y = 38.222 - 0.660x, r^2 = 0.157, p = 5.302 x 10^-110
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean, Epoch=="Anthropocene")))
# ** y = 43.679 - 0.810x, r^2 = 0.413, p = 1.487 x 10^-174
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean, Epoch=="Holocene")))
# ** y = 35.303 - 0.565x, r^2 = 0.259, p = 7.868 x 10^-74
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean, Epoch=="Pleistocene")))
# ** y = 32.702 - 0.591, r^2 = 0.457, p = 1.704 x 10^-42

# Figure 3a-d (size is represented as Width) 
Fig3a = ggplot(data = Clean, 
                      aes(x = Latitude, y = Length_mm,
>>>>>>> main
                      color = Epoch,
                      alpha = 0.4)) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.2, fill="black") + # shade square
<<<<<<< HEAD
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Latitude", y="Size (Width mm)")
Fig3a +
  theme_classic() 
ggsave("Fig3a_Data-Clean_New_Latitude-Width.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Latitude ~ Width_mm, data = Clean_New))
# ** y = 32.837 + 0.057x, r^2 = 0.043, p = 1.056 x 10^-28
## Relationships by Epoch are not important based on global interactive model results (Tab 1)

Fig3b = ggplot(data = Clean_New, 
                      aes(x = Mean_SST, y = Width_mm,
                      color = Epoch,
                      alpha = 0.4)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Mean SST", y="Size (Width mm)")
Fig3b +
  theme_classic() 
ggsave("Fig3b_Data-Clean_New_MeanSST-Width.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Width_mm, data = Clean_New))
# y = 16.050 + 0.005x, r^2 = 2.267 x 10^-4, p = 0.545
ggplotRegression(lm(Mean_SST ~ Width_mm, data = subset(Clean_New, Epoch=="Anthropocene")))
# y = 16.663 - 7.955x10^-5x, r^2 = -0.001, p = 0.993
ggplotRegression(lm(Mean_SST ~ Width_mm, data = subset(Clean_New, Epoch=="Holocene")))
# ** y = 17.009 - 0.065x, r^2 = 0.057, p = 7.759 x 10^-26
ggplotRegression(lm(Mean_SST ~ Width_mm, data = subset(Clean_New, Epoch=="Pleistocene")))
# y = 13.191 - 0.021x, r^2 = 0.001, p = 0.255

Fig3c = ggplot(data = Clean_New, 
                      aes(x = Latitude, y = Globosity_WoverL,
                      color = Epoch,
                      alpha = 0.4)) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.2, fill="black") + # shade square
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Latitude", y="Shape (Globosity, W/L)")
Fig3c +
  theme_classic() 
ggsave("Fig3c_Data-Clean_New_Latitude-Globosity.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = Clean_New))
# ** y = 31.951 + 2.645x, r^2 = 0.013, p = 8.232e-10
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean_New, Epoch=="Anthropocene")))
# ** y = 31.235 + 3.662x, r^2 = 0.028, p = 5.554e-11
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean_New, Epoch=="Holocene")))
# y = 33.297 + 0.697x, r^2 = -1.745e-5, p = 0.322
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean_New, Epoch=="Pleistocene")))
# ** y = 36.22 - 5.014x, r^2 = 0.025, p = 0.004

Fig3d = ggplot(data = Clean_New, 
                      aes(x = Mean_SST, y = Globosity_WoverL,
                      color = Epoch,
                      alpha = 0.4)) +
=======
>>>>>>> main
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
<<<<<<< HEAD
  labs(x="Mean SST", y="Shape (Globosity, W/L)")
Fig3d +
  theme_classic() 
ggsave("Fig3d_Data-Clean_New_MeanSST-Globosity.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = Clean_New))
# ** y = 17.072 - 1.619x, r^2 = 0.001, p = 0.024
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean_New, Epoch=="Anthropocene")))
# ** y = 21.187 - 7.901x, r^2 = 0.083, p = 4.988e-30
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean_New, Epoch=="Holocene")))
# ** y = 13.424 + 4.866x, r^2 = 0.035, p = 3.155e-10
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean_New, Epoch=="Pleistocene")))
# ** y = 13.21 - 0.237x, r^2 = -0.003, p = 0.875

ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean_New))) # p = 0.923
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene"))) # p < 0.0001
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")))
# p < 0.0001
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene"))) # p = 0.03
=======
  labs(x="Latitude", y="Size (Length mm)") +
  theme(legend.position = "none") 
# Regression by Epoch
ggplotRegression(lm(Latitude ~ Length_mm, data = Clean))
# ** y = 32.837 + 0.057x, r^2 = 0.043, p = 1.056 x 10^-28
## Relationships by Epoch are not important based on global interactive model results (Tab 1)

Fig3b = ggplot(data = Clean, 
                      aes(x = Mean_SST, y = Length_mm,
                      color = Epoch,
                      alpha = 0.4)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Mean SST", y="") +
  theme(legend.position = "none") 
# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Length_mm, data = Clean))
# y = 16.050 + 0.005x, r^2 = 2.267 x 10^-4, p = 0.545
ggplotRegression(lm(Mean_SST ~ Length_mm, data = subset(Clean, Epoch=="Anthropocene")))
# y = 16.663 - 7.955x10^-5x, r^2 = -0.001, p = 0.993
ggplotRegression(lm(Mean_SST ~ Length_mm, data = subset(Clean, Epoch=="Holocene")))
# ** y = 17.009 - 0.065x, r^2 = 0.057, p = 7.759 x 10^-26
ggplotRegression(lm(Mean_SST ~ Length_mm, data = subset(Clean, Epoch=="Pleistocene")))
# y = 13.191 - 0.021x, r^2 = 0.001, p = 0.255

Figure3 = plot_grid(Fig3a , Fig3b, labels = c("A", "B"), label_size=12)
ggsave("Fig3_Length.png", dpi = 800)








Fig3c = ggplot(data = Clean, 
                      aes(x = Latitude, y = Globosity_WoverL,
                      color = Epoch,
                      alpha = 0.4)) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.2, fill="black") + # shade square
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Latitude", y="Shape (Globosity, W/L)")
Fig3c +
  theme_classic() 
ggsave("Fig3c_Data-Clean_Latitude-Globosity.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = Clean))
# ** y = 31.951 + 2.645x, r^2 = 0.013, p = 8.232e-10
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean, Epoch=="Anthropocene")))
# ** y = 31.235 + 3.662x, r^2 = 0.028, p = 5.554e-11
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean, Epoch=="Holocene")))
# y = 33.297 + 0.697x, r^2 = -1.745e-5, p = 0.322
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean, Epoch=="Pleistocene")))
# ** y = 36.22 - 5.014x, r^2 = 0.025, p = 0.004

Fig3d = ggplot(data = Clean, 
                      aes(x = Mean_SST, y = Globosity_WoverL,
                      color = Epoch,
                      alpha = 0.4)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Mean SST", y="Shape (Globosity, W/L)")
Fig3d +
  theme_classic() 
ggsave("Fig3d_Data-Clean_MeanSST-Globosity.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = Clean))
# ** y = 17.072 - 1.619x, r^2 = 0.001, p = 0.024
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean, Epoch=="Anthropocene")))
# ** y = 21.187 - 7.901x, r^2 = 0.083, p = 4.988e-30
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean, Epoch=="Holocene")))
# ** y = 13.424 + 4.866x, r^2 = 0.035, p = 3.155e-10
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean, Epoch=="Pleistocene")))
# ** y = 13.21 - 0.237x, r^2 = -0.003, p = 0.875




ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean))) # p = 0.923
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean, Epoch=="Anthropocene"))) # p < 0.0001
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean, Epoch=="Holocene")))
# p < 0.0001
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean, Epoch=="Pleistocene"))) # p = 0.03


# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = Clean))
# 
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = subset(Clean, Epoch=="Anthropocene")))
#
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = subset(Clean, Epoch=="Holocene")))
# **
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = subset(Clean, Epoch=="Pleistocene")))
# 
>>>>>>> main

```

Q3) Do size/shape vary in relation to ecoregion and coastal proximity? Yes.
Figure 4a-f, Supp Fig 2: Size vs shape for ecoregion and coastal proximity
(Clean_New)
```{r}

# install.packages("ggplot2")
library(ggplot2)
# packageVersion("ggplot2") # v3.1.0

# Supplemental Figure 2: What is the relationship between ecoregion, coastal proximity, and SST?
min(Clean_New$Mean_SST) # 9.68
max(Clean_New$Mean_SST) # 19.52831

SuppFig2a <- ggplot(subset(Clean_New, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Mean_SST, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(9.5,20)) +
  scale_y_continuous(name = "Mean SST") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(a) Pleistocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
SuppFig2a
ggsave("SuppFig2a_Data-Clean-New_Ecoregion-Proximity-MeanSST.png", dpi = 800)

SuppFig2b <- ggplot(subset(Clean_New, Epoch=="Holocene"), aes(x = Ecoregion, y = Mean_SST, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(9.5,20)) +
  scale_y_continuous(name = "Mean SST") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(b) Holocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
SuppFig2b
ggsave("SuppFig2b_Data-Clean-New_Ecoregion-Proximity-MeanSST.png", dpi = 800)

SuppFig2c <- ggplot(subset(Clean_New, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Mean_SST, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(9.5,20)) +
  scale_y_continuous(name = "Mean SST") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(c) Anthropocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
SuppFig2c
ggsave("SuppFig2c_Data-Clean-New_Ecoregion-Proximity-MeanSST.png", dpi = 800)

# Figure 4a-f
min(Clean_New$Width_mm) # 3.23
max(Clean_New$Width_mm) # 23.65
min(Clean_New$Globosity_WoverL) # 0.43
max(Clean_New$Globosity_WoverL) # 0.7

<<<<<<< HEAD
Fig4a <- ggplot(subset(Clean_New, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Width_mm, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(3,24)) +
  scale_y_continuous(name = "Size (Width mm)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(a) Pleistocene") +
=======








library("ggsignif")
library("ggpubr")

Fig4a <- ggplot(subset(Clean, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Geometric_Mean, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(3,24)) +
  scale_y_continuous(name = "Mass (Geometric Mean)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(a) Pleistocene") +
  geom_signif(comparisons=list("Ecoregion","Coastal_Proximity"), map_signif_level=TRUE) + 
  stat_compare_means(method = "t.test") +
>>>>>>> main
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4a
<<<<<<< HEAD
ggsave("Fig4a_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

Fig4b <- ggplot(subset(Clean_New, Epoch=="Holocene"), aes(x = Ecoregion, y = Width_mm, fill = Coastal_Proximity)) +
=======


Fig4b <- ggplot(subset(Clean, Epoch=="Holocene"), aes(x = Ecoregion, y = Geometric_Mean, fill = Coastal_Proximity)) +
>>>>>>> main
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(3,24)) +
  scale_y_continuous(name = "Size (Width mm)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(b) Holocene") +
<<<<<<< HEAD
=======
  geom_signif(comparisons=list("Ecoregion"), map_signif_level=TRUE) + 
  stat_compare_means(method = "t.test") +
>>>>>>> main
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4b
ggsave("Fig4b_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

<<<<<<< HEAD
Fig4c <- ggplot(subset(Clean_New, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Width_mm, fill = Coastal_Proximity)) +
=======
Fig4c <- ggplot(subset(Clean, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Geometric_Mean, fill = Coastal_Proximity)) +
>>>>>>> main
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(3,24)) +
  scale_y_continuous(name = "Size (Width mm)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(c) Anthropocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4c
ggsave("Fig4c_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

<<<<<<< HEAD
Fig4d <- ggplot(subset(Clean_New, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
=======
Fig4d <- ggplot(subset(Clean, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
>>>>>>> main
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(.4,.72)) +
  scale_y_continuous(name = "Shape (Globosity, W/L)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(d)") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4d
ggsave("Fig4d_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

<<<<<<< HEAD
Fig4e <- ggplot(subset(Clean_New, Epoch=="Holocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
=======
Fig4e <- ggplot(subset(Clean, Epoch=="Holocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
>>>>>>> main
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(.4,.72)) +
  scale_y_continuous(name = "Shape (Globosity, W/L)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(e)") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4e
ggsave("Fig4e_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

<<<<<<< HEAD
Fig4f <- ggplot(subset(Clean_New, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
=======
Fig4f <- ggplot(subset(Clean, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
>>>>>>> main
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(.4,.72)) +
  scale_y_continuous(name = "Shape (Globosity, W/L)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(f)") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4f
ggsave("Fig4f_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)
<<<<<<< HEAD
=======



Fig5a = plot_grid(sys.plot, dia.plot, labels = "a", label_size=12)
Fig5a   

```

Tables 1-2: Statistics, namely for Figs 3-4 (Clean_New)
```{r}

# Table 1 (for Fig 3)
Clean_Numeric <- Clean_New # This is to have lm() Epoch show up as one instead of breakdown
Clean_Numeric$Epoch<-as.numeric(Clean_Numeric$Epoch)

Fig3a_aov <- aov(Width_mm ~ Latitude + Epoch + Latitude*Epoch, data = Clean_New)
Fig3a_GIM <- lm(Width_mm ~ Latitude + Epoch + Latitude*Epoch, data = Clean_Numeric)

Fig3b_aov <- aov(Width_mm ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_New)
Fig3b_GIM <- lm(Width_mm ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_Numeric)

Fig3c_aov <- aov(Globosity_WoverL ~ Latitude + Epoch + Latitude*Epoch, data = Clean_New)
Fig3c_GIM <- lm(Globosity_WoverL ~ Latitude + Epoch + Latitude*Epoch, data = Clean_Numeric)

Fig3d_aov <- aov(Globosity_WoverL ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_New)
Fig3d_GIM <- lm(Globosity_WoverL ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_Numeric)

# For Supp Fig 2
SuppFig2a_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # interaction by ecoregion only

SuppFig2b_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
SuppFig2b_GIM <- lm(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant for all

SuppFig2c_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
SuppFig2c_GIM <- lm(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant for all

# Table 2 (for Fig 4)
Clean_Numeric <- Clean_New # Need to re-do for filter by Epoch
Clean_Numeric$Ecoregion <- as.numeric(Clean_Numeric$Ecoregion)
Clean_Numeric$Coastal_Proximity <- as.numeric(Clean_Numeric$Coastal_Proximity)

Fig4a_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # no significance
Fig4a_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Pleistocene"))

Fig4b_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
Fig4b_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Holocene")) # significant for all

Fig4c_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
Fig4c_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Anthropocene")) # significant for all

Fig4d_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # no significance
Fig4d_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Pleistocene")) # no significance

Fig4e_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
Fig4e_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Holocene")) # significant for all

Fig4f_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
Fig4f_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Anthropocene")) # no significance between ecoregions

# Tables 1, 2
Tab1 <- list(Fig3a_GIM, Fig3b_GIM, Fig3c_GIM, Fig3d_GIM)
Tab2 <- list(Fig4a_GIM, Fig4b_GIM, Fig4c_GIM, Fig4d_GIM, Fig4e_GIM, Fig4f_GIM) 

# Print summaries of each respective table 
for(i in Tab1){
 
 print(summary(i))
  # r <- summary(i)$adj.r.squared
 # print(r)
}

for(i in Tab2){
 
 print(summary(i))
  # r <- summary(i)$adj.r.squared
 # print(r)
}

```

Supplemental Figure 3: Relationship to human population density
```{r}

SuppFig3 = subset(Clean_New, Epoch!="Pleistocene")
SuppFig3 = subset(SuppFig3, Collection!="PR") # Removed personal research (biases SD)
SuppFig3 = subset(SuppFig3, Population_Density > 0) # 1349 observations

# Create a column that converts population density to categories by standard deviation
min(SuppFig3$Population_Density) # 0.00932179
x <- mean(SuppFig3$Population_Density) # 247.2658
SD <- sd(SuppFig3$Population_Density) # 546.268
max(SuppFig3$Population_Density) # 2708.81

SuppFig3$Human_pop[SuppFig3$Population_Density < (617-546)] <- "Low"
SuppFig3$Human_pop[SuppFig3$Population_Density > (617+546)] <- "High"
SuppFig3$Human_pop[SuppFig3$Population_Density > (617-545) & 
                     SuppFig3$Population_Density < (617+545)] <- "Medium"

# Supp Fig 3a-b 
SuppFig3a = ggplot(data = SuppFig3, 
                      aes(x = Decade, y = Width_mm,
                      color = Human_pop,
                      alpha = 0.4)) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Human_pop)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Decade", y="Size (Width mm)")
SuppFig3a +
  theme_classic() 
ggsave("SuppFig3a_Data-Clean_New_Decade-Width-Human.png", dpi = 800)
# Regression by category
ggplotRegression(lm(Width_mm ~ Decade, data = SuppFig3))
# ** y = 114.07 - 0.053x, r^2 = 0.076, p = 2.118e-23
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="Low")))
# ** y = 147.5 - 0.070x, r^2 = 0.162, p = 1.87e-38
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="Medium")))
# ** y = -197.93 + 0.107x, r^2 = 0.059, p = 0.00014
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="High")))
# ** y = 583.62 - 0.291x, r^2 = 0.394, p = 1.309e-10

SuppFig3b = ggplot(data = SuppFig3, 
                      aes(x = Decade, y = Globosity_WoverL,
                      color = Human_pop,
                      alpha = 0.4)) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Human_pop)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Decade", y="Globosity, W/L")
SuppFig3b +
  theme_classic() 
ggsave("SuppFig3b_Data-Clean_New_Decade-Globosity-Human.png", dpi = 800)

ggplotRegression(lm(Globosity_WoverL ~ Decade, data = SuppFig3))
# y = 0.444 + 6.99e-5x, r^2 = 0.002, p = 0.074
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="Low")))
# ** y = 0.019x - 2.88e-4x, r^2 = 0.028, p = 8.486e-8
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="Medium")))
# ** y = 1.858 - 6.49e-4x, r^2 = 0.101, p = 4.326e-7
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="High")))
# y = 0.795 - 1.11e-4x, r^2 = 0.002, p = 0.249

```

Cope's Rule
```{r}

Copes <- Clean_New[c("Location", "Epoch", "Width_mm", "Length_mm", "Globosity_WoverL")]
Copes <- Copes[complete.cases(Copes), ] # removes NAs from all rows to only compare where all measurements are available

All_Locales <- lm(Width_mm ~ Location + Epoch + Location*Epoch , data = Clean_New) # interaction by ecoregion only
summary(All_Locales)
Merged <- aov(Width_mm ~ Epoch , data = Clean_New) # interaction by ecoregion only
summary(Merged) # yes, significant by size

# bar plot
# width, locations merge by epoch
Copes_new <- Copes %>% select("Location", "Epoch", "Width_mm") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Width_mm) + sd(Width_mm), # upper limit for error bar
         lower = mean(Width_mm) - sd(Width_mm)) # lower limit for error bar


Width <- ggplot(Copes_new, aes(x = Epoch, y = Width_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Width of the error bars
                position = position_dodge(.9))
ggsave("Copes_Widthmm_Epoch-All.png", dpi = 800)

# width, locations separated out
Copes_new <- Copes %>% select("Location", "Epoch", "Width_mm") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Width_mm) + sd(Width_mm), # upper limit for error bar
         lower = mean(Width_mm) - sd(Width_mm)) # lower limit for error bar

Width <- ggplot(Copes_new, aes(x = Location, y = Width_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Width of the error bars
                position = position_dodge(.9))
Width
ggsave("Copes_Widthmm_Epoch-by-Location.png", dpi = 800)


# Length
Copes_new <- Copes %>% select("Location", "Epoch", "Length_mm") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Length_mm) + sd(Length_mm), # upper limit for error bar
         lower = mean(Length_mm) - sd(Length_mm)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Epoch, y = Length_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
ggsave("Copes_Lengthmm_Epoch-All.png", dpi = 800)

# Length, locations separated out
Copes_new <- Copes %>% select("Location", "Epoch", "Length_mm") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Length_mm) + sd(Length_mm), # upper limit for error bar
         lower = mean(Length_mm) - sd(Length_mm)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Location, y = Length_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
Length
ggsave("Copes_Lengthmm_Epoch-by-Location.png", dpi = 800)

# Globosity
Copes_new <- Copes %>% select("Location", "Epoch", "Globosity_WoverL") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Globosity_WoverL) + sd(Globosity_WoverL), # upper limit for error bar
         lower = mean(Globosity_WoverL) - sd(Globosity_WoverL)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Epoch, y = Globosity_WoverL, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
ggsave("Copes_Globosity_Epoch-All.png", dpi = 800)

Copes_new <- Copes %>% select("Location", "Epoch", "Globosity_WoverL") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Globosity_WoverL) + sd(Globosity_WoverL), # upper limit for error bar
         lower = mean(Globosity_WoverL) - sd(Globosity_WoverL)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Location, y = Globosity_WoverL, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
Length
ggsave("Copes_Globosity_Epoch-by-Location.png", dpi = 800)
>>>>>>> main

```

Tables 1-2: Statistics, namely for Figs 3-4 (Clean_New)
```{r}

# Table 1 (for Fig 3)
Clean_Numeric <- Clean_New # This is to have lm() Epoch show up as one instead of breakdown
Clean_Numeric$Epoch<-as.numeric(Clean_Numeric$Epoch)

Fig3a_aov <- aov(Width_mm ~ Latitude + Epoch + Latitude*Epoch, data = Clean_New)
Fig3a_GIM <- lm(Width_mm ~ Latitude + Epoch + Latitude*Epoch, data = Clean_Numeric)

Fig3b_aov <- aov(Width_mm ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_New)
Fig3b_GIM <- lm(Width_mm ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_Numeric)

Fig3c_aov <- aov(Globosity_WoverL ~ Latitude + Epoch + Latitude*Epoch, data = Clean_New)
Fig3c_GIM <- lm(Globosity_WoverL ~ Latitude + Epoch + Latitude*Epoch, data = Clean_Numeric)

Fig3d_aov <- aov(Globosity_WoverL ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_New)
Fig3d_GIM <- lm(Globosity_WoverL ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_Numeric)

# For Supp Fig 2
SuppFig2a_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # interaction by ecoregion only

SuppFig2b_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
SuppFig2b_GIM <- lm(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant for all

SuppFig2c_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
SuppFig2c_GIM <- lm(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant for all

# Table 2 (for Fig 4)
Clean_Numeric <- Clean_New # Need to re-do for filter by Epoch
Clean_Numeric$Ecoregion <- as.numeric(Clean_Numeric$Ecoregion)
Clean_Numeric$Coastal_Proximity <- as.numeric(Clean_Numeric$Coastal_Proximity)

Fig4a_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # no significance
Fig4a_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Pleistocene"))

Fig4b_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
Fig4b_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Holocene")) # significant for all

Fig4c_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
Fig4c_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Anthropocene")) # significant for all

Fig4d_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # no significance
Fig4d_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Pleistocene")) # no significance

Fig4e_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
Fig4e_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Holocene")) # significant for all

Fig4f_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
Fig4f_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Anthropocene")) # no significance between ecoregions

# Tables 1, 2
Tab1 <- list(Fig3a_GIM, Fig3b_GIM, Fig3c_GIM, Fig3d_GIM)
Tab2 <- list(Fig4a_GIM, Fig4b_GIM, Fig4c_GIM, Fig4d_GIM, Fig4e_GIM, Fig4f_GIM) 

# Print summaries of each respective table 
for(i in Tab1){
 
 print(summary(i))
  # r <- summary(i)$adj.r.squared
 # print(r)
}

for(i in Tab2){
 
 print(summary(i))
  # r <- summary(i)$adj.r.squared
 # print(r)
}

```

Supplemental Figure 3: Relationship to human population density
```{r}

SuppFig3 = subset(Clean_New, Epoch!="Pleistocene")
SuppFig3 = subset(SuppFig3, Collection!="PR") # Removed personal research (biases SD)
SuppFig3 = subset(SuppFig3, Population_Density > 0) # 1349 observations

# Create a column that converts population density to categories by standard deviation
min(SuppFig3$Population_Density) # 0.00932179
x <- mean(SuppFig3$Population_Density) # 247.2658
SD <- sd(SuppFig3$Population_Density) # 546.268
max(SuppFig3$Population_Density) # 2708.81

SuppFig3$Human_pop[SuppFig3$Population_Density < (617-546)] <- "Low"
SuppFig3$Human_pop[SuppFig3$Population_Density > (617+546)] <- "High"
SuppFig3$Human_pop[SuppFig3$Population_Density > (617-545) & 
                     SuppFig3$Population_Density < (617+545)] <- "Medium"

# Supp Fig 3a-b 
SuppFig3a = ggplot(data = SuppFig3, 
                      aes(x = Decade, y = Width_mm,
                      color = Human_pop,
                      alpha = 0.4)) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Human_pop)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Decade", y="Size (Width mm)")
SuppFig3a +
  theme_classic() 
ggsave("SuppFig3a_Data-Clean_New_Decade-Width-Human.png", dpi = 800)
# Regression by category
ggplotRegression(lm(Width_mm ~ Decade, data = SuppFig3))
# ** y = 114.07 - 0.053x, r^2 = 0.076, p = 2.118e-23
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="Low")))
# ** y = 147.5 - 0.070x, r^2 = 0.162, p = 1.87e-38
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="Medium")))
# ** y = -197.93 + 0.107x, r^2 = 0.059, p = 0.00014
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="High")))
# ** y = 583.62 - 0.291x, r^2 = 0.394, p = 1.309e-10

SuppFig3b = ggplot(data = SuppFig3, 
                      aes(x = Decade, y = Globosity_WoverL,
                      color = Human_pop,
                      alpha = 0.4)) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Human_pop)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Decade", y="Globosity, W/L")
SuppFig3b +
  theme_classic() 
ggsave("SuppFig3b_Data-Clean_New_Decade-Globosity-Human.png", dpi = 800)

ggplotRegression(lm(Globosity_WoverL ~ Decade, data = SuppFig3))
# y = 0.444 + 6.99e-5x, r^2 = 0.002, p = 0.074
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="Low")))
# ** y = 0.019x - 2.88e-4x, r^2 = 0.028, p = 8.486e-8
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="Medium")))
# ** y = 1.858 - 6.49e-4x, r^2 = 0.101, p = 4.326e-7
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="High")))
# y = 0.795 - 1.11e-4x, r^2 = 0.002, p = 0.249

```

Cope's Rule
```{r}

Copes <- Clean_New[c("Location", "Epoch", "Width_mm", "Length_mm", "Globosity_WoverL")]
Copes <- Copes[complete.cases(Copes), ] # removes NAs from all rows to only compare where all measurements are available

All_Locales <- lm(Width_mm ~ Location + Epoch + Location*Epoch , data = Clean_New) # interaction by ecoregion only
summary(All_Locales)
Merged <- aov(Width_mm ~ Epoch , data = Clean_New) # interaction by ecoregion only
summary(Merged) # yes, significant by size

# bar plot
# width, locations merge by epoch
Copes_new <- Copes %>% select("Location", "Epoch", "Width_mm") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Width_mm) + sd(Width_mm), # upper limit for error bar
         lower = mean(Width_mm) - sd(Width_mm)) # lower limit for error bar


Width <- ggplot(Copes_new, aes(x = Epoch, y = Width_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Width of the error bars
                position = position_dodge(.9))
ggsave("Copes_Widthmm_Epoch-All.png", dpi = 800)

# width, locations separated out
Copes_new <- Copes %>% select("Location", "Epoch", "Width_mm") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Width_mm) + sd(Width_mm), # upper limit for error bar
         lower = mean(Width_mm) - sd(Width_mm)) # lower limit for error bar

Width <- ggplot(Copes_new, aes(x = Location, y = Width_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Width of the error bars
                position = position_dodge(.9))
Width
ggsave("Copes_Widthmm_Epoch-by-Location.png", dpi = 800)


# Length
Copes_new <- Copes %>% select("Location", "Epoch", "Length_mm") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Length_mm) + sd(Length_mm), # upper limit for error bar
         lower = mean(Length_mm) - sd(Length_mm)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Epoch, y = Length_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
ggsave("Copes_Lengthmm_Epoch-All.png", dpi = 800)

# Length, locations separated out
Copes_new <- Copes %>% select("Location", "Epoch", "Length_mm") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Length_mm) + sd(Length_mm), # upper limit for error bar
         lower = mean(Length_mm) - sd(Length_mm)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Location, y = Length_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
Length
ggsave("Copes_Lengthmm_Epoch-by-Location.png", dpi = 800)

# Globosity
Copes_new <- Copes %>% select("Location", "Epoch", "Globosity_WoverL") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Globosity_WoverL) + sd(Globosity_WoverL), # upper limit for error bar
         lower = mean(Globosity_WoverL) - sd(Globosity_WoverL)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Epoch, y = Globosity_WoverL, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
ggsave("Copes_Globosity_Epoch-All.png", dpi = 800)

Copes_new <- Copes %>% select("Location", "Epoch", "Globosity_WoverL") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Globosity_WoverL) + sd(Globosity_WoverL), # upper limit for error bar
         lower = mean(Globosity_WoverL) - sd(Globosity_WoverL)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Location, y = Globosity_WoverL, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
Length
ggsave("Copes_Globosity_Epoch-by-Location.png", dpi = 800)

```