---
title: "Bergmann’s Rule across the Pleistocene, Holocene, and Anthropocene in an Eastern Pacific neogastropod, Californiconus californicus (Reeve, 1844) (Gastropoda: Conidae)"
author: "Malloy and Ul-Hasan et al"
date: "Feb 19, 2026"
output:
  pdf_document: default
---

Last edited [Sabah Ul-Hasan]: Feb 19, 2026

https://github.com/MichaelMalloy/C_californicus_Morphology

Contributors of data by order of contribution: MM, SUH, SC, CAO, JKK, WFG
Contributors of code by order of contribution: SUH, LL, MM, SS, CAO

```{r}

# Install packages
## Required packages
required_packages <- c("tidyverse", "PerformanceAnalytics", "GGally", "cowplot",
                       "ggsignif", "ggpubr", "broom", "effectsize", "nlme",
                       "lme4", "mgcv", "MASS", "car", "lmtest", "boot",
                       "lubridate", "gridExtra", "RCurl", "emmeans", "pwr",
                       "sf", "ggspatial", "maps", "mapdata")

## Function to safely install packages
install_if_missing <- function(packages) {
  for(pkg in packages) {
    if(!require(pkg, character.only = TRUE, quietly = TRUE)) {
      message(paste("Installing", pkg, "..."))
      install.packages(pkg, dependencies = TRUE, repos = "https://cloud.r-project.org", quiet = TRUE)
    }
  }
}

## Install 
install_if_missing(required_packages)

## Load
library(tidyverse)
library(PerformanceAnalytics)
library(GGally)
library(cowplot)
library(ggsignif)
library(ggpubr)
library(broom)
library(effectsize)
library(nlme)
library(lme4)
library(mgcv)
library(MASS)
library(car)
library(lmtest)
library(boot)
library(lubridate)
library(gridExtra)
library(RCurl)
library(emmeans)
library(pwr)
library(sf)
library(ggspatial)
library(maps)
library(mapdata)

# Set seed for reproducibility downstream
set.seed(12345)

# Data upload
## Function to load data from your specific GitHub structure
load_data_from_github <- function() {
  
  cat("=== LOADING DATA FROM GITHUB ===\n")
  
  # Define GitHub base URL
  base_url <- "https://raw.githubusercontent.com/sabahzero/C_californicus_Morphology/main/"
  
  # List of files to load
  files <- list(
    raw = "data/SupplementalTable2_Data-Raw.csv",
    clean = "data/SupplementalTable3_Data-Clean.csv",
    socal = "data/SupplementalTable4_Data-SoCal.csv",
    key = "data/SupplementalTable1_Data-Key.csv",
    inat_raw = "data/Data-iNat-Raw.csv",
    raw2 = "data/Data-Raw.csv"
  )
  
  # Function to try loading a file
  try_load <- function(filename, description) {
    url <- paste0(base_url, filename)
    cat(paste0("\nTrying to load ", description, " from:\n", url, "\n"))
    
    tryCatch({
      data <- read.csv(text = getURL(url), stringsAsFactors = FALSE)
      cat(paste0("✓ Successfully loaded ", description, " (", nrow(data), " rows)\n"))
      return(data)
    }, error = function(e) {
      cat(paste0("✗ Failed to load ", description, ": ", e$message, "\n"))
      return(NULL)
    })
  }
  
  # Try to load each file
  data_list <- list()
  
  # Try SupplementalTable2_Data-Raw.csv first (main raw data)
  data_list$raw <- try_load(files$raw, "raw data")
  
  # If that fails, try data/Data-Raw.csv
  if(is.null(data_list$raw)) {
    data_list$raw <- try_load(files$raw2, "alternative raw data")
  }
  
  # Try already cleaned data
  data_list$clean <- try_load(files$clean, "cleaned data")
  data_list$socal <- try_load(files$socal, "SoCal data")
  data_list$key <- try_load(files$key, "data key")
  data_list$inat <- try_load(files$inat_raw, "iNaturalist data")
  
  # Check what we have
  cat("\n=== DATA LOADING SUMMARY ===\n")
  loaded_files <- names(data_list)[!sapply(data_list, is.null)]
  cat("Successfully loaded:", paste(loaded_files, collapse = ", "), "\n")
  
  if(is.null(data_list$raw) && is.null(data_list$clean)) {
    stop("Could not load any data files. Please check your GitHub repository structure.")
  }
  
  return(data_list)
}

# Function to clean data (if raw data is available)
clean_raw_data <- function(raw_data) {
  
  cat("\n=== CLEANING RAW DATA ===\n")
  
  # Make a copy
  data <- raw_data
  
  # Check required columns
  required_cols <- c("Epoch", "Latitude", "Longitude", "Length_mm", "Width_mm")
  missing_cols <- setdiff(required_cols, names(data))
  
  if(length(missing_cols) > 0) {
    cat("Warning: Missing required columns:", paste(missing_cols, collapse = ", "), "\n")
    cat("Attempting to proceed with available columns...\n")
  }
  
  # Clean data step by step
  cat("\n1. Removing Pliocene specimens...\n")
  if("Epoch" %in% names(data)) {
    data <- data %>% filter(Epoch != "Pliocene")
    cat("Rows after removing Pliocene:", nrow(data), "\n")
  }
  
  cat("\n2. Filtering latitude range...\n")
  if(all(c("Latitude") %in% names(data))) {
    data <- data %>% 
      filter(Latitude > 26.43 & Latitude < 38.16)
    cat("Rows after latitude filtering:", nrow(data), "\n")
  }
  
  cat("\n3. Filtering size outliers...\n")
  if("Length_mm" %in% names(data)) {
    data <- data %>% 
      filter(Length_mm >= 5 & Length_mm <= 50)
    cat("Rows after length filtering:", nrow(data), "\n")
  }
  
  if("Width_mm" %in% names(data)) {
    data <- data %>% 
      filter(Width_mm >= 2 & Width_mm <= 30)
    cat("Rows after width filtering:", nrow(data), "\n")
  }
  
  if(all(c("Width_mm", "Length_mm") %in% names(data))) {
    data <- data %>% 
      mutate(Globosity_WoverL = Width_mm / Length_mm) %>% 
      filter(Globosity_WoverL >= 0.3 & Globosity_WoverL <= 0.8)
    cat("Rows after globosity filtering:", nrow(data), "\n")
  }
  
  cat("\n4. Calculating derived metrics...\n")
  if(all(c("Length_mm", "Width_mm") %in% names(data)) && 
     !("Geometric_Mean" %in% names(data))) {
    data <- data %>% 
      mutate(
        Geometric_Mean = sqrt(Length_mm * Width_mm)
      )
  }
  
  if(all(c("Width_mm", "Length_mm") %in% names(data)) && 
     !("Globosity_WoverL" %in% names(data))) {
    data <- data %>% 
      mutate(
        Globosity_WoverL = Width_mm / Length_mm
      )
  }
  
  cat("\n5. Grouping by lot...\n")
  if("Lot_Number" %in% names(data)) {
    data <- data %>% 
      group_by(Lot_Number) %>% 
      mutate(Specimens_per_Lot = n()) %>% 
      ungroup() %>% 
      filter(Specimens_per_Lot > 2 | Collection == "iNat" | is.na(Collection))
    cat("Rows after lot filtering:", nrow(data), "\n")
  }
  
  cat("\n6. Creating factors...\n")
  if("Epoch" %in% names(data)) {
    data <- data %>% 
      mutate(
        Epoch = factor(Epoch, levels = c("Pleistocene", "Holocene", "Anthropocene"))
      )
  }
  
  if("Collection" %in% names(data)) {
    data <- data %>% 
      mutate(
        Data_Type = ifelse(Collection == "iNat", "iNaturalist", "Museum/Field"),
        Data_Type = factor(Data_Type)
      )
  }
  
  if("Ecoregion" %in% names(data)) {
    data$Ecoregion <- factor(data$Ecoregion)
  }
  
  if("Coastal_Proximity" %in% names(data)) {
    data$Coastal_Proximity <- factor(data$Coastal_Proximity)
  }
  
  cat("\n✓ Data cleaning complete\n")
  cat("Final dataset:", nrow(data), "rows\n")
  
  return(data)
}

# Main data loading function
load_and_clean_data <- function() {
  
  # Load data from GitHub
  data_list <- load_data_from_github()
  
  # Determine which data to use
  if(!is.null(data_list$clean)) {
    cat("\nUsing pre-cleaned data from SupplementalTable3_Data-Clean.csv\n")
    Clean <- data_list$clean
    
    # Ensure factors are set correctly
    if("Epoch" %in% names(Clean)) {
      Clean <- Clean %>% 
        mutate(
          Epoch = factor(Epoch, levels = c("Pleistocene", "Holocene", "Anthropocene"))
        )
    }
    
    if("Collection" %in% names(Clean)) {
      Clean <- Clean %>% 
        mutate(
          Data_Type = factor(ifelse(Collection == "iNat", "iNaturalist", "Museum/Field"))
        )
    }
    
  } else if(!is.null(data_list$raw)) {
    cat("\nCleaning raw data from SupplementalTable2_Data-Raw.csv\n")
    Clean <- clean_raw_data(data_list$raw)
  } else {
    stop("No usable data found")
  }
  
  # Check for SoCal data or create it
  if(!is.null(data_list$socal)) {
    cat("\nUsing pre-existing SoCal data\n")
    SoCal <- data_list$socal
  } else {
    cat("\nCreating SoCal subset...\n")
    socal_counties <- c("Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego")
    
    if("County" %in% names(Clean)) {
      SoCal <- Clean %>% filter(County %in% socal_counties)
      cat("SoCal subset created:", nrow(SoCal), "rows\n")
    } else {
      cat("Warning: County column not found. Cannot create SoCal subset.\n")
      SoCal <- NULL
    }
  }
  
  # Convert to sf object for spatial operations (only if coordinates available)
  if(all(c("Longitude", "Latitude") %in% names(Clean))) {
    cat("\nConverting to sf object for spatial analysis...\n")
    
    # Remove rows with missing coordinates
    Clean_sf <- Clean %>% 
      filter(!is.na(Longitude) & !is.na(Latitude))
    
    if(nrow(Clean_sf) > 0) {
      Clean_sf <- st_as_sf(Clean_sf, coords = c("Longitude", "Latitude"), crs = 4326, remove = FALSE)
      cat("Converted", nrow(Clean_sf), "rows to sf object\n")
      
      # Merge back with unconverted rows
      Clean_not_sf <- Clean %>% 
        filter(is.na(Longitude) | is.na(Latitude))
      
      if(nrow(Clean_not_sf) > 0) {
        Clean_not_sf$geometry <- st_sfc(rep(st_point(), nrow(Clean_not_sf)), crs = 4326)
        Clean <- rbind(Clean_sf, Clean_not_sf)
      } else {
        Clean <- Clean_sf
      }
    } else {
      cat("Warning: No rows with valid coordinates. Skipping sf conversion.\n")
    }
  } else {
    cat("\nWarning: Coordinate columns not found. Skipping sf conversion.\n")
  }
  
  # Save cleaned data (remove geometry for CSV)
  if(inherits(Clean, "sf")) {
    Clean_csv <- as.data.frame(Clean)
    # Remove geometry column if it exists
    if("geometry" %in% names(Clean_csv)) {
      Clean_csv$geometry <- NULL
    }
  } else {
    Clean_csv <- Clean
  }
  
  write.csv(Clean_csv, "Data_Clean.csv", row.names = FALSE)
  
  if(!is.null(SoCal)) {
    write.csv(SoCal, "Data_SoCal.csv", row.names = FALSE)
  }
  
  cat("\n✓ Data saved:\n")
  cat("  Data_Clean.csv\n")
  if(!is.null(SoCal)) cat("  Data_SoCal.csv\n")
  
  return(list(Clean = Clean, SoCal = SoCal))
}

# ==============================================================================
# 2. SIMPLIFIED MAP CREATION
# ==============================================================================

# Function to create simple maps without requiring shapefiles
create_simple_maps <- function(data) {
  
  cat("\n=== CREATING SIMPLE MAPS ===\n")
  
  # Remove geometry if it's an sf object
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  # Check for required columns
  if(!all(c("Longitude", "Latitude", "Epoch") %in% names(data_df))) {
    cat("Warning: Missing required columns for mapping\n")
    return(NULL)
  }
  
  # Define epoch colors
  epoch_colors <- c("Pleistocene" = "#00AFBB", 
                    "Holocene" = "#E7B800", 
                    "Anthropocene" = "#FC4E07")
  
  # Get California and Mexico map data
  cat("Loading map data...\n")
  
  # Get world data
  world <- map_data("world")
  
  # Filter for relevant regions
  study_area <- world %>%
    filter((region %in% c("USA", "Canada", "Mexico") & 
             long >= -130 & long <= -110 & lat >= 15 & lat <= 50) |
           (region == "Mexico" & long >= -130 & long <= -110))
  
  # Create base map
  base_map <- ggplot() +
    geom_polygon(data = study_area,
                 aes(x = long, y = lat, group = group),
                 fill = "gray90", color = "gray70", size = 0.2) +
    coord_fixed(xlim = c(-125, -110), ylim = c(22, 42), ratio = 1.3) +
    theme_minimal() +
    theme(
      panel.grid = element_blank(),
      axis.text = element_text(size = 10),
      axis.title = element_text(size = 12, face = "bold")
    )
  
  # Figure 1A: All collection points
  fig1a <- base_map +
    geom_point(data = data_df,
               aes(x = Longitude, y = Latitude, color = Epoch),
               alpha = 0.3, size = 1) +
    scale_color_manual(values = epoch_colors) +
    labs(title = "All Collection Locations",
         x = "Longitude", y = "Latitude",
         color = "Epoch") +
    theme(legend.position = "bottom")
  
  # Figure 1B: With density contours
  fig1b <- base_map +
    stat_density_2d(data = data_df,
                    aes(x = Longitude, y = Latitude, fill = ..level..),
                    geom = "polygon", alpha = 0.3, bins = 20) +
    geom_point(data = data_df,
               aes(x = Longitude, y = Latitude, color = Epoch),
               alpha = 0.3, size = 1) +
    scale_color_manual(values = epoch_colors) +
    scale_fill_gradient(low = "blue", high = "red", guide = "none") +
    labs(title = "Collection Density",
         x = "Longitude", y = "Latitude",
         color = "Epoch") +
    theme(legend.position = "bottom")
  
  # Combine plots
  figure1 <- plot_grid(fig1a, fig1b, 
                       labels = c("A", "B"), 
                       label_size = 14,
                       ncol = 2)
  
  # Save figure
  ggsave("Figure1_Collection_Maps.png", figure1, 
         width = 14, height = 6, dpi = 300, bg = "white")
  
  cat("✓ Figure 1 saved to: Figure1_Collection_Maps.png\n")
  
  return(list(figure1 = figure1))
}

# ==============================================================================
# 3. CORE ANALYSIS FUNCTIONS (COMPATIBLE WITH OLDER R VERSIONS)
# ==============================================================================

# Function to test statistical assumptions
test_assumptions <- function(data) {
  
  cat("\n=== TESTING STATISTICAL ASSUMPTIONS ===\n")
  
  # Remove geometry if data is sf
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  results <- list()
  
  # Check for required variables
  if(!"Length_mm" %in% names(data_df)) {
    cat("Warning: Length_mm not found in data\n")
    return(NULL)
  }
  
  # 1. Normality test
  cat("\n1. Normality test for Length_mm:\n")
  shapiro_test <- shapiro.test(data_df$Length_mm)
  cat("Shapiro-Wilk: W =", round(shapiro_test$statistic, 4),
      "p =", format.pval(shapiro_test$p.value, digits = 3), "\n")
  
  is_normal <- shapiro_test$p.value > 0.05
  cat("Data is", ifelse(is_normal, "NORMALLY", "NOT NORMALLY"), "distributed\n")
  
  # 2. Homogeneity of variance by epoch
  if("Epoch" %in% names(data_df) && length(unique(data_df$Epoch)) > 1) {
    cat("\n2. Homogeneity of variance by epoch:\n")
    levene_test <- leveneTest(Length_mm ~ Epoch, data = data_df)
    cat("Levene's test: F =", round(levene_test$`F value`[1], 4),
        "p =", format.pval(levene_test$`Pr(>F)`[1], digits = 3), "\n")
    
    equal_var <- levene_test$`Pr(>F)`[1] > 0.05
    cat("Variances are", ifelse(equal_var, "EQUAL", "UNEQUAL"), "across epochs\n")
  }
  
  # 3. Create diagnostic plots
  cat("\n3. Creating diagnostic plots...\n")
  
  # Histogram
  hist_plot <- ggplot(data_df, aes(x = Length_mm)) +
    geom_histogram(aes(y = ..density..), bins = 30, fill = "lightblue", color = "black") +
    geom_density(alpha = 0.5, fill = "red") +
    labs(title = "Distribution of Shell Length",
         x = "Length (mm)", y = "Density") +
    theme_minimal()
  
  # QQ plot
  qq_data <- data.frame(
    Theoretical = qqnorm(data_df$Length_mm, plot.it = FALSE)$x,
    Sample = qqnorm(data_df$Length_mm, plot.it = FALSE)$y
  )
  
  qq_plot <- ggplot(qq_data, aes(x = Theoretical, y = Sample)) +
    geom_point(alpha = 0.5) +
    geom_qq_line(aes(sample = Sample), color = "red") +
    labs(title = "Normal Q-Q Plot",
         x = "Theoretical Quantiles", y = "Sample Quantiles") +
    theme_minimal()
  
  # Save diagnostic plots
  diagnostic_plots <- plot_grid(hist_plot, qq_plot, ncol = 2)
  ggsave("Figure_Diagnostics.png", diagnostic_plots, 
         width = 12, height = 5, dpi = 300)
  
  cat("✓ Diagnostic plots saved to: Figure_Diagnostics.png\n")
  
  return(results)
}

# Function to analyze Bergmann's Rule
analyze_bergmanns_rule <- function(data) {
  
  cat("\n=== ANALYZING BERGMANN'S RULE ===\n")
  
  # Remove geometry if data is sf
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  # Check required columns
  required <- c("Length_mm", "Latitude", "Epoch")
  missing <- setdiff(required, names(data_df))
  
  if(length(missing) > 0) {
    cat("Warning: Missing required columns:", paste(missing, collapse = ", "), "\n")
    return(NULL)
  }
  
  results <- list()
  
  # 1. Overall relationship
  cat("\n1. Overall relationship: Length ~ Latitude\n")
  lm_overall <- lm(Length_mm ~ Latitude, data = data_df)
  summary_overall <- summary(lm_overall)
  
  cat("Intercept:", round(coef(lm_overall)[1], 4), "\n")
  cat("Slope:", round(coef(lm_overall)[2], 4), "mm per degree latitude\n")
  cat("R-squared:", round(summary_overall$r.squared, 4), "\n")
  cat("Adjusted R-squared:", round(summary_overall$adj.r.squared, 4), "\n")
  cat("p-value:", format.pval(summary_overall$coefficients[2, 4], digits = 3), "\n")
  
  # 2. By epoch
  cat("\n2. By epoch: Length ~ Latitude * Epoch\n")
  lm_epoch <- lm(Length_mm ~ Latitude * Epoch, data = data_df)
  anova_epoch <- anova(lm_epoch)
  print(anova_epoch)
  
  # 3. Separate models for each epoch
  cat("\n3. Separate models for each epoch:\n")
  epoch_models <- list()
  
  for(epoch in levels(data_df$Epoch)) {
    epoch_data <- data_df %>% filter(Epoch == epoch)
    
    if(nrow(epoch_data) > 10) {  # Need sufficient data
      lm_epoch_specific <- lm(Length_mm ~ Latitude, data = epoch_data)
      epoch_models[[epoch]] <- lm_epoch_specific
      
      cat("\n", epoch, ":\n", sep = "")
      cat("  n =", nrow(epoch_data), "\n")
      cat("  Slope =", round(coef(lm_epoch_specific)[2], 4), "\n")
      cat("  R² =", round(summary(lm_epoch_specific)$r.squared, 4), "\n")
      cat("  p =", format.pval(summary(lm_epoch_specific)$coefficients[2, 4], digits = 3), "\n")
    }
  }
  
  # 4. Create Bergmann's Rule plot
  cat("\n4. Creating Bergmann's Rule visualization...\n")
  
  epoch_colors <- c("Pleistocene" = "#00AFBB", 
                    "Holocene" = "#E7B800", 
                    "Anthropocene" = "#FC4E07")
  
  bergmann_plot <- ggplot(data_df, aes(x = Latitude, y = Length_mm, color = Epoch)) +
    geom_point(alpha = 0.3, size = 1.5) +
    geom_smooth(aes(fill = Epoch), method = "lm", se = TRUE, alpha = 0.2) +
    geom_smooth(method = "lm", color = "black", se = FALSE, size = 0.5) +
    scale_color_manual(values = epoch_colors) +
    scale_fill_manual(values = epoch_colors) +
    labs(
      title = "Bergmann's Rule in Californiconus californicus",
      x = "Latitude (°N)",
      y = "Shell Length (mm)",
      subtitle = paste("Overall slope =", round(coef(lm_overall)[2], 4), 
                      "mm/degree, R² =", round(summary_overall$r.squared, 3))
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      axis.title = element_text(size = 12, face = "bold"),
      legend.position = "bottom",
      legend.title = element_text(face = "bold")
    )
  
  # Save plot
  ggsave("Figure_Bergmanns_Rule.png", bergmann_plot, 
         width = 10, height = 7, dpi = 300, bg = "white")
  
  cat("✓ Bergmann's Rule plot saved to: Figure_Bergmanns_Rule.png\n")
  
  # 5. Calculate effect sizes
  cat("\n5. Calculating effect sizes...\n")
  
  # Split data by median latitude
  lat_median <- median(data_df$Latitude, na.rm = TRUE)
  data_df$Lat_group <- ifelse(data_df$Latitude > lat_median, "High", "Low")
  
  if(length(unique(data_df$Lat_group)) > 1) {
    cohen_d_result <- cohens_d(Length_mm ~ Lat_group, data = data_df)
    cat("Cohen's d for latitude effect:", round(cohen_d_result$Cohens_d, 3), "\n")
    results$cohen_d <- cohen_d_result
  }
  
  # Store results
  results$models <- list(
    overall = lm_overall,
    epoch_interaction = lm_epoch,
    epoch_specific = epoch_models
  )
  
  results$plot <- bergmann_plot
  
  return(results)
}

# Function to create correlation matrix (compatible with older R)
create_correlation_analysis <- function(data) {
  
  cat("\n=== CREATING CORRELATION ANALYSIS ===\n")
  
  # Remove geometry if data is sf
  if(inherits(data, "sf")) {
    data_df <- as.data.frame(data)
    # Remove geometry column if it exists
    if("geometry" %in% names(data_df)) {
      data_df$geometry <- NULL
    }
  } else {
    data_df <- data
  }
  
  # Select numeric variables (compatible way for older R)
  # Identify numeric columns
  numeric_cols <- sapply(data_df, is.numeric)
  
  # Exclude specific columns if they exist
  exclude_cols <- c("Specimen_Number", "Year", "Decade", "Month")
  numeric_cols[names(data_df) %in% exclude_cols] <- FALSE
  
  # Get numeric variables
  numeric_vars <- data_df[, numeric_cols, drop = FALSE]
  
  # Remove columns with too many NAs
  na_counts <- colSums(is.na(numeric_vars))
  numeric_vars <- numeric_vars[, na_counts < nrow(numeric_vars) * 0.5, drop = FALSE]
  
  if(ncol(numeric_vars) < 2) {
    cat("Warning: Not enough numeric variables for correlation analysis\n")
    cat("Numeric variables found:", paste(names(numeric_vars), collapse = ", "), "\n")
    return(NULL)
  }
  
  # Create correlation plot
  cor_plot <- chart.Correlation(numeric_vars, 
                               histogram = TRUE, 
                               pch = 19,
                               main = "Correlation Matrix of Morphological and Environmental Variables")
  
  # Save plot
  png("Figure_Correlation_Matrix.png", width = 10, height = 10, units = "in", res = 300)
  print(cor_plot)
  dev.off()
  
  cat("✓ Correlation matrix saved to: Figure_Correlation_Matrix.png\n")
  
  # Calculate specific correlations of interest
  cat("\nKey correlations:\n")
  
  if(all(c("Length_mm", "Width_mm") %in% names(data_df))) {
    cor_length_width <- cor.test(data_df$Length_mm, data_df$Width_mm, use = "complete.obs")
    cat("Length ~ Width: r =", round(cor_length_width$estimate, 4),
        "p =", format.pval(cor_length_width$p.value, digits = 3), "\n")
  }
  
  if(all(c("Length_mm", "Latitude") %in% names(data_df))) {
    cor_length_lat <- cor.test(data_df$Length_mm, data_df$Latitude, use = "complete.obs")
    cat("Length ~ Latitude: r =", round(cor_length_lat$estimate, 4),
        "p =", format.pval(cor_length_lat$p.value, digits = 3), "\n")
  }
  
  if(all(c("Length_mm", "Globosity_WoverL") %in% names(data_df))) {
    cor_length_shape <- cor.test(data_df$Length_mm, data_df$Globosity_WoverL, use = "complete.obs")
    cat("Length ~ Globosity: r =", round(cor_length_shape$estimate, 4),
        "p =", format.pval(cor_length_shape$p.value, digits = 3), "\n")
  }
  
  return(cor_plot)
}

# ==============================================================================
# 4. MAIN ANALYSIS PIPELINE
# ==============================================================================

run_analysis_pipeline <- function() {
  
  cat(paste0(rep("=", 60), collapse = ""), "\n")
  cat("CALIFORNICONUS CALIFORNICUS MORPHOLOGY ANALYSIS\n")
  cat("Bergmann's Rule across Pleistocene, Holocene, and Anthropocene\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n\n")
  
  # 1. Load and clean data
  data_list <- load_and_clean_data()
  Clean <- data_list$Clean
  
  cat("\n=== DATASET SUMMARY ===\n")
  cat("Total specimens:", nrow(Clean), "\n")
  
  if("Epoch" %in% names(Clean)) {
    cat("\nBy epoch:\n")
    epoch_table <- table(Clean$Epoch)
    print(epoch_table)
  }
  
  if("Data_Type" %in% names(Clean)) {
    cat("\nBy data type:\n")
    data_type_table <- table(Clean$Data_Type)
    print(data_type_table)
  }
  
  # 2. Create maps
  cat("\n=== CREATING VISUALIZATIONS ===\n")
  maps <- create_simple_maps(Clean)
  
  # 3. Test assumptions
  assumptions <- test_assumptions(Clean)
  
  # 4. Correlation analysis
  cor_plot <- create_correlation_analysis(Clean)
  
  # 5. Bergmann's Rule analysis
  bergmann_results <- analyze_bergmanns_rule(Clean)
  
  # 6. Generate summary statistics
  cat("\n=== GENERATING SUMMARY STATISTICS ===\n")
  
  if(inherits(Clean, "sf")) {
    Clean_df <- as.data.frame(Clean)
    # Remove geometry column if it exists
    if("geometry" %in% names(Clean_df)) {
      Clean_df$geometry <- NULL
    }
  } else {
    Clean_df <- Clean
  }
  
  # Create summary by epoch
  if("Epoch" %in% names(Clean_df) && "Length_mm" %in% names(Clean_df)) {
    summary_stats <- data.frame()
    
    for(epoch in levels(Clean_df$Epoch)) {
      epoch_data <- Clean_df %>% filter(Epoch == epoch)
      
      if(nrow(epoch_data) > 0) {
        epoch_summary <- data.frame(
          Epoch = epoch,
          N = nrow(epoch_data),
          Mean_Length = round(mean(epoch_data$Length_mm, na.rm = TRUE), 2),
          SD_Length = round(sd(epoch_data$Length_mm, na.rm = TRUE), 2),
          Min_Length = round(min(epoch_data$Length_mm, na.rm = TRUE), 2),
          Max_Length = round(max(epoch_data$Length_mm, na.rm = TRUE), 2),
          stringsAsFactors = FALSE
        )
        summary_stats <- rbind(summary_stats, epoch_summary)
      }
    }
    
    cat("\nSummary statistics by epoch:\n")
    print(summary_stats)
    
    write.csv(summary_stats, "Summary_Statistics.csv", row.names = FALSE)
    cat("\n✓ Summary statistics saved to: Summary_Statistics.csv\n")
  }
  
  # 7. Create model comparison table
  cat("\n=== CREATING MODEL COMPARISON ===\n")
  
  if(!is.null(bergmann_results)) {
    
    # Get sample sizes
    n_pleistocene <- sum(Clean_df$Epoch == "Pleistocene", na.rm = TRUE)
    n_holocene <- sum(Clean_df$Epoch == "Holocene", na.rm = TRUE)
    n_anthropocene <- sum(Clean_df$Epoch == "Anthropocene", na.rm = TRUE)
    
    model_comparison <- data.frame(
      Model = c("Length ~ Latitude (Overall)", 
                "Length ~ Latitude * Epoch",
                "Length ~ Latitude (Pleistocene)",
                "Length ~ Latitude (Holocene)",
                "Length ~ Latitude (Anthropocene)"),
      n = c(
        nrow(Clean_df),
        nrow(Clean_df),
        n_pleistocene,
        n_holocene,
        n_anthropocene
      ),
      Slope = c(
        round(coef(bergmann_results$models$overall)[2], 4),
        NA,
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(coef(bergmann_results$models$epoch_specific$Pleistocene)[2], 4) else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(coef(bergmann_results$models$epoch_specific$Holocene)[2], 4) else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(coef(bergmann_results$models$epoch_specific$Anthropocene)[2], 4) else NA
      ),
      R_squared = c(
        round(summary(bergmann_results$models$overall)$r.squared, 4),
        round(summary(bergmann_results$models$epoch_interaction)$r.squared, 4),
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(summary(bergmann_results$models$epoch_specific$Pleistocene)$r.squared, 4) else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(summary(bergmann_results$models$epoch_specific$Holocene)$r.squared, 4) else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          round(summary(bergmann_results$models$epoch_specific$Anthropocene)$r.squared, 4) else NA
      ),
      P_value = c(
        format.pval(summary(bergmann_results$models$overall)$coefficients[2, 4], digits = 3),
        if(!is.null(anova(bergmann_results$models$epoch_interaction))) 
          format.pval(anova(bergmann_results$models$epoch_interaction)$`Pr(>F)`[1], digits = 3) else NA,
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          format.pval(summary(bergmann_results$models$epoch_specific$Pleistocene)$coefficients[2, 4], digits = 3) else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          format.pval(summary(bergmann_results$models$epoch_specific$Holocene)$coefficients[2, 4], digits = 3) else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          format.pval(summary(bergmann_results$models$epoch_specific$Anthropocene)$coefficients[2, 4], digits = 3) else NA
      ),
      Supports_Bergmann = c(
        ifelse(summary(bergmann_results$models$overall)$coefficients[2, 4] < 0.05, "Yes", "No"),
        if(!is.null(anova(bergmann_results$models$epoch_interaction))) 
          ifelse(anova(bergmann_results$models$epoch_interaction)$`Pr(>F)`[3] < 0.05, "Varies", "Consistent") else NA,
        if("Pleistocene" %in% names(bergmann_results$models$epoch_specific)) 
          ifelse(summary(bergmann_results$models$epoch_specific$Pleistocene)$coefficients[2, 4] < 0.05, "Yes", "No") else NA,
        if("Holocene" %in% names(bergmann_results$models$epoch_specific)) 
          ifelse(summary(bergmann_results$models$epoch_specific$Holocene)$coefficients[2, 4] < 0.05, "Yes", "No") else NA,
        if("Anthropocene" %in% names(bergmann_results$models$epoch_specific)) 
          ifelse(summary(bergmann_results$models$epoch_specific$Anthropocene)$coefficients[2, 4] < 0.05, "Yes", "No") else NA
      )
    )
    
    write.csv(model_comparison, "Model_Comparison.csv", row.names = FALSE)
    cat("\n✓ Model comparison saved to: Model_Comparison.csv\n")
    
    # Print key findings
    cat("\n=== KEY FINDINGS ===\n")
    
    overall_slope <- coef(bergmann_results$models$overall)[2]
    overall_p <- summary(bergmann_results$models$overall)$coefficients[2, 4]
    
    if(overall_p < 0.05) {
      cat(paste0("\n✓ BERGMANN'S RULE SUPPORTED: Shell length increases with latitude\n",
                 "  Slope: ", round(overall_slope, 4), " mm per degree latitude\n",
                 "  R²: ", round(summary(bergmann_results$models$overall)$r.squared, 3), "\n",
                 "  p-value: ", format.pval(overall_p, digits = 3), "\n"))
    } else {
      cat("\n✗ BERGMANN'S RULE NOT SUPPORTED: No significant relationship with latitude\n")
    }
    
    # Check epoch differences
    if(!is.null(anova(bergmann_results$models$epoch_interaction))) {
      epoch_p <- anova(bergmann_results$models$epoch_interaction)$`Pr(>F)`[3]
      if(!is.na(epoch_p) && epoch_p < 0.05) {
        cat("\n✓ EPOCH EFFECT: Relationship varies significantly between epochs\n")
      } else if(!is.na(epoch_p)) {
        cat("\n✗ NO EPOCH EFFECT: Relationship consistent across epochs\n")
      }
    }
  }
  
  # Return results
  results <- list(
    data = Clean,
    assumptions = assumptions,
    bergmann = bergmann_results,
    maps = maps,
    correlation = cor_plot
  )
  
  # Save results
  saveRDS(results, "Analysis_Results.rds")
  cat("\n✓ Complete analysis results saved to: Analysis_Results.rds\n")
  
  return(results)
}

# ==============================================================================
# 5. EXECUTE ANALYSIS
# ==============================================================================

# Run the analysis
cat("\n")
cat(paste0(rep("=", 60), collapse = ""), "\n")
cat("STARTING ANALYSIS\n")
cat(paste0(rep("=", 60), collapse = ""), "\n\n")

tryCatch({
  results <- run_analysis_pipeline()
  
  cat("\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n")
  cat("ANALYSIS COMPLETE!\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n\n")
  
  cat("Output files generated:\n")
  cat("1. Data_Clean.csv - Cleaned dataset\n")
  cat("2. Figure1_Collection_Maps.png - Collection location maps\n")
  cat("3. Figure_Diagnostics.png - Diagnostic plots\n")
  cat("4. Figure_Correlation_Matrix.png - Correlation matrix\n")
  cat("5. Figure_Bergmanns_Rule.png - Bergmann's Rule visualization\n")
  cat("6. Summary_Statistics.csv - Summary statistics\n")
  cat("7. Model_Comparison.csv - Model comparison table\n")
  cat("8. Analysis_Results.rds - Complete analysis results\n\n")
  
}, error = function(e) {
  cat("\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n")
  cat("ANALYSIS FAILED WITH ERROR:\n")
  cat(e$message, "\n")
  cat(paste0(rep("=", 60), collapse = ""), "\n\n")
  
  # Try to load pre-existing data files
  cat("Attempting to load pre-existing data files...\n")
  
  if(file.exists("Data_Clean.csv")) {
    cat("Loading Data_Clean.csv...\n")
    Clean <- read.csv("Data_Clean.csv", stringsAsFactors = FALSE)
    
    # Run basic analyses on existing data
    cat("Running basic analyses...\n")
    
    # Create simple correlation plot (compatible way)
    numeric_cols <- sapply(Clean, is.numeric)
    numeric_vars <- Clean[, numeric_cols, drop = FALSE]
    
    if(ncol(numeric_vars) >= 2) {
      png("Figure_Correlation_Basic.png", width = 8, height = 8, units = "in", res = 300)
      chart.Correlation(numeric_vars, histogram = TRUE, pch = 19)
      dev.off()
      cat("Basic correlation plot saved to: Figure_Correlation_Basic.png\n")
    }
    
    # Check for Bergmann's Rule
    if(all(c("Length_mm", "Latitude") %in% names(Clean))) {
      lm_result <- lm(Length_mm ~ Latitude, data = Clean)
      cat("\nBasic Bergmann's Rule analysis:\n")
      cat("Slope:", round(coef(lm_result)[2], 4), "\n")
      cat("R²:", round(summary(lm_result)$r.squared, 4), "\n")
      cat("p-value:", format.pval(summary(lm_result)$coefficients[2, 4], digits = 3), "\n")
    }
  } else {
    cat("No existing data files found.\n")
  }
})

# Save workspace
if(exists("results")) {
  save.image("C_californicus_Analysis_Workspace.RData")
  cat("Workspace saved to: C_californicus_Analysis_Workspace.RData\n")
} else {
  cat("No results to save.\n")
}

# Print session info
cat("\n=== SESSION INFORMATION ===\n")
print(sessionInfo())

```

Jan 14, 2026 
Data integration and clean-up
```{r}
# Data Integration Script for C. californicus Morphology Study
# Author: Based on original script by Malloy and Ul-Hasan et al.
# Date: 2026-01-15
# Purpose: Integrate SupplementalTable2_Data-iNat-Raw.csv into existing dataset

# install.packages("tidyverse")
# install.packages("RCurl")

library(tidyverse)
library(RCurl)

# ---------------------------
# 1. LOAD EXISTING DATA
# ---------------------------

# Load original raw data from GitHub
Raw_original <- read.csv(text = getURL("https://raw.githubusercontent.com/sabahzero/C_californicus_Morphology/main/data/Data-iNat-Raw"))

# Load the Data Key for reference
Key <- read.csv(text = getURL("https://raw.githubusercontent.com/sabahzero/C_californicus_Morphology/main/SupplementalTable1_Data-Key.csv"))

# ---------------------------
# 2. LOAD AND PREPARE NEW iNat DATA
# ---------------------------

# Assuming SupplementalTable2_Data-iNat-Raw.csv is in your working directory
# If not, adjust the path accordingly
iNat_new <- read.csv("https://raw.githubusercontent.com/sabahzero/C_californicus_Morphology/main/data/Data-Raw.csv", stringsAsFactors = FALSE)

# Clean the column names to match the original dataset
names(iNat_new) <- gsub("observation_", "", names(iNat_new))
names(iNat_new) <- gsub("latidude", "latitude", names(iNat_new))

# Create a mapping function for iNaturalist data transformation
transform_iNat_data <- function(iNat_df) {
  
  # Initialize an empty dataframe with the structure of Raw_original
  # but ensure all columns have the correct data types
  transformed <- data.frame(
    matrix(ncol = ncol(Raw_original), nrow = nrow(iNat_df))
  )
  colnames(transformed) <- colnames(Raw_original)
  
  # First, convert all columns to character to avoid type mismatches
  transformed[] <- lapply(transformed, as.character)
  
  # Now fill in the transformed dataframe
  # Convert numeric IDs to sequential numbers
  transformed$Specimen_Number <- as.character(
    max(as.numeric(Raw_original$Specimen_Number), na.rm = TRUE) + 
      seq_len(nrow(iNat_df))
  )
  
  transformed$Collection <- "iNat"
  transformed$Lot_Number <- paste0("iNat_", iNat_df$id)
  transformed$Collector <- iNat_df$login
  
  # Ensure Latitude and Longitude are numeric
  transformed$Latitude <- as.character(as.numeric(iNat_df$latitude))
  transformed$Longitude <- as.character(as.numeric(iNat_df$longitude))
  
  # Convert length and width (adjust conversion factor as needed)
  # First check if the columns exist and have data
  if("length" %in% names(iNat_df) && "width" %in% names(iNat_df)) {
    # Convert to numeric, handling empty strings
    length_vals <- as.numeric(iNat_df$length)
    width_vals <- as.numeric(iNat_df$width)
    
    # Apply conversion (adjust /10 as needed based on your data)
    transformed$Length_mm <- as.character(ifelse(
      is.na(length_vals), 
      NA, 
      length_vals / 10  # Example conversion - ADJUST AS NEEDED
    ))
    
    transformed$Width_mm <- as.character(ifelse(
      is.na(width_vals), 
      NA, 
      width_vals / 10    # Example conversion - ADJUST AS NEEDED
    ))
    
    # Calculate derived metrics where we have both measurements
    has_both <- !is.na(length_vals) & !is.na(width_vals)
    if(any(has_both)) {
      length_num <- as.numeric(transformed$Length_mm)
      width_num <- as.numeric(transformed$Width_mm)
      
      transformed$Geometric_Mean <- as.character(
        sqrt(length_num * width_num)
      )
      
      transformed$Globosity_WoverL <- as.character(
        width_num / length_num
      )
    }
  }
  
  # Set epoch to Anthropocene for iNaturalist observations
  transformed$Epoch <- "Anthropocene"
  
  # Parse observation date
  # Handle different date formats
  obs_date <- tryCatch({
    as.Date(iNat_df$date, format = "%m/%d/%y")
  }, error = function(e) {
    # Try alternative format
    as.Date(iNat_df$date, format = "%Y-%m-%d")
  })
  
  transformed$Year <- as.character(as.numeric(format(obs_date, "%Y")))
  transformed$Decade <- as.character(floor(as.numeric(transformed$Year) / 10) * 10)
  transformed$Month <- as.character(as.numeric(format(obs_date, "%m")))
  
  # Assign season based on month
  month_num <- as.numeric(transformed$Month)
  transformed$Season <- case_when(
    month_num %in% c(12, 1, 2) ~ "Winter",
    month_num %in% c(3, 4, 5) ~ "Spring",
    month_num %in% c(6, 7, 8) ~ "Summer",
    month_num %in% c(9, 10, 11) ~ "Fall",
    TRUE ~ NA_character_
  )
  
  # Determine county based on latitude
  lat_num <- as.numeric(transformed$Latitude)
  transformed$County <- case_when(
    lat_num >= 38.0 ~ "Sonoma",
    lat_num >= 37.5 ~ "San Mateo",
    lat_num >= 36.0 ~ "Monterey",
    lat_num >= 35.5 ~ "San Luis Obispo",
    lat_num >= 34.5 ~ "Santa Barbara",
    lat_num >= 34.0 ~ "Ventura",
    lat_num >= 33.5 ~ "Los Angeles",
    lat_num >= 33.0 ~ "Orange",
    lat_num >= 32.5 ~ "San Diego",
    TRUE ~ "Baja California"
  )
  
  # Determine ecoregion based on location
  transformed$Ecoregion <- ifelse(
    lat_num > 34.5, 
    "Northern California", 
    "Southern California Bight"
  )
  
  # Set coastal proximity
  transformed$Coastal_Proximity <- "Coastal"
  
  # Create Location code based on County
  transformed$Location <- case_when(
    transformed$County == "San Diego" ~ "SD",
    transformed$County == "Los Angeles" ~ "LA",
    transformed$County == "Orange" ~ "ORA",
    transformed$County == "Ventura" ~ "VEN",
    transformed$County == "Santa Barbara" ~ "SB",
    TRUE ~ substr(gsub(" ", "", transformed$County), 1, 3)
  )
  
  # Set all other columns to NA
  other_cols <- setdiff(
    colnames(transformed),
    c("Specimen_Number", "Collection", "Lot_Number", "Collector", 
      "Latitude", "Longitude", "Length_mm", "Width_mm", 
      "Geometric_Mean", "Globosity_WoverL", "Epoch", "Year", 
      "Decade", "Month", "Season", "County", "Ecoregion", 
      "Coastal_Proximity", "Location")
  )
  
  for(col in other_cols) {
    transformed[[col]] <- NA_character_
  }
  
  return(transformed)
}

# Transform the new iNaturalist data
iNat_transformed <- transform_iNat_data(iNat_new)

# ---------------------------
# 3. MERGE DATASETS WITH TYPE HANDLING
# ---------------------------

# First, ensure all columns in Raw_original are character type for safe binding
Raw_original_chr <- Raw_original
Raw_original_chr[] <- lapply(Raw_original_chr, as.character)

# Now bind the rows
Raw_combined <- bind_rows(Raw_original_chr, iNat_transformed)

# Convert back to appropriate data types
Raw_combined <- Raw_combined %>%
  mutate(
    Specimen_Number = as.numeric(Specimen_Number),
    Latitude = as.numeric(Latitude),
    Longitude = as.numeric(Longitude),
    Length_mm = as.numeric(Length_mm),
    Width_mm = as.numeric(Width_mm),
    Geometric_Mean = as.numeric(Geometric_Mean),
    Globosity_WoverL = as.numeric(Globosity_WoverL),
    Year = as.numeric(Year),
    Decade = as.numeric(Decade),
    Month = as.numeric(Month),
    Mean_SST = as.numeric(Mean_SST),
    Mean_SST_Sdev = as.numeric(Mean_SST_Sdev),
    Aperture_mm = as.numeric(Aperture_mm),
    Thickness_mm = as.numeric(Thickness_mm)
  )

# Reset Specimen_Number to be sequential
Raw_combined$Specimen_Number <- seq_len(nrow(Raw_combined))

# ---------------------------
# 4. APPLY ORIGINAL CLEANING PIPELINE
# ---------------------------

# Replicate the cleaning steps from your original script
Clean <- Raw_combined %>%
  # Filter outliers (replicating original logic)
  # Note: using ifelse to handle NAs in Thickness_mm and Aperture_mm
  filter(ifelse(is.na(Thickness_mm), TRUE, Thickness_mm < 2)) %>%
  filter(!(Width_mm < 15 & !is.na(Thickness_mm) & Thickness_mm > 2)) %>%
  filter(ifelse(is.na(Aperture_mm), TRUE, Aperture_mm < 6)) %>%
  filter(!(Length_mm > 20 & !is.na(Aperture_mm) & Aperture_mm < 2)) %>%
  filter(Width_mm > 2) %>%
  # Handle NAs in Globosity_WoverL
  filter(
    ifelse(
      is.na(Globosity_WoverL), 
      TRUE, 
      Globosity_WoverL > 0.42 & Globosity_WoverL < 0.71
    )
  ) %>%
  # Remove Pliocene specimens
  filter(!Epoch == "Pliocene") %>%
  # Filter latitude range
  filter(Latitude > 26.43 & Latitude < 38.16) %>%
  # Group by lot and count specimens
  group_by(Lot_Number) %>%
  mutate(Specimens_per_Lot = n()) %>%
  ungroup() %>%
  # Keep catalogs with >2 in lot and iNat (1 per 'lot')
  filter(Specimens_per_Lot > 2 | Collection == "iNat")

# ---------------------------
# 5. CREATE SoCal SUBSET
# ---------------------------

# Define SoCal counties (from original script)
socal_counties <- c("Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego")

SoCal <- Clean %>%
  filter(County %in% socal_counties)

# Order the locations
socal_locations <- c("SB", "NLA", "LA", "SLA", "SD", "ANA", "SCZI", "SRI", "SMI", "CAT", "SBI", "SNI", "SCI")
SoCal$Location <- factor(SoCal$Location, levels = socal_locations)

# ---------------------------
# 6. SAVE THE INTEGRATED DATASETS
# ---------------------------

# Save the integrated datasets
write.csv(Raw_combined, "SupplementalTable2_Data-Raw-Integrated.csv", row.names = FALSE)
write.csv(Clean, "SupplementalTable3_Data-Clean-Integrated.csv", row.names = FALSE)
write.csv(SoCal, "SupplementalTable4_Data-SoCal-Integrated.csv", row.names = FALSE)

# ---------------------------
# 7. VALIDATION AND SUMMARY
# ---------------------------

cat("Data Integration Complete!\n")
cat("=================================\n")
cat("Original raw data rows:", nrow(Raw_original), "\n")
cat("New iNaturalist rows:", nrow(iNat_new), "\n")
cat("Combined raw data rows:", nrow(Raw_combined), "\n")
cat("Cleaned data rows:", nrow(Clean), "\n")
cat("SoCal subset rows:", nrow(SoCal), "\n")
cat("\nCollection breakdown in Clean dataset:\n")
print(table(Clean$Collection))

cat("\niNaturalist observations by County:\n")
iNat_by_county <- Clean %>%
  filter(Collection == "iNat") %>%
  count(County)
print(iNat_by_county)

# ---------------------------
# 8. QUICK DATA QUALITY CHECKS
# ---------------------------

cat("\n=== Data Quality Checks ===\n")

# Check for NAs in key columns
cat("\nMissing values in key columns:\n")
key_cols <- c("Length_mm", "Width_mm", "Latitude", "Longitude", "Epoch")
missing_counts <- sapply(key_cols, function(col) {
  sum(is.na(Clean[[col]]))
})
print(missing_counts)

# Check measurement ranges
cat("\nMeasurement ranges:\n")
if(!all(is.na(Clean$Length_mm))) {
  cat("Length (mm):", range(Clean$Length_mm, na.rm = TRUE), "\n")
}
if(!all(is.na(Clean$Width_mm))) {
  cat("Width (mm):", range(Clean$Width_mm, na.rm = TRUE), "\n")
}
if(!all(is.na(Clean$Globosity_WoverL))) {
  cat("Globosity (W/L):", range(Clean$Globosity_WoverL, na.rm = TRUE), "\n")
}

cat("\nTo use the integrated data in your RMarkdown script, you can either:\n")
cat("1. Use the local files created by this script, or\n")
cat("2. Upload them to GitHub and update the URLs in your script.\n")
```

Statistical analysis of data distribution
```{r}
# Analysis of Clean Dataset: Museum/Field vs iNaturalist Data

# install.packages("broom")
# install.packages("effectsize")
# install.packages("pwr")
# install.packages("lubridate")
# install.packages("gridExtra")

library(tidyverse)
library(broom)
library(effectsize)
library(pwr)
library(lubridate)
library(gridExtra)

# Load the Clean dataset
Clean <- read.csv("https://raw.githubusercontent.com/sabahzero/C_californicus_Morphology/main/SupplementalTable3_Data-Clean.csv", stringsAsFactors = FALSE)

# ------------------------------------------------------------
# 1. DATA PREPARATION AND OVERVIEW
# ------------------------------------------------------------

cat("=== DATA OVERVIEW ===\n")
cat("Total specimens in Clean dataset:", nrow(Clean), "\n")
cat("Unique collections:", paste(unique(Clean$Collection), collapse=", "), "\n\n")

# Create data groups
Clean <- Clean %>%
  mutate(
    Data_Type = ifelse(Collection == "iNat", "iNaturalist", "Museum/Field"),
    # Ensure numeric types
    Length_mm = as.numeric(Length_mm),
    Width_mm = as.numeric(Width_mm),
    Geometric_Mean = as.numeric(Geometric_Mean),
    Globosity_WoverL = as.numeric(Globosity_WoverL),
    Year = as.numeric(Year),
    Decade = as.numeric(Decade)
  )

# Summary by data type
data_type_summary <- Clean %>%
  group_by(Data_Type) %>%
  summarise(
    N = n(),
    Percent = round(n()/nrow(Clean)*100, 1),
    Mean_Length = mean(Length_mm, na.rm = TRUE),
    SD_Length = sd(Length_mm, na.rm = TRUE),
    Mean_Width = mean(Width_mm, na.rm = TRUE),
    SD_Width = sd(Width_mm, na.rm = TRUE),
    Mean_Globosity = mean(Globosity_WoverL, na.rm = TRUE),
    SD_Globosity = sd(Globosity_WoverL, na.rm = TRUE),
    .groups = 'drop'
  )

cat("Data Type Summary:\n")
print(data_type_summary)

# ------------------------------------------------------------
# 2. CHRONOLOGICAL ANALYSIS
# ------------------------------------------------------------

cat("\n=== CHRONOLOGICAL ANALYSIS ===\n")

# Check year ranges
year_summary <- Clean %>%
  filter(!is.na(Year)) %>%
  group_by(Data_Type) %>%
  summarise(
    Min_Year = min(Year, na.rm = TRUE),
    Max_Year = max(Year, na.rm = TRUE),
    Year_Range = max(Year, na.rm = TRUE) - min(Year, na.rm = TRUE),
    Median_Year = median(Year, na.rm = TRUE),
    .groups = 'drop'
  )

cat("\nYear ranges by data type:\n")
print(year_summary)

# Visualize temporal distribution
p1 <- Clean %>%
  filter(!is.na(Year)) %>%
  ggplot(aes(x = Year, fill = Data_Type)) +
  geom_histogram(binwidth = 5, alpha = 0.7, position = "identity") +
  labs(title = "Temporal Distribution of Specimens",
       x = "Year", y = "Count") +
  theme_minimal() +
  theme(legend.position = "bottom")

p2 <- Clean %>%
  filter(!is.na(Year) & Year >= 2000) %>%
  ggplot(aes(x = Year, fill = Data_Type)) +
  geom_histogram(binwidth = 1, alpha = 0.7, position = "identity") +
  labs(title = "Temporal Distribution (2000+)",
       x = "Year", y = "Count") +
  theme_minimal() +
  theme(legend.position = "bottom")

# ------------------------------------------------------------
# 3. BATCH EFFECTS ANALYSIS
# ------------------------------------------------------------

cat("\n=== BATCH EFFECTS ANALYSIS ===\n")

# Function to perform comprehensive comparison
compare_groups <- function(data, var_name) {
  formula <- as.formula(paste(var_name, "~ Data_Type"))
  
  # Descriptive statistics
  desc_stats <- data %>%
    group_by(Data_Type) %>%
    summarise(
      Mean = mean(.data[[var_name]], na.rm = TRUE),
      SD = sd(.data[[var_name]], na.rm = TRUE),
      Median = median(.data[[var_name]], na.rm = TRUE),
      IQR = IQR(.data[[var_name]], na.rm = TRUE),
      N = sum(!is.na(.data[[var_name]])),
      .groups = 'drop'
    )
  
  # Statistical tests
  # 1. T-test (if parametric)
  t_test <- tryCatch({
    t.test(as.formula(paste(var_name, "~ Data_Type")), data = data)
  }, error = function(e) NULL)
  
  # 2. Wilcoxon test (non-parametric)
  wilcox_test <- tryCatch({
    wilcox.test(as.formula(paste(var_name, "~ Data_Type")), data = data)
  }, error = function(e) NULL)
  
  # 3. Cohen's d effect size
  cohen_d <- tryCatch({
    cohens_d(as.formula(paste(var_name, "~ Data_Type")), data = data)
  }, error = function(e) NULL)
  
  # 4. ANOVA with other factors
  anova_test <- tryCatch({
    aov(formula, data = data) %>% summary()
  }, error = function(e) NULL)
  
  return(list(
    variable = var_name,
    descriptive = desc_stats,
    t_test = t_test,
    wilcox_test = wilcox_test,
    cohen_d = cohen_d,
    anova = anova_test
  ))
}

# Analyze key morphological variables
morph_vars <- c("Length_mm", "Width_mm", "Geometric_Mean", "Globosity_WoverL")

batch_effects <- list()
for (var in morph_vars) {
  cat(sprintf("\n--- Analysis for %s ---\n", var))
  result <- compare_groups(Clean, var)
  batch_effects[[var]] <- result
  
  # Print descriptive stats
  print(result$descriptive)
  
  # Print statistical test results
  if (!is.null(result$t_test)) {
    cat(sprintf("\nT-test: t = %.3f, p = %.4f\n", 
                result$t_test$statistic, result$t_test$p.value))
  }
  
  if (!is.null(result$wilcox_test)) {
    cat(sprintf("Wilcoxon test: W = %.0f, p = %.4f\n", 
                result$wilcox_test$statistic, result$wilcox_test$p.value))
  }
  
  if (!is.null(result$cohen_d)) {
    cat(sprintf("Cohen's d: %.3f (Effect size: %s)\n", 
                result$cohen_d$Cohens_d, 
                interpret_cohens_d(result$cohen_d$Cohens_d)))
  }
}

# ------------------------------------------------------------
# 4. NORMALITY ASSESSMENT
# ------------------------------------------------------------

cat("\n=== NORMALITY ASSESSMENT ===\n")

# Function to assess normality
assess_normality <- function(data, var_name) {
  results <- list()
  
  for (type in c("Museum/Field", "iNaturalist")) {
    subset_data <- data %>% 
      filter(Data_Type == type & !is.na(.data[[var_name]]))
    
    if (nrow(subset_data) > 3) {
      # Shapiro-Wilk test
      shapiro <- shapiro.test(subset_data[[var_name]])
      
      # QQ plot
      qq_data <- data.frame(
        Theoretical = qqnorm(subset_data[[var_name]], plot.it = FALSE)$x,
        Sample = qqnorm(subset_data[[var_name]], plot.it = FALSE)$y
      )
      
      results[[type]] <- list(
        n = nrow(subset_data),
        shapiro_W = shapiro$statistic,
        shapiro_p = shapiro$p.value,
        qq_data = qq_data,
        is_normal = shapiro$p.value > 0.05
      )
    }
  }
  return(results)
}

# Create normality assessment plots
normality_plots <- list()
for (var in morph_vars) {
  normality_results <- assess_normality(Clean, var)
  
  # Create density plots
  p_density <- Clean %>%
    filter(!is.na(.data[[var]])) %>%
    ggplot(aes(x = .data[[var]], fill = Data_Type)) +
    geom_density(alpha = 0.5) +
    labs(title = paste("Distribution of", var),
         x = var, y = "Density") +
    theme_minimal() +
    facet_wrap(~ Data_Type, ncol = 2)
  
  # Create QQ plots
  if ("iNaturalist" %in% names(normality_results)) {
    p_qq_iNat <- ggplot(normality_results[["iNaturalist"]]$qq_data, 
                       aes(x = Theoretical, y = Sample)) +
      geom_point(alpha = 0.5) +
      geom_qq_line(aes(sample = Sample), color = "red") +
      labs(title = paste("QQ Plot - iNaturalist:", var),
           x = "Theoretical Quantiles", y = "Sample Quantiles") +
      theme_minimal()
  }
  
  if ("Museum/Field" %in% names(normality_results)) {
    p_qq_museum <- ggplot(normality_results[["Museum/Field"]]$qq_data, 
                         aes(x = Theoretical, y = Sample)) +
      geom_point(alpha = 0.5) +
      geom_qq_line(aes(sample = Sample), color = "red") +
      labs(title = paste("QQ Plot - Museum/Field:", var),
           x = "Theoretical Quantiles", y = "Sample Quantiles") +
      theme_minimal()
  }
  
  normality_plots[[var]] <- list(
    density = p_density,
    qq_iNat = if(exists("p_qq_iNat")) p_qq_iNat else NULL,
    qq_museum = if(exists("p_qq_museum")) p_qq_museum else NULL,
    results = normality_results
  )
  
  # Print normality test results
  cat(sprintf("\n--- Normality tests for %s ---\n", var))
  for (type in names(normality_results)) {
    cat(sprintf("%s: Shapiro-Wilk W = %.4f, p = %.4f (Normal: %s)\n",
                type, 
                normality_results[[type]]$shapiro_W,
                normality_results[[type]]$shapiro_p,
                ifelse(normality_results[[type]]$is_normal, "Yes", "No")))
  }
}

# ------------------------------------------------------------
# 5. DATA TRANSFORMATION OPTIONS
# ------------------------------------------------------------

cat("\n=== DATA TRANSFORMATION OPTIONS ===\n")

transform_data <- function(x) {
  transformations <- list(
    original = x,
    log = log(x + 1),
    sqrt = sqrt(x),
    cube_root = sign(x) * abs(x)^(1/3),
    rank = rank(x),
    arcsin_sqrt = asin(sqrt(x))  # For proportions like Globosity
  )
  return(transformations)
}

# Test transformations for non-normal variables
test_transformations <- function(data, var_name, data_type) {
  subset_data <- data %>% 
    filter(Data_Type == data_type & !is.na(.data[[var_name]])) %>%
    pull(var_name)
  
  if (length(subset_data) > 3) {
    transformations <- transform_data(subset_data)
    
    cat(sprintf("\nTransformations for %s (%s):\n", var_name, data_type))
    
    for (trans_name in names(transformations)) {
      if (all(is.finite(transformations[[trans_name]]))) {
        shapiro <- shapiro.test(transformations[[trans_name]])
        cat(sprintf("  %s: W = %.4f, p = %.4f\n", 
                    trans_name, shapiro$statistic, shapiro$p.value))
      }
    }
  }
}

# Test transformations for each variable showing non-normality
for (var in morph_vars) {
  for (type in c("Museum/Field", "iNaturalist")) {
    test_transformations(Clean, var, type)
  }
}

# ------------------------------------------------------------
# 6. POWER ANALYSIS
# ------------------------------------------------------------

cat("\n=== POWER ANALYSIS ===\n")

# Function to calculate power for different effect sizes
calculate_power <- function(n1, n2, alpha = 0.05) {
  effect_sizes <- c(0.2, 0.5, 0.8)  # Small, medium, large
  
  power_results <- data.frame(
    Effect_Size = effect_sizes,
    Power = sapply(effect_sizes, function(d) {
      pwr.t2n.test(n1 = n1, n2 = n2, d = d, sig.level = alpha)$power
    })
  )
  
  # Calculate detectable effect size with 80% power
  detectable_effect <- pwr.t2n.test(
    n1 = n1, n2 = n2, power = 0.80, sig.level = alpha
  )$d
  
  return(list(
    power_table = power_results,
    detectable_effect = detectable_effect
  ))
}

# Get sample sizes
n_iNat <- sum(Clean$Data_Type == "iNaturalist", na.rm = TRUE)
n_museum <- sum(Clean$Data_Type == "Museum/Field", na.rm = TRUE)

cat(sprintf("\nSample sizes:\n"))
cat(sprintf("iNaturalist: %d specimens\n", n_iNat))
cat(sprintf("Museum/Field: %d specimens\n", n_museum))

# Calculate power
power_analysis <- calculate_power(n_iNat, n_museum)

cat("\nStatistical power for detecting effects:\n")
print(power_analysis$power_table)

cat(sprintf("\nMinimum detectable effect size with 80%% power: %.3f\n", 
            power_analysis$detectable_effect))
cat("Interpretation: ", interpret_cohens_d(power_analysis$detectable_effect), "\n")

# Power for ANOVA (if considering multiple groups within Museum/Field)
if (n_museum > 0) {
  # Assuming 8 museum collections as groups
  anova_power <- pwr.anova.test(
    k = 9,  # 8 museum collections + iNaturalist
    n = min(n_iNat/9, n_museum/8),  # Balanced design approximation
    f = 0.25,  # Medium effect size
    sig.level = 0.05
  )
  
  cat(sprintf("\nANOVA power analysis:\n"))
  cat(sprintf("Groups: 9 (8 museum collections + iNaturalist)\n"))
  cat(sprintf("Approximate sample per group: %.0f\n", min(n_iNat/9, n_museum/8)))
  cat(sprintf("Power to detect medium effect (f=0.25): %.3f\n", anova_power$power))
}

# ------------------------------------------------------------
# 7. SPATIAL AND TEMPORAL OVERLAP ANALYSIS
# ------------------------------------------------------------

cat("\n=== SPATIAL AND TEMPORAL OVERLAP ===\n")

# Temporal overlap calculation
temporal_overlap <- Clean %>%
  filter(!is.na(Year)) %>%
  group_by(Data_Type) %>%
  summarise(
    Year_Range = paste(min(Year), "-", max(Year)),
    Overlap_Start = max(min(Year[Data_Type == "iNaturalist"]), 
                       min(Year[Data_Type == "Museum/Field"])),
    Overlap_End = min(max(Year[Data_Type == "iNaturalist"]), 
                     max(Year[Data_Type == "Museum/Field"])),
    Overlap_Years = ifelse(Overlap_End >= Overlap_Start, 
                          Overlap_End - Overlap_Start + 1, 0),
    .groups = 'drop'
  ) %>%
  distinct()

cat("\nTemporal overlap analysis:\n")
print(temporal_overlap)

# Spatial overlap (by county)
spatial_overlap <- Clean %>%
  filter(!is.na(County) & !is.na(Data_Type)) %>%
  group_by(County, Data_Type) %>%
  summarise(N = n(), .groups = 'drop') %>%
  pivot_wider(names_from = Data_Type, values_from = N, values_fill = 0) %>%
  mutate(
    Total = `Museum/Field` + iNaturalist,
    Overlap_Score = pmin(`Museum/Field`, iNaturalist) / pmax(`Museum/Field`, iNaturalist),
    Overlap_Score = ifelse(is.na(Overlap_Score), 0, Overlap_Score)
  ) %>%
  arrange(desc(Total))

cat("\nSpatial overlap by county:\n")
print(spatial_overlap)

# ------------------------------------------------------------
# 8. COMPREHENSIVE VISUALIZATION
# ------------------------------------------------------------

cat("\n=== GENERATING COMPREHENSIVE VISUALIZATIONS ===\n")

# Create a multi-panel figure
final_plot <- function() {
  # Panel 1: Data composition
  p1 <- ggplot(data_type_summary, aes(x = Data_Type, y = N, fill = Data_Type)) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = paste0(N, " (", Percent, "%)")), 
              vjust = -0.5, size = 3) +
    labs(title = "Sample Sizes by Data Type", 
         x = "Data Type", y = "Number of Specimens") +
    theme_minimal() +
    theme(legend.position = "none")
  
  # Panel 2: Morphological comparison
  comparison_data <- Clean %>%
    select(Data_Type, Length_mm, Width_mm, Globosity_WoverL) %>%
    pivot_longer(cols = c(Length_mm, Width_mm, Globosity_WoverL),
                 names_to = "Variable", values_to = "Value") %>%
    filter(!is.na(Value))
  
  p2 <- ggplot(comparison_data, aes(x = Data_Type, y = Value, fill = Data_Type)) +
    geom_boxplot(alpha = 0.7) +
    facet_wrap(~ Variable, scales = "free_y", ncol = 3) +
    labs(title = "Morphological Comparison by Data Type",
         x = "Data Type", y = "Value") +
    theme_minimal() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Panel 3: Temporal distribution
  p3 <- Clean %>%
    filter(!is.na(Year)) %>%
    ggplot(aes(x = Year, fill = Data_Type)) +
    geom_density(alpha = 0.5) +
    labs(title = "Temporal Distribution Density",
         x = "Year", y = "Density") +
    theme_minimal() +
    theme(legend.position = "bottom")
  
  # Arrange plots
  grid.arrange(p1, p2, p3, ncol = 1, heights = c(1, 2, 1))
}

# Save the comprehensive plot
png("Data_Type_Analysis.png", width = 10, height = 12, units = "in", res = 300)
final_plot()
dev.off()

cat("\n=== ANALYSIS COMPLETE ===\n")
cat("\nKey findings have been saved to:\n")
cat("1. Console output (statistical results)\n")
cat("2. Data_Type_Analysis.png (comprehensive visualization)\n")
cat("3. Individual density and QQ plots for normality assessment\n")

# ------------------------------------------------------------
# 9. SUMMARY RECOMMENDATIONS
# ------------------------------------------------------------

cat("\n=== SUMMARY AND RECOMMENDATIONS ===\n\n")

# Based on analysis, provide recommendations
recommendations <- list(
  "Batch Effects" = if(any(sapply(batch_effects, function(x) 
    !is.null(x$t_test) && x$t_test$p.value < 0.05))) {
    "Significant batch effects detected. Consider including Data_Type as a covariate in analyses."
  } else {
    "No significant batch effects detected between data types."
  },
  
  "Normality" = if(any(sapply(normality_plots, function(x) 
    any(sapply(x$results, function(y) !y$is_normal))))){
    "Some variables show non-normal distributions. Consider data transformations or non-parametric tests."
  } else {
    "Data appears normally distributed for most variables."
  },
  
  "Statistical Power" = if(power_analysis$detectable_effect < 0.5) {
    sprintf("Good statistical power (detectable effect size: %.2f).", 
            power_analysis$detectable_effect)
  } else {
    sprintf("Limited power for small effects (detectable effect size: %.2f).", 
            power_analysis$detectable_effect)
  },
  
  "Temporal Comparability" = if(temporal_overlap$Overlap_Years[1] > 10) {
    sprintf("Good temporal overlap (%d years). iNaturalist data is comparable to museum data.", 
            temporal_overlap$Overlap_Years[1])
  } else {
    "Limited temporal overlap. Consider temporal stratification in analyses."
  }
)

for (i in seq_along(recommendations)) {
  cat(names(recommendations)[i], ": ", recommendations[[i]], "\n\n")
}
```

Updates to plots
```{r}
# OPTIMIZED ANALYSIS SCRIPT FOR C. californicus MORPHOLOGY
# Using Clean dataset only with DRY principles

# ==============================================================================
# 0. SETUP AND DATA LOADING
# ==============================================================================

# Install and load required packages (run once)
required_packages <- c("tidyverse", "PerformanceAnalytics", "GGally", 
                       "cowplot", "ggsignif", "ggpubr", "broom", "effectsize")
new_packages <- required_packages[!(required_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

# Load all packages
library(tidyverse)
library(PerformanceAnalytics)
library(GGally)
library(cowplot)
library(ggsignif)
library(ggpubr)
library(broom)
library(effectsize)

# Load Clean dataset directly
Clean <- read.csv("https://raw.githubusercontent.com/sabahzero/C_californicus_Morphology/main/SupplementalTable3_Data-Clean.csv", 
                  stringsAsFactors = FALSE)

# Ensure numeric types for key variables
Clean <- Clean %>%
  mutate(
    Length_mm = as.numeric(Length_mm),
    Width_mm = as.numeric(Width_mm),
    Geometric_Mean = as.numeric(Geometric_Mean),
    Globosity_WoverL = as.numeric(Globosity_WoverL),
    Mean_SST = as.numeric(Mean_SST),
    Latitude = as.numeric(Latitude),
    Year = as.numeric(Year),
    Decade = as.numeric(Decade),
    Epoch = factor(Epoch, levels = c("Pleistocene", "Holocene", "Anthropocene")),
    Data_Type = ifelse(Collection == "iNat", "iNaturalist", "Museum/Field")
  )

# ==============================================================================
# 1. REUSABLE FUNCTIONS
# ==============================================================================

# Function for consistent plotting aesthetics
epoch_colors <- c("Pleistocene" = "#00AFBB", 
                  "Holocene" = "#E7B800", 
                  "Anthropocene" = "#FC4E07")

theme_custom <- function(base_size = 11) {
  theme_minimal(base_size = base_size) +
    theme(
      plot.title = element_text(size = base_size + 3, face = "bold"),
      axis.title = element_text(face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
}

# Function for regression plots with consistent formatting
ggplotRegression <- function(fit, xlab = NULL, ylab = NULL) {
  require(ggplot2)
  
  coef_summary <- summary(fit)$coefficients
  slope <- ifelse(nrow(coef_summary) > 1, coef_summary[2, 1], NA)
  p_value <- ifelse(nrow(coef_summary) > 1, coef_summary[2, 4], NA)
  
  ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
    geom_point(alpha = 0.4, size = 1.5) +
    stat_smooth(method = "lm", col = "red", se = FALSE) +
    labs(
      title = sprintf("R² = %.3f, Slope = %.3f, p = %.2e", 
                      summary(fit)$adj.r.squared, slope, p_value),
      x = ifelse(is.null(xlab), names(fit$model)[2], xlab),
      y = ifelse(is.null(ylab), names(fit$model)[1], ylab)
    ) +
    theme_custom()
}

# Function for correlation matrix plots
plot_correlation_matrix <- function(data, variables, filename = NULL) {
  corr_matrix <- chart.Correlation(data %>% select(all_of(variables)))
  
  if (!is.null(filename)) {
    dev.copy(png, filename, width = 10, height = 10, units = "in", res = 300)
    dev.off()
  }
  
  return(corr_matrix)
}

# Function for epoch-specific boxplots
create_epoch_boxplot <- function(data, x_var, y_var, fill_var = NULL, 
                                 y_limits = NULL, title = "", 
                                 filename = NULL) {
  
  plot <- ggplot(data, aes_string(x = x_var, y = y_var)) +
    geom_boxplot(aes_string(fill = ifelse(is.null(fill_var), "Epoch", fill_var)), 
                 alpha = 0.7) +
    scale_fill_manual(values = if(is.null(fill_var)) epoch_colors else "Set2") +
    labs(title = title, x = x_var, y = y_var) +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  if (!is.null(y_limits)) {
    plot <- plot + coord_cartesian(ylim = y_limits)
  }
  
  if (!is.null(filename)) {
    ggsave(filename, plot, width = 8, height = 6, dpi = 300)
  }
  
  return(plot)
}

# ==============================================================================
# 2. DATA OVERVIEW AND CORRELATION ANALYSIS (Figure 2b)
# ==============================================================================

cat("=== DATA OVERVIEW ===\n")
cat("Total specimens:", nrow(Clean), "\n")
cat("Collections:", paste(sort(unique(Clean$Collection)), collapse = ", "), "\n")
cat("Epoch distribution:\n")
print(table(Clean$Epoch))

# Define morphological variables
morph_vars <- c("Length_mm", "Width_mm", "Geometric_Mean", "Globosity_WoverL")

# Figure 2b: Correlation matrix
Fig2b_data <- Clean %>%
  select(Epoch, all_of(morph_vars)) %>%
  rename(
    "Length (size)" = Length_mm,
    "Width (size)" = Width_mm,
    "geomM (mass)" = Geometric_Mean,
    "Globosity (shape)" = Globosity_WoverL
  )

# Custom function for scatter plots with regression
my_fn <- function(data, mapping, ...) {
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point(alpha = 0.6, size = 1) + 
    geom_smooth(method = lm, color = "black", se = FALSE, ...)
  p
}

Fig2b_chart <- ggpairs(
  Fig2b_data,
  aes(color = Epoch, alpha = 0.6),
  columns = 2:5,
  lower = list(continuous = my_fn),
  upper = list(continuous = wrap("cor", size = 3)),
  progress = FALSE
) +
  scale_color_manual(values = epoch_colors) +
  scale_fill_manual(values = epoch_colors) +
  theme_classic()

ggsave("Figure2b_Shell-Allometry.png", Fig2b_chart, width = 10, height = 10, dpi = 300)

# Statistical correlations
correlation_tests <- list(
  "Length~Width" = lm(Length_mm ~ Width_mm, data = Clean),
  "Length~Globosity" = lm(Length_mm ~ Globosity_WoverL, data = Clean),
  "Width~Globosity" = lm(Width_mm ~ Globosity_WoverL, data = Clean),
  "Length~Geometric_Mean" = lm(Length_mm ~ Geometric_Mean, data = Clean),
  "Width~Geometric_Mean" = lm(Width_mm ~ Geometric_Mean, data = Clean),
  "Globosity~Geometric_Mean" = lm(Globosity_WoverL ~ Geometric_Mean, data = Clean)
)

cat("\n=== CORRELATION STATISTICS ===\n")
for (i in seq_along(correlation_tests)) {
  test <- correlation_tests[[i]]
  cat(sprintf("%s: R² = %.4f, p = %.2e\n", 
              names(correlation_tests)[i],
              summary(test)$adj.r.squared,
              summary(test)$coefficients[2, 4]))
}

# ==============================================================================
# 3. LATITUDE AND SST ANALYSIS (Figure 3 & Supplemental Figure 1)
# ==============================================================================

# Supplemental Figure 1: Latitude vs SST
SuppFig1 <- ggplot(Clean, aes(x = Latitude, y = Mean_SST, color = Epoch)) +
  annotate("rect", ymin = -Inf, ymax = Inf, xmin = 32.56, xmax = 34.49,
           alpha = 0.2, fill = "gray") +
  geom_point(alpha = 0.4, size = 1) +
  geom_smooth(aes(fill = Epoch), method = lm, se = TRUE, alpha = 0.2) +
  geom_smooth(method = lm, color = "black", se = FALSE) +
  scale_color_manual(values = epoch_colors) +
  scale_fill_manual(values = epoch_colors) +
  labs(x = "Latitude", y = "Mean SST (°C)") +
  theme_custom()

ggsave("SuppFig1_Latitude-SST.png", SuppFig1, width = 8, height = 6, dpi = 300)

# Function for latitude/SST scatter plots
create_scatter_plot <- function(data, x_var, y_var, 
                                x_label = NULL, y_label = NULL,
                                show_epoch_regression = TRUE) {
  
  plot <- ggplot(data, aes_string(x = x_var, y = y_var, color = "Epoch")) +
    geom_point(alpha = 0.4, size = 1) +
    scale_color_manual(values = epoch_colors) +
    labs(x = ifelse(is.null(x_label), x_var, x_label),
         y = ifelse(is.null(y_label), y_var, y_label)) +
    theme_custom() +
    theme(legend.position = "none")
  
  if (show_epoch_regression) {
    plot <- plot + 
      geom_smooth(aes(fill = Epoch), method = lm, se = TRUE, alpha = 0.2) +
      scale_fill_manual(values = epoch_colors)
  }
  
  plot <- plot + geom_smooth(method = lm, color = "black", se = FALSE)
  
  return(plot)
}

# Figure 3a-d: Size and shape vs Latitude and SST
Fig3a <- create_scatter_plot(Clean, "Latitude", "Length_mm", 
                             "Latitude", "Size (Length, mm)")
Fig3b <- create_scatter_plot(Clean, "Mean_SST", "Length_mm",
                             "Mean SST (°C)", "Size (Length, mm)")
Fig3c <- create_scatter_plot(Clean, "Latitude", "Globosity_WoverL",
                             "Latitude", "Shape (Globosity, W/L)")
Fig3d <- create_scatter_plot(Clean, "Mean_SST", "Globosity_WoverL",
                             "Mean SST (°C)", "Shape (Globosity, W/L)")

# Combine plots
Figure3 <- plot_grid(Fig3a, Fig3b, Fig3c, Fig3d, 
                     labels = c("A", "B", "C", "D"), 
                     label_size = 12, ncol = 2)
ggsave("Figure3_Size-Shape-Lat-SST.png", Figure3, width = 12, height = 10, dpi = 300)

# ==============================================================================
# 4. ECOREGION AND COASTAL PROXIMITY ANALYSIS (Figure 4 & Supplemental Figure 2)
# ==============================================================================

# Supplemental Figure 2: SST by ecoregion and proximity
create_sst_boxplot <- function(epoch_data, title) {
  ggplot(epoch_data, aes(x = Ecoregion, y = Mean_SST, fill = Coastal_Proximity)) +
    geom_boxplot(alpha = 0.7) +
    coord_cartesian(ylim = c(9.5, 20)) +
    labs(title = title, x = "Ecoregion", y = "Mean SST (°C)") +
    scale_fill_brewer(palette = "Set2") +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

SuppFig2a <- create_sst_boxplot(filter(Clean, Epoch == "Pleistocene"), "(a) Pleistocene")
SuppFig2b <- create_sst_boxplot(filter(Clean, Epoch == "Holocene"), "(b) Holocene")
SuppFig2c <- create_sst_boxplot(filter(Clean, Epoch == "Anthropocene"), "(c) Anthropocene")

# Save supplemental figures
ggsave("SuppFig2a_SST-Pleistocene.png", SuppFig2a, width = 8, height = 6, dpi = 300)
ggsave("SuppFig2b_SST-Holocene.png", SuppFig2b, width = 8, height = 6, dpi = 300)
ggsave("SuppFig2c_SST-Anthropocene.png", SuppFig2c, width = 8, height = 6, dpi = 300)

# Figure 4: Morphology by ecoregion and proximity
create_morphology_boxplot <- function(epoch_data, y_var, y_label, 
                                      y_limits = NULL, title = "") {
  
  plot <- ggplot(epoch_data, aes(x = Ecoregion, y = .data[[y_var]], 
                                 fill = Coastal_Proximity)) +
    geom_boxplot(alpha = 0.7) +
    labs(title = title, x = "Ecoregion", y = y_label) +
    scale_fill_brewer(palette = "Set2") +
    theme_custom() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  if (!is.null(y_limits)) {
    plot <- plot + coord_cartesian(ylim = y_limits)
  }
  
  return(plot)
}

# Create all Figure 4 plots
fig4_plots <- list()
epochs <- c("Pleistocene", "Holocene", "Anthropocene")
variables <- list(
  list(var = "Geometric_Mean", label = "Mass (Geometric Mean)", limits = c(3, 24)),
  list(var = "Globosity_WoverL", label = "Shape (Globosity, W/L)", limits = c(0.4, 0.72))
)

for (i in seq_along(epochs)) {
  epoch_data <- filter(Clean, Epoch == epochs[i])
  
  for (j in seq_along(variables)) {
    var_info <- variables[[j]]
    plot_index <- (i - 1) * 2 + j
    
    fig4_plots[[plot_index]] <- create_morphology_boxplot(
      epoch_data, 
      var_info$var, 
      var_info$label,
      var_info$limits,
      paste0("(", letters[plot_index], ") ", epochs[i])
    )
  }
}

# Combine and save Figure 4
Figure4 <- plot_grid(plotlist = fig4_plots, ncol = 2, labels = "AUTO", label_size = 12)
ggsave("Figure4_Ecoregion-Proximity.png", Figure4, width = 12, height = 14, dpi = 300)

# ==============================================================================
# 5. STATISTICAL MODELS (Tables 1-2)
# ==============================================================================

# Function to fit and summarize models
fit_model <- function(formula, data, model_name = "") {
  model <- lm(formula, data = data)
  cat(sprintf("\n=== %s ===\n", model_name))
  print(summary(model))
  return(model)
}

# Table 1: Latitude and SST models
cat("\n" + strrep("=", 60) + "\n")
cat("TABLE 1: LATITUDE AND SST MODELS\n")
cat(strrep("=", 60) + "\n")

table1_models <- list(
  "Width ~ Latitude + Epoch + Latitude:Epoch" = 
    lm(Width_mm ~ Latitude + Epoch + Latitude:Epoch, data = Clean),
  "Width ~ Mean_SST + Epoch + Mean_SST:Epoch" = 
    lm(Width_mm ~ Mean_SST + Epoch + Mean_SST:Epoch, data = Clean),
  "Globosity ~ Latitude + Epoch + Latitude:Epoch" = 
    lm(Globosity_WoverL ~ Latitude + Epoch + Latitude:Epoch, data = Clean),
  "Globosity ~ Mean_SST + Epoch + Mean_SST:Epoch" = 
    lm(Globosity_WoverL ~ Mean_SST + Epoch + Mean_SST:Epoch, data = Clean)
)

for (i in seq_along(table1_models)) {
  cat(sprintf("\n%s:\n", names(table1_models)[i]))
  print(summary(table1_models[[i]]))
}

# Table 2: Ecoregion and proximity models
cat("\n" + strrep("=", 60) + "\n")
cat("TABLE 2: ECOREGION AND PROXIMITY MODELS\n")
cat(strrep("=", 60) + "\n")

table2_models <- list()
for (epoch in epochs) {
  epoch_data <- filter(Clean, Epoch == epoch)
  
  table2_models[[paste0(epoch, "_Width")]] <- 
    lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion:Coastal_Proximity, 
       data = epoch_data)
  
  table2_models[[paste0(epoch, "_Globosity")]] <- 
    lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion:Coastal_Proximity, 
       data = epoch_data)
}

for (i in seq_along(table2_models)) {
  cat(sprintf("\n%s:\n", names(table2_models)[i]))
  print(summary(table2_models[[i]]))
}

# ==============================================================================
# 6. HUMAN POPULATION DENSITY ANALYSIS (Supplemental Figure 3)
# ==============================================================================

# Prepare data for human population analysis
SuppFig3_data <- Clean %>%
  filter(Epoch != "Pleistocene", Collection != "PR", Population_Density > 0) %>%
  mutate(
    Human_pop = cut(Population_Density,
                    breaks = quantile(Population_Density, probs = c(0, 0.33, 0.66, 1)),
                    labels = c("Low", "Medium", "High"),
                    include.lowest = TRUE)
  )

# Create population density plots
create_population_plot <- function(data, y_var, y_label) {
  ggplot(data, aes(x = Decade, y = .data[[y_var]], color = Human_pop)) +
    geom_point(alpha = 0.4, size = 1) +
    geom_smooth(aes(fill = Human_pop), method = lm, se = TRUE, alpha = 0.2) +
    geom_smooth(method = lm, color = "black", se = FALSE) +
    labs(x = "Decade", y = y_label) +
    scale_color_brewer(palette = "Set1") +
    scale_fill_brewer(palette = "Set1") +
    theme_custom()
}

SuppFig3a <- create_population_plot(SuppFig3_data, "Width_mm", "Size (Width, mm)")
SuppFig3b <- create_population_plot(SuppFig3_data, "Globosity_WoverL", "Shape (Globosity, W/L)")

ggsave("SuppFig3a_Population-Width.png", SuppFig3a, width = 8, height = 6, dpi = 300)
ggsave("SuppFig3b_Population-Globosity.png", SuppFig3b, width = 8, height = 6, dpi = 300)

# ==============================================================================
# 7. COPE'S RULE ANALYSIS
# ==============================================================================

# Prepare Cope's Rule data
Copes_data <- Clean %>%
  select(Location, Epoch, Width_mm, Length_mm, Globosity_WoverL) %>%
  drop_na()

# Function for Cope's Rule bar plots
create_copes_plot <- function(data, y_var, y_label, group_by_location = FALSE) {
  
  plot_data <- if (group_by_location) {
    data %>%
      group_by(Epoch, Location) %>%
      summarise(
        mean_value = mean(.data[[y_var]], na.rm = TRUE),
        se = sd(.data[[y_var]], na.rm = TRUE) / sqrt(n()),
        .groups = "drop"
      )
  } else {
    data %>%
      group_by(Epoch) %>%
      summarise(
        mean_value = mean(.data[[y_var]], na.rm = TRUE),
        se = sd(.data[[y_var]], na.rm = TRUE) / sqrt(n()),
        .groups = "drop"
      )
  }
  
  plot <- ggplot(plot_data, aes(x = if(group_by_location) Location else Epoch, 
                                y = mean_value, fill = Epoch)) +
    geom_bar(stat = "identity", position = position_dodge(0.9)) +
    geom_errorbar(aes(ymin = mean_value - se, ymax = mean_value + se),
                  width = 0.2, position = position_dodge(0.9)) +
    labs(x = if(group_by_location) "Location" else "Epoch", y = y_label) +
    scale_fill_manual(values = epoch_colors) +
    theme_custom() +
    theme(axis.text.x = element_text(angle = if(group_by_location) 45 else 0, hjust = 1))
  
  return(plot)
}

# Create Cope's Rule plots
Copes_Width <- create_copes_plot(Copes_data, "Width_mm", "Width (mm)")
Copes_Width_byLoc <- create_copes_plot(Copes_data, "Width_mm", "Width (mm)", TRUE)
Copes_Globosity <- create_copes_plot(Copes_data, "Globosity_WoverL", "Globosity (W/L)")
Copes_Globosity_byLoc <- create_copes_plot(Copes_data, "Globosity_WoverL", "Globosity (W/L)", TRUE)

# Save Cope's Rule plots
ggsave("Copes_Width_Epoch.png", Copes_Width, width = 8, height = 6, dpi = 300)
ggsave("Copes_Width_byLocation.png", Copes_Width_byLoc, width = 10, height = 6, dpi = 300)
ggsave("Copes_Globosity_Epoch.png", Copes_Globosity, width = 8, height = 6, dpi = 300)
ggsave("Copes_Globosity_byLocation.png", Copes_Globosity_byLoc, width = 10, height = 6, dpi = 300)

# ==============================================================================
# 8. SUMMARY AND EXPORT
# ==============================================================================

cat("\n" + strrep("=", 60) + "\n")
cat("ANALYSIS COMPLETE\n")
cat(strrep("=", 60) + "\n")

# Summary statistics
summary_stats <- Clean %>%
  group_by(Epoch, Data_Type) %>%
  summarise(
    N = n(),
    Mean_Length = mean(Length_mm, na.rm = TRUE),
    SD_Length = sd(Length_mm, na.rm = TRUE),
    Mean_Width = mean(Width_mm, na.rm = TRUE),
    SD_Width = sd(Width_mm, na.rm = TRUE),
    Mean_Globosity = mean(Globosity_WoverL, na.rm = TRUE),
    SD_Globosity = sd(Globosity_WoverL, na.rm = TRUE),
    .groups = "drop"
  )

cat("\nSummary Statistics by Epoch and Data Type:\n")
print(summary_stats, n = Inf)

# List generated files
cat("\nGenerated Files:\n")
cat("1. Figure2b_Shell-Allometry.png - Correlation matrix\n")
cat("2. Figure3_Size-Shape-Lat-SST.png - Latitude/SST relationships\n")
cat("3. Figure4_Ecoregion-Proximity.png - Ecoregion/proximity analysis\n")
cat("4. SuppFig1_Latitude-SST.png - Supplemental latitude vs SST\n")
cat("5. SuppFig2a/b/c_SST-*.png - Supplemental SST by ecoregion\n")
cat("6. SuppFig3a/b_Population-*.png - Human population effects\n")
cat("7. Copes_*.png - Cope's Rule analysis\n")

# Save summary data
write.csv(summary_stats, "Summary_Statistics.csv", row.names = FALSE)
cat("\nSummary statistics saved to: Summary_Statistics.csv\n")

cat("\n=== END OF ANALYSIS ===\n")
```

Prior
Data Clean-up        Post-processing: Raw -> Clean + SoCal       
*Fig 1a-b            Map of C. californicus range                
*Fig 2b              Shell allometry (Fig 2a of shell manual)             
*Fig 3a-d,Supp Fig 1 Size or Shape and Lat / Temp, Bergmann's    
*Fig 4a-f,Supp Fig 2 Size or Shape and Ecoregion + Proxmity      
*Tables 1-2          Statistics breakdown (of all figures)      
*Supp Fig 3          Correlations to human population density    
*Cope's Rule         Size by location ~ Epoch                    

Overview of data and post-processing
```{r}

###### Upload data ######
# Need package for downloading data from Github repo
## install.packages("RCurl")
library("RCurl") 
## packageVersion("RCurl") # v1.98.1.1 

Raw <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable2_Data-Raw.csv")) 

Key <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable1_Data-Key.csv")) 

# Ensure blanks are read as NAs
###### Upload data ######


###### Data structure ######
# Quick view of RawData
str(Raw) 

# Plots of shell morphology measurements as a point of reference for outliers 
morphos = c("Length_mm", "Width_mm", "Aperture_mm", "Thickness_mm", "Globosity_WoverL", "Geometric_Mean")

# Need package for correlation charts
## install.packages("PerformanceAnalytics")
library("PerformanceAnalytics") 
## packageVersion("PerformanceAnalytics") # v2.0.4

# Need package pipes ( %>% or "then" )
## install.packages("tidyverse")
library("tidyverse") 
## packageVersion("tidyverse") # v1.3.0

# Check for outliers by correlation matrix
Raw_CorrMatrix = chart.Correlation(Raw %>% 
                                     select(morphos)) 
Raw_CorrMatrix
###### Data structure ######


###### Clean-up ######
# Organization of locations in dataset
## Re-order counties, North to South
counties = c("Sonoma", "San Mateo", "Monterey", "San Luis Obispo", "Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego", "Baja California")
Raw$County = factor(Raw$County, levels=counties) # order counties
## Order SoCal counties and locales for when subsetting SoCal data
socal = c("Santa Barbara", "Ventura", "Los Angeles", "Orange", "San Diego")
socal_locations = c("SB", "NLA", "LA", "SLA", "SD", "ANA", "SCZI", "SRI", "SMI", "CAT", "SBI", "SNI", "SCI") # socal location order
la = "Los Angeles"

# Filter outliers (min/max values identified by Raw correlations)
## Subset for including back in as part of Clean / filtered set (below)
No_TA = subset(Raw, is.na(Raw$Thickness_mm)) # 775 no T or A
Raw_iNat = subset(Raw, Raw$Collection=="iNat") # 99 from iNat
## Confirmed to be operator error (measuring or datasheet input)
OutliersRemoved = Raw %>% # 2996 obs 
  filter(Thickness_mm < 2) %>% # low T
  filter(!(Width_mm < 15 & Thickness_mm > 2)) %>% # too thick for W
  filter(Aperture_mm < 6) %>%  # small A
  filter(!(Length_mm > 20 & Aperture_mm <2)) %>% # A too small for L
  rbind(No_TA) %>%  # adds 775 shells back in 
  filter(Width_mm > 2) %>%  # too narrow (note this gets rid of iNat)
  rbind(Raw_iNat) %>% # add iNat back in
  filter(Globosity_WoverL > 0.42 & Globosity_WoverL < 0.71 ) %>% 
  data.frame # ensure it's a data frame
## Remove 25 shells from Pliocene (too old, few)
OutliersRemoved=subset(OutliersRemoved, !Epoch=="Pliocene")
## Filter specimens that may skew in being in the open ocean and/or too north or south
min(OutliersRemoved$Latitude) # 26.43
max(OutliersRemoved$Latitude) # 38.16
### Ensure updated Clean dataset has 3 or more specimens per lot
Clean = OutliersRemoved %>% 
  filter(Latitude > 26.43) %>% 
  filter(Latitude < 38.16) %>% 
  group_by(Lot_Number) %>% 
  mutate(Specimens_per_Lot = length((Lot_Number))) %>%
  ungroup %>% 
  filter(Specimens_per_Lot > 2 | Collection == "iNat") # keep catalogs with > 2 in lot and iNat (1 per 'lot')
## Validation check for NA values for final data set
cat("\n\n",
    sum(is.na(Clean$Globosity_WoverL)), "have no globosity ratio,\n", # 0
    sum(is.na(Clean$Geometric_Mean)), "have no geometric mean,\n", # 99 [iNat]
    sum(is.na(Clean$Width_mm)), "have no width measurement,\n", # 99 [iNat]
    sum(is.na(Clean$Length_mm)), "have no length measurement,\n", # 99 [iNat]
    sum(is.na(Clean$Thickness_mm)), "have no thickness measurement,\n", # 742 = 99 [iNat] + 338 PR + 273 Pleisto
    sum(is.na(Clean$Aperture_mm)), "have no aperture measurement" ) # 742 = 99 [iNat] + 338 PR + 273 Pleisto
## Order the data
Clean %>% 
  group_by(Lot_Number) %>% 
  summarise(N = length(Specimen_Number))
Clean %>% 
  group_by(Epoch) %>% 
  summarise(N = length(Specimen_Number)) 
## Re-do of correlation chart
Clean_CorrMatrix = chart.Correlation(Clean %>% 
                                       select(morphos)) # much improved
Clean_CorrMatrix
dev.copy(png,'Figure2b_Clean.jpg')
dev.off()

# Subset Clean (2908 of 2996 from Raw) to SoCal (2608)
SoCal = Clean %>% 
  filter(County %in% 
           socal) 
SoCal$Location = factor(SoCal$Location, levels=socal_locations) # order 
# Distribution of shells by county *most are in LA (1372)
SoCal %>% 
  group_by(County) %>% 
  summarise(N = length(Epoch)) %>% 
  data.frame 
# Check correlation for SoCal subset
SoCal_CorrMatrix = chart.Correlation(SoCal %>% 
                                       select(morphos)) # Also looks good
SoCal_CorrMatrix
dev.copy(png,'Figure2b_SoCal.jpg')
dev.off()

# Subset 
LA = SoCal %>% 
  filter(County %in% 
           la) 
LA$Location = factor(LA$Location, levels=socal_locations) # order 
# Check correlation for LA subset
LA_CorrMatrix = chart.Correlation(LA %>%
                                    select(morphos)) #
LA_CorrMatrix
dev.copy(png,'Figure2b_LA.jpg')
dev.off()
###### Clean-up ######



###### Save cleaned-up data for downstream visualization and analysis ######
# Data structure for Clean and SoCal subset
sort(unique(Clean$Collection)) # Unique collections: 9
sort(unique(Clean$Lot_Number)) # Unique lots: 382 
sort(unique(Clean$County)) # Unique counties: 9
sort(unique(Clean$Location)) # Unique locations: 19
sort(unique(Clean$Coastal_Proximity)) # Unique location types: 2
sort(unique(Clean$Habitat)) # Unique habitats: 4
sort(unique(Clean$Depth)) # Unique depths: 3

sort(unique(SoCal$Collection)) # Unique collections: 9
sort(unique(SoCal$Lot_Number)) # Unique lots: 339
sort(unique(SoCal$County)) # Unique counties: 4
sort(unique(SoCal$Location)) # Unique locations: 13

write.csv(Clean, file = "SupplementalTable3_Data-Clean.csv")
write.csv(SoCal, file = "SupplementalTable4_Data-SoCal.csv") 

# Validation check 
Clean <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable3_Data-Clean.csv")) 
SoCal <- read.csv(text=getURL("https://raw.githubusercontent.com/MichaelMalloy/C_californicus_Morphology/master/SupplementalTable4_Data-SoCal.csv")) 


###### Save cleaned-up data for downstream visualization and analysis ######

# Subset data by Epoch to color code points
Anthro = subset(Clean, Epoch=="Anthropocene") # 1492
Holo = subset(Clean, Epoch=="Holocene") # 1106
Pleisto = subset(Clean, Epoch=="Pleistocene") # 305

Anthro_SoCal = subset(SoCal, Epoch=="Anthropocene") # 1366
Holo_SoCal = subset(SoCal, Epoch=="Holocene") # 1020
Pleisto_SoCal = subset(SoCal, Epoch=="Pleistocene") # 211

```

Figure 1a-b: Maps of shell locations (Clean and SoCal Data) 
```{r}

# Need packages for map coordinates
## if(!requireNamespace("devtools")) install.packages("devtools")
## devtools::install_github("dkahle/ggmap")
library("ggmap")
## packageVersion("ggmap") # v3.0.0
## install.packages("rgdal")
library("rgdal")
## packageVersion("rgdal") # v1.4.8
# Need package for plots of maps
## install.packages("ggplot2")
library("ggplot2")
## packageVersion("ggplot2") # v3.2.1
# Need package for gridding maps as one figure
## install.packages("cowplot")
library("cowplot")
## packageVersion("cowplot") # v1.0.0

# This code is personalized by account and only runs if you have a key
## Instructions: https://github.com/dkahle/ggmap
## Visit your personal Google Maps Platform to retrieve key if you have one
## ggmap::register_google(key = " ") 

# Load polygons for ecoregions and call them after downloaded in your folder 
## shape files from http://www.marineregions.org/ (also in repo)
NC_poly=readOGR( "northern_ecoregions" , layer="ecoregions") 
SCB_poly=readOGR( "southern_ecoregions" , layer="ecoregions") 
MT_poly=readOGR( "magdaelena_ecoregions" , layer="ecoregions") 

# Figure 1a
## Parameters for All points
All_bbox <- make_bbox(lat = Latitude, lon = Longitude, data = Clean)
All_bbox
##     left        bottom    right       top 
##     -122.786    27.359    -114.734    37.721 
All=get_map(location=c(-125,20,-105,40), # adjust to include regions
            zoom = 5, 
            maptype = "terrain") 
## Create box for zoom of SoCal
SoCal_box = data.frame(x=c(-121,-121,-116.5,-116.5,-121), 
       y=c(35,32,32,35,35)) 
## Plot map
Clean_map = ggmap(All) +
  geom_polygon(data = NC_poly, aes(x = long, y = lat),
               fill = "white", alpha = 0.5) + # white
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) + # blue
  geom_polygon(data = MT_poly, aes(x = long, y = lat),
               fill = "#666666", alpha = 0.5) + # grey
  stat_density_2d(data = Clean, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # color fill, black
  geom_point(data = Anthro, aes(x = Longitude, y = Latitude), 
             color = "#FC4E07", # orange
             alpha = 0.15,
             size = 1) + 
  geom_point(data = Holo, aes(x = Longitude,y = Latitude), 
             color = "#E7B800", # gold
             alpha = 0.15,
             size = 1) + 
  geom_point(data = Pleisto, aes(x = Longitude, y = Latitude), 
             color = "#00AFBB", # blue
             alpha = 0.15,
             size = 1) + 
  geom_point(data = Key, aes(x = Longitude, y = Latitude), # asbsent data
             color = "#FC4E07", # orange
             shape = 4,
             size = 1.5) + # manually add Punta Abreojos white circle
  geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude") 
Clean_map

# Figure 1b
SoCal_bbox = make_bbox(lat = Latitude, lon = Longitude, data = SoCal)
SoCal_bbox
##    left        bottom    right       top
##    -120.6685   32.4635 -116.9615   34.5865 

SoCal=get_map(location=c(-121,32,-116.5,35), 
            zoom = 7, 
            maptype = "satellite") 

SoCal$Latitude = as.numeric(SoCal$Latitude)
SoCal$Longitude = as.numeric(SoCal$Longitude)

SoCal = data.frame(SoCal)

SoCal_map = ggmap(SoCal) +
  stat_density_2d(data = SoCal, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .35,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + 
  geom_point(data = Anthro_SoCal, aes(x = Longitude, y = Latitude), 
             color = "#FC4E07",
             alpha = 0.15,
             size = 2) + 
  geom_point(data = Holo_SoCal, aes(x = Longitude,y = Latitude), 
             color = "#E7B800", 
             alpha = 0.15,
             size = 2) + 
  geom_point(data = Pleisto_SoCal, aes(x = Longitude, y = Latitude), 
             color = "#00AFBB",
             alpha = 0.15,
             size = 2) + 
    geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude")
SoCal_map


ggsave("Figure1b_Data-Clean-SoCal_Map.png", dpi = 800)


ggsave("Figure1a_Data-Clean_Map.png", dpi = 800)

##### Create Map (Figure 1a-b) #####
#  Need a personalized account with API key (map key) for Google maps
## Instructions: https://github.com/dkahle/ggmap
## ggmap::register_google(key = " ") 
### More helpful info: https://www.littlemissdata.com/blog/maps 

# Figure 1a: Larger map
## Parameters for All points
All_bbox <- make_bbox(lat = Latitude, lon = Longitude, data = df)
All_bbox # Informs us what to set for parameters
##     left        bottom    right       top 
##     -122.786    27.359    -114.734    37.721 
All=get_map(location=c(-129,20,-105,42), # adjust to include regions
            zoom = 5, 
            maptype = "terrain") # can adjust this type as well
## Create box for zoom of SoCal
SoCal_box = data.frame(x=c(-121,-121,-116.5,-116.5,-121), 
       y=c(35,32,32,35,35)) 
## Plot map 
Figure1a = ggmap(All) + # Google map of parameters
  geom_polygon(data = NC_poly, aes(x = long, y = lat), # polygons (shape files)
               fill = "white", alpha = 0.5) + 
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) +
  geom_polygon(data = MT_poly, aes(x = long, y = lat),
               fill = "#666666", alpha = 0.5) + # grey
  stat_density_2d(data = df, # hypothetical population density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250, # can adjust this for resolution
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # fills density layers as black 
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) + # colors by Epoch
  geom_point(data = df, aes(x = Longitude, y = Latitude, color = Epoch), 
             alpha = 0.15,
             size = 1) + # layers points
  geom_path(data=SoCal_box, aes(x,y), color="black", lwd=1) + # insert box
  labs(x="Longitude", y="Latitude") +
  theme(legend.position="none")
Figure1a

# Figure 1b: Zoom-in of SoCal
SoCal=get_map(location=c(-121,32,-116.5,35), # adjust based on SoCal_box
            maptype = "terrain-background") # can adjust this type as well
## Plot map
Figure1b = ggmap(SoCal) +
  geom_polygon(data = NC_poly, aes(x = long, y = lat),
               fill = "white", alpha = 0.5) + # white
  geom_polygon(data = SCB_poly, aes(x = long, y = lat),
               fill = "blue", alpha = 0.3) + # blue
    stat_density_2d(data = df, # density
                  aes(x = Longitude,
                      y = Latitude,
                      fill = stat(level)),
                  alpha = .25,
                  bins = 250,
                  geom = "polygon") +
  scale_fill_gradientn(colors = "black") + # color fill, black
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_point(data = df, aes(x = Longitude, y = Latitude, color = Epoch), 
             alpha = 0.15,
             size = 1) +
  theme(axis.title.x=element_blank(), axis.title.y=element_blank())
Figure1b # Note the ecoregions cut-out if polygon not fully included

# Grid using cowplot, and save
Figure1 = plot_grid(Figure1a, Figure1b, labels = c("A", "B"), label_size=12)
Figure1   

ggsave("Figure1.png", dpi = 800) 

##### Create Map (Figure 1a-b) #####

```

Figure 2b: Are shell size, mass, or shape independent or dependent metrics of each other?
```{r}

# install.packages("GGally")
library(GGally)
# packageVersion("GGally") # v1.4.0

# Create a function for including linear regression line
my_fn <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    geom_smooth(method=lm, color="black", ...) 
  p
}

# Figure 2b
Fig2b = Clean[,c("Epoch", "Length_mm", "Width_mm", "Globosity_WoverL", "Geometric_Mean")] 
# Rename columns
names(Fig2b) = c("Epoch", "Length (size)", "Width (size)", "Globosity (shape)", "geomM (mass)") 
# Create figure (automatically excludes NAs)
Fig2b_chart = ggpairs(Fig2b, aes(color = Epoch, alpha = 0.85), lower = list(continuous = my_fn), upper = list(continuous = wrap("cor", size = 4.25)))
# Change colors manually with a loop through each plot
for(i in 1:Fig2b_chart$nrow) {
  for(j in 1:Fig2b_chart$ncol){
    Fig2b_chart[i,j] <- Fig2b_chart[i,j] + 
        scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
        scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB"))  
  }
}

Fig2b_chart + 
  theme_classic() # theme for figure
ggsave("Figure2b_Data-Clean-New_Shell-Allometry.png", dpi = 800)

# Statistics
names(Fig2b) = c("Epoch", "Length", "Width", "Globosity", "geomM") 
summary(lm(Length ~ Width, data = Fig2b)) # Radj = 0.9534, p < 2.2e-16
summary(lm(Length ~ Globosity, data = Fig2b)) # Radj = 0.0853, p < 2.2e-16
summary(lm(Width ~ Globosity, data = Fig2b)) # Radj = 0.007822, p = 2.796e-06 
summary(lm(Length ~ geomM, data = Fig2b)) # Radj = 0.9884, p < 2.2e-16
summary(lm(Width ~ geomM, data = Fig2b)) # Radj = 0.988, p < 2.2e-16
summary(lm(Globosity ~ geomM, data = Fig2b)) # Radj = 0.03698, p < 2.2e-16

# Length = Width, dependent on each other and synonymous with size
# Globosity (shape) is an independent metric of size 
## Width slightly more independent (horizontal line to globosity) than Length
### Thus, use Width as proxy for size 
# geomM (mass) is synonymous with size in this instance, and independent of shape
## Do we want to then only use geomM (mass) and globosity (shape)?
### Note Width is still the most horizontal of the 3 (Length, Width, geomM) for Fig 2b ()

```

Q2) Do size/shape vary by epoch in relation to latitude or temperature? Yes.
Figure 3a-d, Supp Figure 1: Comparison of size and shape for Latitude vs SST (Clean)
```{r}

# Regression Plot function for y = mx + b
ggplotRegression <- function (fit) {

require(ggplot2)

ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                     "Intercept =",signif(fit$coef[[1]],5 ),
                     " Slope =",signif(fit$coef[[2]], 5),
                     " P =",signif(summary(fit)$coef[2,4], 5)))
}

# Supplemental Figure 1: What is the relationship between latitude and SST?
SuppFig1 = ggplot(data = Clean, 
                      aes(x = Latitude, y = Mean_SST,
                      color = Epoch,
                      alpha = 0.4)) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.2, fill="black") + # shade square
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Latitude", y="Mean SST")
SuppFig1 +
  theme_classic() 
ggsave("SuppFig1_Data-Clean-New_Latitude-MeanSST.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Latitude, data = Clean))
# ** y = 38.222 - 0.660x, r^2 = 0.157, p = 5.302 x 10^-110
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean, Epoch=="Anthropocene")))
# ** y = 43.679 - 0.810x, r^2 = 0.413, p = 1.487 x 10^-174
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean, Epoch=="Holocene")))
# ** y = 35.303 - 0.565x, r^2 = 0.259, p = 7.868 x 10^-74
ggplotRegression(lm(Mean_SST ~ Latitude, data = subset(Clean, Epoch=="Pleistocene")))
# ** y = 32.702 - 0.591, r^2 = 0.457, p = 1.704 x 10^-42

# Figure 3a-d (size is represented as Width) 
Fig3a = ggplot(data = Clean, 
                      aes(x = Latitude, y = Length_mm,
                      color = Epoch,
                      alpha = 0.4)) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.2, fill="black") + # shade square
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Latitude", y="Size (Length mm)") +
  theme(legend.position = "none") 
# Regression by Epoch
ggplotRegression(lm(Latitude ~ Length_mm, data = Clean))
# ** y = 32.837 + 0.057x, r^2 = 0.043, p = 1.056 x 10^-28
## Relationships by Epoch are not important based on global interactive model results (Tab 1)

Fig3b = ggplot(data = Clean, 
                      aes(x = Mean_SST, y = Length_mm,
                      color = Epoch,
                      alpha = 0.4)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Mean SST", y="") +
  theme(legend.position = "none") 
# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Length_mm, data = Clean))
# y = 16.050 + 0.005x, r^2 = 2.267 x 10^-4, p = 0.545
ggplotRegression(lm(Mean_SST ~ Length_mm, data = subset(Clean, Epoch=="Anthropocene")))
# y = 16.663 - 7.955x10^-5x, r^2 = -0.001, p = 0.993
ggplotRegression(lm(Mean_SST ~ Length_mm, data = subset(Clean, Epoch=="Holocene")))
# ** y = 17.009 - 0.065x, r^2 = 0.057, p = 7.759 x 10^-26
ggplotRegression(lm(Mean_SST ~ Length_mm, data = subset(Clean, Epoch=="Pleistocene")))
# y = 13.191 - 0.021x, r^2 = 0.001, p = 0.255

Figure3 = plot_grid(Fig3a , Fig3b, labels = c("A", "B"), label_size=12)
ggsave("Fig3_Length.png", dpi = 800)








Fig3c = ggplot(data = Clean, 
                      aes(x = Latitude, y = Globosity_WoverL,
                      color = Epoch,
                      alpha = 0.4)) +
  annotate("rect",ymin=-Inf,ymax=Inf,xmin=32.56,xmax=34.49, # adjust by transformation
            alpha=0.2, fill="black") + # shade square
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Latitude", y="Shape (Globosity, W/L)")
Fig3c +
  theme_classic() 
ggsave("Fig3c_Data-Clean_Latitude-Globosity.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = Clean))
# ** y = 31.951 + 2.645x, r^2 = 0.013, p = 8.232e-10
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean, Epoch=="Anthropocene")))
# ** y = 31.235 + 3.662x, r^2 = 0.028, p = 5.554e-11
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean, Epoch=="Holocene")))
# y = 33.297 + 0.697x, r^2 = -1.745e-5, p = 0.322
ggplotRegression(lm(Latitude ~ Globosity_WoverL, data = subset(Clean, Epoch=="Pleistocene")))
# ** y = 36.22 - 5.014x, r^2 = 0.025, p = 0.004

Fig3d = ggplot(data = Clean, 
                      aes(x = Mean_SST, y = Globosity_WoverL,
                      color = Epoch,
                      alpha = 0.4)) +
  scale_color_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  scale_fill_manual(values=c("#FC4E07", "#E7B800", "#00AFBB")) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Epoch)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(aes(color = Epoch, fill = Epoch), method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Mean SST", y="Shape (Globosity, W/L)")
Fig3d +
  theme_classic() 
ggsave("Fig3d_Data-Clean_MeanSST-Globosity.png", dpi = 800)
# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = Clean))
# ** y = 17.072 - 1.619x, r^2 = 0.001, p = 0.024
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean, Epoch=="Anthropocene")))
# ** y = 21.187 - 7.901x, r^2 = 0.083, p = 4.988e-30
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean, Epoch=="Holocene")))
# ** y = 13.424 + 4.866x, r^2 = 0.035, p = 3.155e-10
ggplotRegression(lm(Mean_SST ~ Globosity_WoverL, data = subset(Clean, Epoch=="Pleistocene")))
# ** y = 13.21 - 0.237x, r^2 = -0.003, p = 0.875




ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean))) # p = 0.923
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean, Epoch=="Anthropocene"))) # p < 0.0001
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean, Epoch=="Holocene")))
# p < 0.0001
ggplotRegression(lm(Mean_SST ~ Coastal_Proximity, data = subset(Clean, Epoch=="Pleistocene"))) # p = 0.03


# Regression by Epoch
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = Clean))
# 
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = subset(Clean, Epoch=="Anthropocene")))
#
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = subset(Clean, Epoch=="Holocene")))
# **
ggplotRegression(lm(Mean_SST ~ Geometric_Mean, data = subset(Clean, Epoch=="Pleistocene")))
# 

```

Q3) Do size/shape vary in relation to ecoregion and coastal proximity? Yes.
Figure 4a-f, Supp Fig 2: Size vs shape for ecoregion and coastal proximity
(Clean_New)
```{r}

# install.packages("ggplot2")
library(ggplot2)
# packageVersion("ggplot2") # v3.1.0

# Supplemental Figure 2: What is the relationship between ecoregion, coastal proximity, and SST?
min(Clean_New$Mean_SST) # 9.68
max(Clean_New$Mean_SST) # 19.52831

SuppFig2a <- ggplot(subset(Clean_New, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Mean_SST, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(9.5,20)) +
  scale_y_continuous(name = "Mean SST") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(a) Pleistocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
SuppFig2a
ggsave("SuppFig2a_Data-Clean-New_Ecoregion-Proximity-MeanSST.png", dpi = 800)

SuppFig2b <- ggplot(subset(Clean_New, Epoch=="Holocene"), aes(x = Ecoregion, y = Mean_SST, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(9.5,20)) +
  scale_y_continuous(name = "Mean SST") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(b) Holocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
SuppFig2b
ggsave("SuppFig2b_Data-Clean-New_Ecoregion-Proximity-MeanSST.png", dpi = 800)

SuppFig2c <- ggplot(subset(Clean_New, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Mean_SST, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(9.5,20)) +
  scale_y_continuous(name = "Mean SST") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(c) Anthropocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
SuppFig2c
ggsave("SuppFig2c_Data-Clean-New_Ecoregion-Proximity-MeanSST.png", dpi = 800)

# Figure 4a-f
min(Clean_New$Width_mm) # 3.23
max(Clean_New$Width_mm) # 23.65
min(Clean_New$Globosity_WoverL) # 0.43
max(Clean_New$Globosity_WoverL) # 0.7









library("ggsignif")
library("ggpubr")

Fig4a <- ggplot(subset(Clean, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Geometric_Mean, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(3,24)) +
  scale_y_continuous(name = "Mass (Geometric Mean)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(a) Pleistocene") +
  geom_signif(comparisons=list("Ecoregion","Coastal_Proximity"), map_signif_level=TRUE) + 
  stat_compare_means(method = "t.test") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4a


Fig4b <- ggplot(subset(Clean, Epoch=="Holocene"), aes(x = Ecoregion, y = Geometric_Mean, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(3,24)) +
  scale_y_continuous(name = "Size (Width mm)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(b) Holocene") +
  geom_signif(comparisons=list("Ecoregion"), map_signif_level=TRUE) + 
  stat_compare_means(method = "t.test") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4b
ggsave("Fig4b_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

Fig4c <- ggplot(subset(Clean, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Geometric_Mean, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(3,24)) +
  scale_y_continuous(name = "Size (Width mm)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(c) Anthropocene") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4c
ggsave("Fig4c_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

Fig4d <- ggplot(subset(Clean, Epoch=="Pleistocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(.4,.72)) +
  scale_y_continuous(name = "Shape (Globosity, W/L)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(d)") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4d
ggsave("Fig4d_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

Fig4e <- ggplot(subset(Clean, Epoch=="Holocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(.4,.72)) +
  scale_y_continuous(name = "Shape (Globosity, W/L)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(e)") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4e
ggsave("Fig4e_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)

Fig4f <- ggplot(subset(Clean, Epoch=="Anthropocene"), aes(x = Ecoregion, y = Globosity_WoverL, fill = Coastal_Proximity)) +
  geom_boxplot(alpha=0.7) +
  coord_cartesian(ylim = c(.4,.72)) +
  scale_y_continuous(name = "Shape (Globosity, W/L)") +
  scale_x_discrete(name = "Ecoregion") +
  ggtitle("(f)") +
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
        text = element_text(size = 12, family = "Tahoma"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) 
Fig4f
ggsave("Fig4f_Data-Clean-New_Ecoregion-Proximity-Width.png", dpi = 800)



Fig5a = plot_grid(sys.plot, dia.plot, labels = "a", label_size=12)
Fig5a   

```

Tables 1-2: Statistics, namely for Figs 3-4 (Clean_New)
```{r}

# Table 1 (for Fig 3)
Clean_Numeric <- Clean_New # This is to have lm() Epoch show up as one instead of breakdown
Clean_Numeric$Epoch<-as.numeric(Clean_Numeric$Epoch)

Fig3a_aov <- aov(Width_mm ~ Latitude + Epoch + Latitude*Epoch, data = Clean_New)
Fig3a_GIM <- lm(Width_mm ~ Latitude + Epoch + Latitude*Epoch, data = Clean_Numeric)

Fig3b_aov <- aov(Width_mm ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_New)
Fig3b_GIM <- lm(Width_mm ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_Numeric)

Fig3c_aov <- aov(Globosity_WoverL ~ Latitude + Epoch + Latitude*Epoch, data = Clean_New)
Fig3c_GIM <- lm(Globosity_WoverL ~ Latitude + Epoch + Latitude*Epoch, data = Clean_Numeric)

Fig3d_aov <- aov(Globosity_WoverL ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_New)
Fig3d_GIM <- lm(Globosity_WoverL ~ Mean_SST + Epoch + Mean_SST*Epoch, data = Clean_Numeric)

# For Supp Fig 2
SuppFig2a_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # interaction by ecoregion only

SuppFig2b_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
SuppFig2b_GIM <- lm(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant for all

SuppFig2c_aov <- aov(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
SuppFig2c_GIM <- lm(Mean_SST ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant for all

# Table 2 (for Fig 4)
Clean_Numeric <- Clean_New # Need to re-do for filter by Epoch
Clean_Numeric$Ecoregion <- as.numeric(Clean_Numeric$Ecoregion)
Clean_Numeric$Coastal_Proximity <- as.numeric(Clean_Numeric$Coastal_Proximity)

Fig4a_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # no significance
Fig4a_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Pleistocene"))

Fig4b_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
Fig4b_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Holocene")) # significant for all

Fig4c_aov <- aov(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
Fig4c_GIM <- lm(Width_mm ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Anthropocene")) # significant for all

Fig4d_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Pleistocene")) # no significance
Fig4d_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Pleistocene")) # no significance

Fig4e_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Holocene")) # significant interaction
Fig4e_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Holocene")) # significant for all

Fig4f_aov <- aov(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_New, Epoch=="Anthropocene")) # significant interaction
Fig4f_GIM <- lm(Globosity_WoverL ~ Ecoregion + Coastal_Proximity + Ecoregion*Coastal_Proximity, data = subset(Clean_Numeric, Epoch=="Anthropocene")) # no significance between ecoregions

# Tables 1, 2
Tab1 <- list(Fig3a_GIM, Fig3b_GIM, Fig3c_GIM, Fig3d_GIM)
Tab2 <- list(Fig4a_GIM, Fig4b_GIM, Fig4c_GIM, Fig4d_GIM, Fig4e_GIM, Fig4f_GIM) 

# Print summaries of each respective table 
for(i in Tab1){
 
 print(summary(i))
  # r <- summary(i)$adj.r.squared
 # print(r)
}

for(i in Tab2){
 
 print(summary(i))
  # r <- summary(i)$adj.r.squared
 # print(r)
}

```

Supplemental Figure 3: Relationship to human population density
```{r}

SuppFig3 = subset(Clean_New, Epoch!="Pleistocene")
SuppFig3 = subset(SuppFig3, Collection!="PR") # Removed personal research (biases SD)
SuppFig3 = subset(SuppFig3, Population_Density > 0) # 1349 observations

# Create a column that converts population density to categories by standard deviation
min(SuppFig3$Population_Density) # 0.00932179
x <- mean(SuppFig3$Population_Density) # 247.2658
SD <- sd(SuppFig3$Population_Density) # 546.268
max(SuppFig3$Population_Density) # 2708.81

SuppFig3$Human_pop[SuppFig3$Population_Density < (617-546)] <- "Low"
SuppFig3$Human_pop[SuppFig3$Population_Density > (617+546)] <- "High"
SuppFig3$Human_pop[SuppFig3$Population_Density > (617-545) & 
                     SuppFig3$Population_Density < (617+545)] <- "Medium"

# Supp Fig 3a-b 
SuppFig3a = ggplot(data = SuppFig3, 
                      aes(x = Decade, y = Width_mm,
                      color = Human_pop,
                      alpha = 0.4)) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Human_pop)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Decade", y="Size (Width mm)")
SuppFig3a +
  theme_classic() 
ggsave("SuppFig3a_Data-Clean_New_Decade-Width-Human.png", dpi = 800)
# Regression by category
ggplotRegression(lm(Width_mm ~ Decade, data = SuppFig3))
# ** y = 114.07 - 0.053x, r^2 = 0.076, p = 2.118e-23
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="Low")))
# ** y = 147.5 - 0.070x, r^2 = 0.162, p = 1.87e-38
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="Medium")))
# ** y = -197.93 + 0.107x, r^2 = 0.059, p = 0.00014
ggplotRegression(lm(Width_mm ~ Decade, data = subset(SuppFig3, Human_pop=="High")))
# ** y = 583.62 - 0.291x, r^2 = 0.394, p = 1.309e-10

SuppFig3b = ggplot(data = SuppFig3, 
                      aes(x = Decade, y = Globosity_WoverL,
                      color = Human_pop,
                      alpha = 0.4)) +
  geom_smooth(se = FALSE, method = lm) +
  geom_smooth(method = lm, color = "black") + # 95% confidence
  geom_point(aes(color = Human_pop)) + 
  geom_errorbar(stat="summary", fun.data="mean_se", fun.args = list(mult = 1.96), width=0) +
    geom_smooth(method=lm, se = TRUE, alpha = 0.35) + # 95% confidence interval
  labs(x="Decade", y="Globosity, W/L")
SuppFig3b +
  theme_classic() 
ggsave("SuppFig3b_Data-Clean_New_Decade-Globosity-Human.png", dpi = 800)

ggplotRegression(lm(Globosity_WoverL ~ Decade, data = SuppFig3))
# y = 0.444 + 6.99e-5x, r^2 = 0.002, p = 0.074
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="Low")))
# ** y = 0.019x - 2.88e-4x, r^2 = 0.028, p = 8.486e-8
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="Medium")))
# ** y = 1.858 - 6.49e-4x, r^2 = 0.101, p = 4.326e-7
ggplotRegression(lm(Globosity_WoverL ~ Decade, data = subset(SuppFig3, Human_pop=="High")))
# y = 0.795 - 1.11e-4x, r^2 = 0.002, p = 0.249

```

Cope's Rule
```{r}

Copes <- Clean_New[c("Location", "Epoch", "Width_mm", "Length_mm", "Globosity_WoverL")]
Copes <- Copes[complete.cases(Copes), ] # removes NAs from all rows to only compare where all measurements are available

All_Locales <- lm(Width_mm ~ Location + Epoch + Location*Epoch , data = Clean_New) # interaction by ecoregion only
summary(All_Locales)
Merged <- aov(Width_mm ~ Epoch , data = Clean_New) # interaction by ecoregion only
summary(Merged) # yes, significant by size

# bar plot
# width, locations merge by epoch
Copes_new <- Copes %>% select("Location", "Epoch", "Width_mm") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Width_mm) + sd(Width_mm), # upper limit for error bar
         lower = mean(Width_mm) - sd(Width_mm)) # lower limit for error bar


Width <- ggplot(Copes_new, aes(x = Epoch, y = Width_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Width of the error bars
                position = position_dodge(.9))
ggsave("Copes_Widthmm_Epoch-All.png", dpi = 800)

# width, locations separated out
Copes_new <- Copes %>% select("Location", "Epoch", "Width_mm") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Width_mm) + sd(Width_mm), # upper limit for error bar
         lower = mean(Width_mm) - sd(Width_mm)) # lower limit for error bar

Width <- ggplot(Copes_new, aes(x = Location, y = Width_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Width of the error bars
                position = position_dodge(.9))
Width
ggsave("Copes_Widthmm_Epoch-by-Location.png", dpi = 800)


# Length
Copes_new <- Copes %>% select("Location", "Epoch", "Length_mm") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Length_mm) + sd(Length_mm), # upper limit for error bar
         lower = mean(Length_mm) - sd(Length_mm)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Epoch, y = Length_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
ggsave("Copes_Lengthmm_Epoch-All.png", dpi = 800)

# Length, locations separated out
Copes_new <- Copes %>% select("Location", "Epoch", "Length_mm") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Length_mm) + sd(Length_mm), # upper limit for error bar
         lower = mean(Length_mm) - sd(Length_mm)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Location, y = Length_mm, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
Length
ggsave("Copes_Lengthmm_Epoch-by-Location.png", dpi = 800)

# Globosity
Copes_new <- Copes %>% select("Location", "Epoch", "Globosity_WoverL") %>%
  group_by(Epoch) %>%            # group data for limit calculation
  mutate(upper = mean(Globosity_WoverL) + sd(Globosity_WoverL), # upper limit for error bar
         lower = mean(Globosity_WoverL) - sd(Globosity_WoverL)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Epoch, y = Globosity_WoverL, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
ggsave("Copes_Globosity_Epoch-All.png", dpi = 800)

Copes_new <- Copes %>% select("Location", "Epoch", "Globosity_WoverL") %>%
  group_by(Epoch, Location) %>%            # group data for limit calculation
  mutate(upper = mean(Globosity_WoverL) + sd(Globosity_WoverL), # upper limit for error bar
         lower = mean(Globosity_WoverL) - sd(Globosity_WoverL)) # lower limit for error bar

Length <- ggplot(Copes_new, aes(x = Location, y = Globosity_WoverL, fill = Epoch)) +
  stat_summary(fun.y = mean, geom = "bar", position = position_dodge(width = .9),
               size = 3) +
  geom_errorbar(aes(ymin = lower, ymax = upper),
                width = .2,                    # Length of the error bars
                position = position_dodge(.9))
Length
ggsave("Copes_Globosity_Epoch-by-Location.png", dpi = 800)

```
